---
title: "Week 1 screening summary"
subtitle: "Task 1 — confirmed case-study stocks, SAG status, and advice vs catch"
author: "Laurence Kell\\Sea++"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
  word_document:
    toc: true
    fig_width: 7
    fig_height: 5
---

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
root  ="C:/active/blueMarine"

stocks=read.csv(file.path(root,"data/reference/stocks.csv"), stringsAsFactors=FALSE)
stocks$name=factor(stocks$name, levels=unique(stocks$name))
sids  =as.character(stocks$sid)

suppressPackageStartupMessages(suppressWarnings(library(knitr)))

dir.create(file.path(root, "tex/figs"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "data/results"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "Rmd/cache", "01_screening"), recursive = TRUE, showWarnings = FALSE)
knitr::opts_chunk$set(
  echo      = FALSE, 
  warning   = FALSE, 
  message   = FALSE, 
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(root, "Rmd/cache", "01_screening/"),
  fig.path  = file.path(root, "tex/figs", "01_screening-"),
  fig.width = 9, 
  fig.height= 5, 
  dev       ="png")

Fig=0
Tab=0
```

```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(ggplotFL)
  library(ggplot2)
  library(FLBacktest)
  library(kableExtra)
  library(jsonlite)
  library(FLasher)
})
source(file.path(root, "R/plot_theme.R"))

theme_set(theme_bluemarine_html())
```

```{r data}
source("C:/active/blueMarine/R/loadStocks.R")

oms = FLStocks(dlply(stocks, .(sid), function(x) loadFLStock(x,file.path(root,"data/WGCSE"))))
oms = FLStocks(llply(oms, cleanStock))

inventory = stockInventory(stocks, oms)
```

# Purpose

Week~1 contract deliverable: screen candidate ICES Category~1 stocks, summarise
SAG status and advice vs catch, compare local `FLStock` objects with SAG, and
**confirm the case-study list** for the backtest.

# Confirmed case-study stocks

Local WGCSE `FLStock` objects from `data/WGCSE/`, windowed to each stock's
advice year.


**Table `r Tab = Tab + 1; Tab`**. Confirmed case-study stocks. Display
`name`, ICES `sid`, WGCSE `rdata` file, and the R `object` inside that file
come from `data/reference/stocks.csv`. Advice year is the terminal year used
to window the `FLStock`.

```{r stocks-inventory}
knitr::kable(
  inventory[, c("name", "sid", "rdata", "object")],
  col.names = c("Stock", "SID", "RData", "Object"),
  row.names = FALSE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed"))

knitr::kable(
  inventory[, c("name", "sid", "ices_area", "end", "minyear", "maxyear",
                "minage", "maxage")],
  col.names = c("Stock", "SID", "ICES area", "Advice",
                "Min", "Max", "Min", "Max"),
  row.names = FALSE,
  align = c("l", "l", "l", rep("r", 5))) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed")) %>%
  add_header_above(c(" " = 3, "Year" = 3, "Age" = 2))
```

**Decision.** Proceed with all six stocks:

1. Irish Sea cod (`cod.27.7a`)
2. Irish Sea whiting (`whg.27.7a`)
3. Celtic Sea cod (`cod.27.7e-k`)
4. Celtic Sea whiting (`whg.27.7b-ce-k`)
5. Celtic Sea haddock (`had.27.7b-k`)
6. NE Atlantic mackerel (`mac.27.nea`)

Irish Sea stocks use slightly earlier advice years than the Celtic Sea gadoids,
reflecting available local assessment objects. Mackerel remains included as the
Coastal States / ICES MSY-rule case study.

# Screening summary

Screening uses local assessments in `data/WGCSE/` and multi-year SAG downloads
(assessment years 2014–2026) for status and reference points.

## SAG coverage

```{r load-sag}
assYears = 2014:2026
cache = file.path(root, "data/interim/sag_stocks.RData")
dir.create(dirname(cache), recursive = TRUE, showWarnings = FALSE)

if (file.exists(cache)) {
  load(cache)
} else {
  sag = getSag(sids, assYears)
  save(sag, file = cache)
}

sag$ts    = subset(sag$ts,    sid %in% sids)
sag$rfpts = subset(sag$rfpts, sid %in% sids)

sag_cov = ddply(sag$ts, .(sid), summarise,
                assYears  = paste(range(assYear, na.rm = TRUE), collapse = "-"),
                n_assYear = length(unique(assYear)))
sag_cov = merge(stocks[, c("name", "sid")], sag_cov, by = "sid", all.x = TRUE)
sag_cov = sag_cov[match(sids, sag_cov$sid), ]
```

**Table `r Tab = Tab + 1; Tab`**. SAG assessment-year coverage.

```{r sag-coverage}
knitr::kable(
  sag_cov[, c("name", "sid", "assYears", "n_assYear")],
  col.names = c("Stock", "SID", "Assessment years", "n"),
  row.names = FALSE)
```

Celtic Sea gadoids and Irish Sea cod typically have about ten assessment years;
Irish Sea whiting and mackerel have slightly fewer but still sufficient for
screening. No stock was dropped solely for SAG incompleteness.

## ICES advice vs reported catch

Advised catch from the ICES Advice and Scenarios Database; reported catch from
the latest SAG assessment year per stock.

```{r asd-advice, fig.height=10, fig.width=10}
advice = ldply(sids, function(sid) {
  url = paste0("https://asd.ices.dk/API/getAdviceViewRecord?StockCode=",
               URLencode(sid, reserved = TRUE))
  d = tryCatch(fromJSON(url), error = function(e) NULL)
  if (is.null(d) || !NROW(d)) return(NULL)
  d = as.data.frame(d, stringsAsFactors = FALSE)
  data.frame(
    sid        = sid,
    assYear    = as.integer(d$assessmentYear),
    adviceYear = as.integer(substr(as.character(d$adviceApplicableFrom), 1, 4)),
    advice     = suppressWarnings(as.numeric(d$adviceValue)),
    stringsAsFactors = FALSE)
})
advice = merge(advice, stocks[, c("sid", "name")], by = "sid")
advice = subset(advice, is.finite(adviceYear) & is.finite(advice))

latest_ass = aggregate(assYear ~ sid, data = sag$ts, FUN = max)
reported = merge(sag$ts, latest_ass, by = c("sid", "assYear"))
reported = merge(reported, stocks[, c("sid", "name")], by = "sid")

advice_plot = rbind(
  data.frame(sid=advice$sid, year = advice$adviceYear,
             series = "ICES advice", data = advice$advice,
             stringsAsFactors = FALSE),
  data.frame(sid=reported$sid, year = reported$year,
             series = "Reported catch", data = reported$catch,
             stringsAsFactors = FALSE))
advice_plot       =subset(advice_plot, year >= 2010 & is.finite(data))
advice_plot       =merge(advice_plot,stocks,by="sid")
advice_plot$name  =factor(advice_plot$name, levels = levels(stocks$name))
advice_plot$series=factor(advice_plot$series,
                            levels = c("Reported catch", "ICES advice"))

ggplot(advice_plot, aes(year, data, colour = series, linetype = series)) +
  geom_line(linewidth = 0.85) +
  geom_point(size = 1.2, alpha = 0.9) +
  scale_catch_advice_colour() +
  scale_catch_advice_linetype() +
  facets_stock_row() +
  labs(x = NULL, y = "Catch (t)", colour = NULL, linetype = NULL)

```

**Figure `r Fig = Fig + 1; Fig`**. ICES advised catch (ASD) and reported catch (SAG).

## Historical SSB retrospectives

```{r sag-ssb, fig.height=10, fig.width=10}
ts_plot = merge(sag$ts, stocks[, c("sid", "name")], by = "sid")
ts_plot = subset(ts_plot, is.finite(ssb))
ts_plot$name = factor(ts_plot$name, levels = levels(stocks$name))

p_ssb = ggplot(ts_plot, aes(year, ssb, group = assYear, colour = factor(assYear))) +
  geom_line(linewidth = 0.65, alpha = 0.9) +
  facet_grid(name ~ ., scales = "free_y") +
  labs(x = NULL, y = "SSB", colour = "Assessment") +
  theme(legend.position = "bottom")
p_ssb

dir.create(file.path(root, "tex/figs"), recursive = TRUE, showWarnings = FALSE)
ggsave(file.path(root, "tex/figs/week1_sag_ssb.pdf"), p_ssb,
       width = 10, height = 8)
ggsave(file.path(root, "tex/figs/week1_sag_ssb.png"), p_ssb,
       width = 10, height = 8, dpi = 150)
```

**Figure `r Fig = Fig + 1; Fig`**. SAG SSB by assessment year (retrospective fan).

## Reference-point stability

Reference points relative to the latest assessment year per stock
(values near 1 indicate little revision).

```{r sag-rfpts, fig.height=10, fig.width=11}
rp_vars = intersect(c("fpa", "bpa", "blim", "fmsy", "msybtrigger"),
                    names(sag$rfpts))
rp_lbl  = c(fpa = "F[pa]", bpa = "B[pa]", blim = "B[lim]",
            fmsy = "F[MSY]", msybtrigger = "MSYB[trigger]")

rfs = do.call(rbind, lapply(rp_vars, function(v) {
  data.frame(
    sid = sag$rfpts$sid,
    assYear = sag$rfpts$assYear,
    qname = v,
    data = sag$rfpts[[v]],
    stringsAsFactors = FALSE)
}))
rfs = subset(rfs, is.finite(data) & data > 0)
rfs = merge(rfs, stocks[, c("sid", "name")], by = "sid")
rfs$name = factor(rfs$name, levels = levels(stocks$name))

latest = aggregate(assYear ~ sid + qname, data = rfs, FUN = max)
names(latest)[3] = "assYear_max"
rfs = merge(rfs, latest, by = c("sid", "qname"))
rfs_lat = subset(rfs, assYear == assYear_max)[, c("sid", "qname", "data")]
names(rfs_lat)[3] = "data_latest"
rfs = merge(rfs, rfs_lat, by = c("sid", "qname"))
rfs$rel = rfs$data / rfs$data_latest
rfs$qname = factor(as.character(rfs$qname), levels = rp_vars)

ggplot(rfs, aes(assYear, rel)) +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey50") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.4, alpha = 0.85) +
  facet_grid(name ~ qname, scales = "free_y",
             labeller = labeller(qname = as_labeller(rp_lbl, label_parsed))) +
  labs(x = "Assessment year", y = "Relative to latest")
```

**Figure `r Fig = Fig + 1; Fig`**. SAG reference points relative to the latest assessment.

Across stocks, reference points are relatively stable once scaled to the latest
assessment, with larger revisions mainly where benchmarks or advice frameworks
changed.

## Latest stock status

**Table `r Tab = Tab + 1; Tab`**. Latest SAG terminal year.
SSB: green above \(B_{\mathrm{trigger}}\), amber between \(B_{\lim}\) and
\(B_{\mathrm{trigger}}\), red below \(B_{\lim}\);
\(F/F_{\mathrm{MSY}}\): green at or below 1, red above.

```{r sag-latest, results='asis'}
latest_ass = aggregate(assYear ~ sid, data = sag$ts, FUN = max)
ts_now = merge(sag$ts, latest_ass, by = c("sid", "assYear"))
sag_now = do.call(rbind, lapply(split(ts_now, ts_now$sid), function(d) {
  d = d[order(d$year), ]
  ok = is.finite(d$ssb) & is.finite(d$f)
  if (!any(ok)) ok = is.finite(d$ssb)
  if (!any(ok)) ok = seq_len(nrow(d))
  d[max(which(ok)),
    c("sid", "assYear", "year", "ssb", "f", "rec", "catch"),
    drop = FALSE]
}))
rownames(sag_now) = NULL

rp_now = merge(sag$rfpts, latest_ass, by = c("sid", "assYear"))
keep = intersect(c("fmsy", "blim", "msybtrigger", "fpa", "bpa"), names(rp_now))
if (length(keep))
  sag_now = merge(sag_now, rp_now[, c("sid", keep), drop = FALSE],
                  by = "sid", all.x = TRUE)

tab = merge(stocks[, c("name", "sid")], sag_now, by = "sid", all.x = TRUE)
tab = tab[match(sids, tab$sid), ]
tab$ssb_blim  = tab$ssb / tab$blim
tab$ssb_btrig = tab$ssb / tab$msybtrigger
tab$f_fmsy    = tab$f / tab$fmsy

fmt_num = function(x, digits = 0)
  ifelse(is.finite(x), formatC(x, format = "f", digits = digits), "")

ssb_ratio_bg = ifelse(is.finite(tab$ssb_blim) & tab$ssb_blim < 1, "#F4CCCC",
               ifelse(is.finite(tab$ssb_btrig) & tab$ssb_btrig < 1, "#FCE5CD",
               ifelse(is.finite(tab$ssb_btrig), "#D9EAD3", "#FFFFFF")))
ssb_blim_bg  = ifelse(!is.finite(tab$ssb_blim), "#FFFFFF",
               ifelse(tab$ssb_blim < 1, "#F4CCCC", "#D9EAD3"))
f_ratio_bg   = ifelse(!is.finite(tab$f_fmsy), "#FFFFFF",
               ifelse(tab$f_fmsy > 1, "#F4CCCC", "#D9EAD3"))
ssb_bg = ifelse(is.finite(tab$ssb) & is.finite(tab$blim) & tab$ssb < tab$blim,
                "#F4CCCC",
         ifelse(is.finite(tab$ssb) & is.finite(tab$msybtrigger) &
                  tab$ssb < tab$msybtrigger,
                "#FCE5CD",
         ifelse(is.finite(tab$ssb), "#D9EAD3", "#FFFFFF")))

out = data.frame(
  name = tab$name,
  assYear = tab$assYear,
  year = tab$year,
  ssb_btrig = fmt_num(tab$ssb_btrig, 2),
  ssb_blim  = fmt_num(tab$ssb_blim, 2),
  f_fmsy    = fmt_num(tab$f_fmsy, 2),
  ssb = fmt_num(tab$ssb, 0),
  f   = fmt_num(tab$f, 3),
  fmsy = fmt_num(tab$fmsy, 3),
  blim = fmt_num(tab$blim, 0),
  msybtrigger = fmt_num(tab$msybtrigger, 0),
  stringsAsFactors = FALSE)

print(knitr::kable(
  out, format = "html", escape = TRUE,
  col.names = c("Stock", "Assessment", "Year",
                "SSB/Btrig", "SSB/Blim", "F/FMSY",
                "SSB", "F", "FMSY", "Blim", "MSYBtrigger"),
  row.names = FALSE,
  align = c("l", "r", "r", rep("r", 8))) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "condensed")) %>%
  column_spec(4, background = ssb_ratio_bg) %>%
  column_spec(5, background = ssb_blim_bg) %>%
  column_spec(6, background = f_ratio_bg) %>%
  column_spec(7, background = ssb_bg) %>%
  column_spec(8, background = f_ratio_bg))
```

**Screening findings (status).**

- **Celtic Sea cod** remains depleted (SSB well below \(B_{\lim}\)) with
  \(F\) far above \(F_{\mathrm{MSY}}\) — a priority backtest case.
- **Celtic Sea haddock** and **mackerel** are below \(B_{\mathrm{trigger}}\)
  with \(F > F_{\mathrm{MSY}}\).
- **Celtic Sea whiting** is below \(B_{\mathrm{trigger}}\) but with
  \(F < F_{\mathrm{MSY}}\) in the latest SAG year.
- **Irish Sea cod** is below \(B_{\lim}\) with very low recent \(F\) in SAG.
- **Irish Sea whiting** SAG terminal years need care later; the stock is
  retained with diagnostics flagged for assessment reconstruction.

These contrasts support keeping the full set for a cross-stock synthesis of
advice adherence.

# SAG vs FLStock

Scaled series using latest SAG reference points: SSB / \(B_{\mathrm{trigger}}\),
\(\bar{F}/F_{\mathrm{MSY}}\), catch / MSY proxy
(\(B_{\mathrm{trigger}}(1-e^{-F_{\mathrm{MSY}}})\)), recruitment / mean SAG
recruitment.

```{r sag-fl-compare, fig.height=10, fig.width=10}
sag_ts = merge(sag$ts, latest_ass, by = c("sid", "assYear"))
sag_ts = merge(sag_ts, stocks[, c("sid", "name")], by = "sid")
sag_ts$source = "SAG"

fl_ts = ldply(names(oms), function(id) {
  stk = oms[[id]]
  if (is.null(stk)) return(NULL)
  df = model.frame(FLQuants(
    ssb   = ssb(stk),
    f     = fbar(stk),
    rec   = stock.n(stk)[1],
    catch = catch(stk)), drop = TRUE)
  df$sid = id
  df$source = "FLStock"
  df
})
fl_ts = merge(fl_ts, stocks[, c("sid", "name")], by = "sid")

scale_one = function(df, sag_now) {
  ldply(split(df, df$sid), function(d) {
    rp = sag_now[match(d$sid[1], sag_now$sid), ]
    msy = if (is.finite(rp$msybtrigger) && is.finite(rp$fmsy))
      rp$msybtrigger * (1 - exp(-rp$fmsy)) else NA_real_
    rbar = mean(d$rec[d$source == "SAG"], na.rm = TRUE)
    if (!is.finite(rbar) || rbar == 0)
      rbar = mean(d$rec, na.rm = TRUE)
    rbind(
      data.frame(sid = d$sid, name = d$name, year = d$year,
                 source = d$source, qname = "SSB",
                 data = d$ssb / rp$msybtrigger, stringsAsFactors = FALSE),
      data.frame(sid = d$sid, name = d$name, year = d$year,
                 source = d$source, qname = "F",
                 data = d$f / rp$fmsy, stringsAsFactors = FALSE),
      data.frame(sid = d$sid, name = d$name, year = d$year,
                 source = d$source, qname = "Catch",
                 data = d$catch / msy, stringsAsFactors = FALSE),
      data.frame(sid = d$sid, name = d$name, year = d$year,
                 source = d$source, qname = "Rec",
                 data = d$rec / rbar, stringsAsFactors = FALSE))
  })
}

both = rbind(
  sag_ts[, c("sid", "name", "year", "source", "ssb", "f", "rec", "catch")],
  fl_ts[,  c("sid", "name", "year", "source", "ssb", "f", "rec", "catch")])
cmp = scale_one(both, sag_now)
cmp = subset(cmp, is.finite(data))
cmp$name  = factor(cmp$name, levels = levels(stocks$name))
cmp$qname = factor(cmp$qname, levels = c("SSB", "F", "Catch", "Rec"))

ggplot(cmp, aes(year, data, colour = source)) +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey50") +
  geom_line(linewidth = 0.8) +
  scale_colour_manual(values = c(SAG = "#1B9E77", FLStock = "#D95F02")) +
  facet_grid(name ~ qname, scales = "free_y") +
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(legend.position = "bottom",
        strip.text.y = element_text(angle = 0))
```

**Figure `r Fig = Fig + 1; Fig`**. SAG vs `FLStock`, scaled by SAG reference points (line at 1).

**Table `r Tab = Tab + 1; Tab`**. Terminal overlapping year on the scaled
series (\(r\) = SAG/FL).

```{r sag-fl-table}
term = ddply(cmp, .(sid, name, qname), function(d) {
  yrs = sort(intersect(d$year[d$source == "SAG"],
                       d$year[d$source == "FLStock"]))
  if (!length(yrs)) return(NULL)
  y = max(yrs)
  sag_v = d$data[d$source == "SAG" & d$year == y][1]
  fl_v  = d$data[d$source == "FLStock" & d$year == y][1]
  data.frame(year = y, SAG = sag_v, FL = fl_v, r = sag_v / fl_v)
})

qord = intersect(c("SSB", "F", "Catch", "Rec"), unique(as.character(term$qname)))
term_wide = reshape(term[, c("name", "year", "qname", "SAG", "FL", "r")],
                    idvar = c("name", "year"), timevar = "qname",
                    direction = "wide")
cols = unlist(lapply(qord, function(q) paste(c("SAG", "FL", "r"), q, sep = ".")))
num = intersect(cols, names(term_wide))
term_wide[num] = lapply(term_wide[num], function(x) round(x, 3))
term_wide = term_wide[match(levels(stocks$name), term_wide$name),
                      c("name", "year", cols)]

knitr::kable(
  term_wide,
  col.names = c("Stock", "Year", rep(c("SAG", "FL", "r"), length(qord))),
  row.names = FALSE,
  align = c("l", "r", rep("r", length(cols)))) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed")) %>%
  add_header_above(c(" " = 2, setNames(rep(3, length(qord)), qord)))
```

# Implications for the backtest

| Topic | Week~1 conclusion |
|-------|-------------------|
| Case studies | All six proposal stocks **confirmed** |
| Advice rules | ICES MSY approach for gadoids; Coastal States / ICES MSY rule for mackerel |
| Backtest window | Align peels with each stock's advice year and SAG coverage |
| Priority contrasts | Depleted Celtic Sea cod vs mixed gadoid/mackerel status |
| Next tasks | Assessment reconstruction; advice-rule coding; perfect-management counterfactuals |

```{r save}
dir.create(file.path(root, "data/results"), recursive = TRUE, showWarnings = FALSE)

screening = list(
  stocks    = stocks,
  inventory = inventory,
  sag       = sag,
  sag_now   = sag_now,
  sag_cov   = sag_cov,
  advice    = advice,
  oms       = oms,
  cmp       = cmp,
  term      = term,
  date      = Sys.Date())
save(screening, file = file.path(root, "data/results/01_screening.RData"))
```

Results saved to `data/results/01_screening.RData`.
Downstream: `02.0_condition_om.Rmd` / `02.2_dynamics.Rmd`.
