---
title: "Long-term fwd vs FLBRP"
subtitle: "Operating Model gate — deterministic Fmsy projection should reach refpts(eq)['msy',]"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
suppressPackageStartupMessages(suppressWarnings(library(knitr)))
root <- "C:/active/blueMarine"
dir.create(file.path(root, "tex/figs"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "data/results"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "Rmd/cache", "02.1_lterm_eq"), recursive = TRUE, showWarnings = FALSE)
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(root, "Rmd/cache", "02.1_lterm_eq/"),
  fig.path  = file.path(root, "tex/figs", "02.1_lterm_eq-"),
  fig.width = 9,
  fig.height= 5,
  dev       = "png")
Fig=0
Tab=0
```

```{r libs, echo=FALSE}
suppressPackageStartupMessages({
  library(FLCore)
  library(FLBRP)
  library(icesdata)
  library(ggplotFL)
  library(ggplot2)
  library(FLasher)
  library(FLBacktest)
})
source(file.path(root, "R/plot_theme.R"))
source(file.path(root, "R/loadStocks.R"))
source(file.path(root, "R/om.R"))
theme_set(theme_bluemarine_html())
load_om(root)
nYears   = 50L
avgYears = 10L
tol      = 0.05
```

# Purpose

A 50-year deterministic `FLasher::fwd` of each conditioned `FLStock` at
FLBRP \(F_{\mathrm{MSY}}\) should finish at `refpts(eq)["msy", ]`
(SSB, yield, recruitment, \(F\)).

The projection uses `fwdWindow(stk, eq)` so `fwd` sees the same mean \(M\),
weights, maturity and selectivity as `FLBRP`. Yield is **landings** (FLBRP
`yield` is not catch-with-discards). A stock passes when the mean of the last
10 years is within `r 100*tol`% of the FLBRP value.

# One stock

Irish Sea cod: extend the stock, project at \(F_{\mathrm{MSY}}\), compare
the terminal year with `refpts`.

```{r one-stock}
sid = "cod.27.7a"
stk = oms[[sid]]
eq  = eqs[[sid]]
rp  = refpts(eq)
msy = c(harvest = c(rp["msy", "harvest"]),
        ssb     = c(rp["msy", "ssb"]),
        yield   = c(rp["msy", "yield"]),
        rec     = c(rp["msy", "rec"]))

endYr = dims(stk)$maxyear + nYears
yrs   = (dims(stk)$maxyear + 1):endYr
stkW  = fwdWindow(stk, eq, end = endYr)
fmsy  = FLQuant(msy[["harvest"]], dimnames = list(year = ac(yrs)))
omFmsy = FLasher::fwd(stkW, fbar = fmsy, sr = eq)

term = window(omFmsy, start = endYr - avgYears + 1)
data.frame(
  metric = c("F", "SSB", "yield", "rec"),
  FLBRP  = unname(msy[c("harvest", "ssb", "yield", "rec")]),
  fwd    = c(mean(c(fbar(term))), mean(c(ssb(term))),
             mean(c(landings(term))), mean(c(rec(term)))),
  row.names = NULL)
```

```{r one-stock-plot, fig.height=7, fig.width=9, echo=FALSE}
x0 = dims(stk)$maxyear
plot(FLStocks("fwd" = omFmsy),
     metrics = list(Rec = rec, SSB = ssb, Catch = landings, F = fbar)) +
  geom_flpar(data = FLPars(
    Rec   = FLPar("Rmsy" = msy[["rec"]]),
    SSB   = FLPar("Bmsy" = msy[["ssb"]]),
    Catch = FLPar("MSY"  = msy[["yield"]]),
    F     = FLPar("Fmsy" = msy[["harvest"]])),
    x = rep(x0, 4)) +
  theme_bluemarine_html() +
  labs(x = "Year", title = stock_name(sid, stocks))
```

**Figure `r Fig = Fig + 1; Fig`**. Irish Sea cod: 50-year `fwd` at FLBRP
\(F_{\mathrm{MSY}}\). Horizontal lines are `refpts(eq)["msy", ]`.

# All stocks

```{r all-stocks}
omStks = if (is(oms, "FLStocks")) oms else FLStocks(oms)
eqBrps = if (is(eqs, "FLBRPs")) eqs else FLBRPs(eqs)
lt = ltermEq(omStks, eqBrps, nyears = nYears, avgYears = avgYears,
             tol = tol, biology = "window",
             labels = setNames(as.character(stocks$name),
                               as.character(stocks$sid)))
compare = lt$compare
compare$name = factor(compare$name, levels = as.character(stocks$name))
passTab = ltermPass(compare)
fwds = FLStocks(lapply(lt$byStock, function(x) x$fwds[["window"]]))
lterm_eq = list(compare = compare, pass = passTab, fwds = fwds,
                nYears = nYears, avgYears = avgYears, tol = tol)
save(lterm_eq, file = file.path(root, "data/results/03.3_lterm_eq.RData"))
```

**Table `r Tab = Tab + 1; Tab`**. Mean of the last `r avgYears` years vs
`refpts(eq)["msy", ]`.

```{r compare-table, echo=FALSE}
cmp = compare
cmp$flbrp = signif(cmp$flbrp, 4)
cmp$longTerm = signif(cmp$longTerm, 4)
cmp$ratio = round(cmp$ratio, 3)
knitr::kable(
  cmp[, c("name", "metric", "flbrp", "longTerm", "ratio", "pass")],
  col.names = c("Stock", "Metric", "FLBRP MSY", "Long-term mean",
                "fwd / FLBRP", "Pass"),
  row.names = FALSE)
```

**Table `r Tab = Tab + 1; Tab`**. Pass if \(F\), SSB, yield and recruitment
are all within `r 100*tol`%.

```{r pass-table, echo=FALSE}
knitr::kable(passTab, row.names = FALSE)
```

```{r ratio-plot, fig.height=6, fig.width=9, echo=FALSE}
rd = subset(compare, is.finite(ratio))
ggplot(rd, aes(name, ratio)) +
  geom_hline(yintercept = 1, linetype = 2, colour = bm_col("ref_strong")) +
  geom_hline(yintercept = c(1 - tol, 1 + tol), linetype = 3,
             colour = bm_col("muted")) +
  geom_point(size = 2.2) +
  facet_wrap(~ metric, scales = "free_y", ncol = 2) +
  labs(x = NULL, y = "fwd / FLBRP MSY") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

**Figure `r Fig = Fig + 1; Fig`**. Long-term `fwd` / FLBRP MSY. Dotted lines
are \(\pm\)`r 100*tol`%.

```{r all-plots, fig.height=7, fig.width=9, echo=FALSE}
for (id in names(fwds)) {
  eq = eqs[[id]]
  rp = refpts(eq)
  print(
    plot(FLStocks("fwd" = fwds[[id]]),
         metrics = list(Rec = rec, SSB = ssb, Catch = landings, F = fbar)) +
      geom_flpar(data = FLPars(
        Rec   = FLPar("Rmsy" = c(rp["msy", "rec"])),
        SSB   = FLPar("Bmsy" = c(rp["msy", "ssb"])),
        Catch = FLPar("MSY"  = c(rp["msy", "yield"])),
        F     = FLPar("Fmsy" = c(rp["msy", "harvest"]))),
        x = rep(dims(oms[[id]])$maxyear, 4)) +
      theme_bluemarine_html() +
      labs(x = "Year", title = stock_name(id, stocks)))
}
```

**Figure `r Fig = Fig + 1; Fig`**. Each Operating Model: `fwd` at FLBRP \(F_{\mathrm{MSY}}\)
with MSY reference lines. The last years should sit on the lines.

Output: `data/results/03.3_lterm_eq.RData`.
