---
title: "Stock and Operating Model dynamics"
subtitle: "Task 3 — dynamics, SPR0, process error, fwd vs FLBRP, realised status"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
mathjax: true
---

# Purpose

Summarise **historical dynamics** and Operating Model diagnostics for the six case-study
stocks, and check that long-term `FLasher::fwd` at \(F_{\mathrm{MSY}}\)
matches `FLBRP` before the Operating Models are treated as final. Operating Models
themselves are built in
[`02.0_condition_om.Rmd`](02.0_condition_om.Rmd) (`oms`, `eqs`, `eqls`).
The detailed 50-year trajectories are in
[`02.1_lterm_eq.Rmd`](02.1_lterm_eq.Rmd).

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
suppressPackageStartupMessages(suppressWarnings(library(knitr)))
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- bm_root()
dir.create(file.path(root, "tex/figs"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "data/results"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "data/interim/03.0_om"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "Rmd/cache", "02.2_dynamics"), recursive = TRUE, showWarnings = FALSE)
knitr::opts_chunk$set(
  echo      = FALSE, 
  warning   = FALSE, 
  message   = FALSE, 
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(root, "Rmd/cache", "02.2_dynamics/"),
  fig.path  = file.path(root, "tex/figs", "02.2_dynamics-"),
  fig.width = 9, 
  fig.height= 5, 
  dev       ="png")
Fig=0
Tab=0
```

```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(ggplotFL)
  library(ggplot2)
  library(FLBacktest)
  library(reshape2)
  library(data.table)
  library(jsonlite)
  library(FLasher)
})
source(file.path(root, "R/plot_theme.R"))
source(file.path(root, "R/loadStocks.R"))
source(file.path(root, "R/om.R"))
theme_set(theme_bluemarine_html())
```

```{r load-om}
load_om(root)
eqls_file = file.path(root, "data/interim/03.0_om/eqls.RData")
if (!file.exists(eqls_file))
  stop("Missing ", eqls_file, " — knit Rmd/02.0_condition_om.Rmd first.", call.=FALSE)
load(eqls_file)
tol = 0.05

# ldply over stocks: stop with a clear message instead of dropping failures.
map_stock = function(obj, fn, what = "stock") {
  out = tryCatch(
    ldply(obj, function(z) {
      r = tryCatch(fn(z), error = function(e) {
        stop(what, " failed: ", conditionMessage(e), call. = FALSE)
      })
      if (is.null(r)) return(NULL)
      r = as.data.frame(r)
      if (!NROW(r)) return(NULL)
      r
    }),
    error = function(e) stop(conditionMessage(e), call. = FALSE)
  )
  if (is.null(out) || !NROW(out))
    stop("No rows from ", what, " — check inputs.", call. = FALSE)
  out
}
scen_rows = function(srr, scen, dat) {
  if (is.null(dat) || !NROW(dat))
    stop("Empty rows for SRR=", srr, " Scenario=", scen, call. = FALSE)
  cbind(SRR = srr, Scenario = scen, dat)
}
```

# Biological variation

The historical backtest keeps these year-specific series rather than replacing them with means. Weight-at-age, \(M\) and maturity, together with recruitment residuals, are the observed productivity history.

**Table `r Tab=Tab+1; Tab`**. Whether natural mortality, maturity and mass
vary by age or year in each `FLStock`.

```{r vary}
vary = ldply(oms, checkVariation)
if (!("sid" %in% names(vary)) && (".id" %in% names(vary)))
  vary$sid = as.character(vary$.id)
vary = transform(vary,
  m.by.age   = ifelse(mAge,   "varies by age",  "constant by age"),
  m.by.yr    = ifelse(mYr,    "varies by year", "constant by year"),
  mat.by.yr  = ifelse(matYr,  "varies by year", "constant by year"),
  mass.by.yr = ifelse(massYr, "varies by year", "constant by year"))
vary = merge(vary, stocks[, c("sid","name")], by="sid")
vary$name = factor(vary$name, levels=levels(stocks$name))
vary = vary[order(vary$name), ]
knitr::kable(
  vary[, c("name","sid","m.by.age","m.by.yr","mat.by.yr","mass.by.yr")],
  col.names = c("Stock","SID","M (age)","M (year)","Maturity","Mass"),
  row.names = FALSE)
save(vary, file=file.path(root, "data/interim/03.0_om", "vary.RData"))
```

# \(F_{\mathrm{MSY}}\): `fwd` vs `FLBRP`

Gate from [`02.1_lterm_eq.Rmd`](02.1_lterm_eq.Rmd): 50-year deterministic
`fwd` at FLBRP \(F_{\mathrm{MSY}}\) (`fwdWindow(stk, eq)`) vs
`refpts(eq)["msy", ]`. Yield is landings.

```{r lterm-eq, cache=FALSE}
lterm_file = file.path(root, "data/results/03.3_lterm_eq.RData")
if (!file.exists(lterm_file))
  stop("Missing ", lterm_file, " — knit Rmd/02.1_lterm_eq.Rmd first.",
       call. = FALSE)
load(lterm_file)
lterm = lterm_eq
lterm_pass = lterm$pass
save(lterm, lterm_pass,
     file=file.path(root, "data/interim/03.0_om", "lterm.RData"))
```

**Table `r Tab=Tab+1; Tab`**. Long-term `fwd` / FLBRP MSY.

```{r lterm-eq-table}
cmp = lterm$compare
cmp$flbrp = signif(cmp$flbrp, 4)
cmp$longTerm = signif(cmp$longTerm, 4)
cmp$ratio = round(cmp$ratio, 3)
keep = intersect(c("name", "metric", "flbrp", "longTerm", "ratio", "pass"),
                 names(cmp))
knitr::kable(
  cmp[, keep],
  col.names = c("Stock", "Metric", "FLBRP MSY",
                "Long-term mean", "Ratio", "Pass")[seq_along(keep)],
  row.names = FALSE)
```

**Table `r Tab=Tab+1; Tab`**. Pass/fail (all four metrics within
`r 100*tol`%).

```{r lterm-eq-pass}
knitr::kable(lterm_pass, row.names = FALSE)
```

```{r lterm-eq-plot, fig.height=6, fig.width=9}
rd = subset(lterm$compare, is.finite(ratio))
if (nrow(rd)) {
  ggplot(rd, aes(name, ratio)) +
    geom_hline(yintercept = 1, linetype = 2, colour = bm_col("ref_strong")) +
    geom_hline(yintercept = c(1 - tol, 1 + tol), linetype = 3,
               colour = bm_col("muted")) +
    geom_point(size = 2.2) +
    facet_wrap(~ metric, scales = "free_y", ncol = 2) +
    labs(x = NULL, y = "fwd / FLBRP MSY") +
    theme(axis.text.x = element_text(angle = 30, hjust = 1))
}
```

**Figure `r Fig=Fig+1; Fig`**. Long-term `fwd` / FLBRP MSY. Dotted lines are
\(\pm\)`r 100*tol`%. Full trajectories: [`02.1_lterm_eq.Rmd`](02.1_lterm_eq.Rmd).

# Dynamic \(B_0\)

Unfished SSB by year from observed recruitment and year-specific SPR0
(`icesdata::spr0Yr`). No stock–recruitment fit.

**Figure `r Fig=Fig+1; Fig`**. Historical SSB and dynamic \(B_0\)
(\(R_t \times \mathrm{SPR}_{0,t}\)).

```{r dynB0, fig.height=8, fig.width=10}
dynB0 = FLQuants()
for (id in sids)
  dynB0[[id]] = rec(oms[[id]]) * spr0Yr(oms[[id]])

b0_plot = rbind(
  transform(as.data.frame(FLQuants(lapply(oms[sids], ssb)), drop=TRUE), series="SSB"),
  transform(as.data.frame(dynB0, drop=TRUE), series="dynB0"))
names(b0_plot)[names(b0_plot)=="qname"] = "sid"
b0_plot = merge(b0_plot, stocks, by="sid")
b0_plot$name = factor(b0_plot$name, levels=levels(stocks$name))
b0_plot = subset(b0_plot, is.finite(data))

ggplot(b0_plot, aes(year, data, colour=series)) +
  geom_line(linewidth=0.75) +
  scale_colour_manual(values=c(SSB=bm_col("advice"), dynB0=bm_col("catch"))) +
  facets_stock_wrap(ncol=2) +
  labs(x=NULL, y="Biomass", colour=NULL)
```

# Process error (production Operating Model)

**Figure `r Fig=Fig+1; Fig`**. Process error on the production Operating Model,
\((SSB_{t+1}-SSB_t+C_t-SP_t)/SSB_t\), from `processError`.

```{r pe, fig.height=8, fig.width=10}
pe = ldply(sids, function(id) {
  rtn = tryCatch(processError(eqs[[id]]), error = function(e)
    stop("processError(", id, "): ", conditionMessage(e), call. = FALSE))
  q = if (is(rtn, "FLQuants") && "pe" %in% names(rtn)) rtn[["pe"]] else rtn
  d = as.data.frame(q, drop = TRUE)
  if (!("year" %in% names(d) && "data" %in% names(d)))
    stop("processError(", id, ") did not return year/data columns.", call. = FALSE)
  cbind(sid = id, d)
}, .id = NULL)
ggplot_stock_ts(pe, y="data", ylab=expression((SSB[t+1]-SSB[t]+C[t]-SP[t])/SSB[t]),
                hline=0)
save(pe, file=file.path(root, "data/interim/03.0_om", "pe.RData"))
```

# SPR\(_0\)

**Figure `r Fig=Fig+1; Fig`**. Year-specific spawners per recruit in the
absence of fishing, \(\mathrm{SPR}_{0,t}\).

```{r spr0, fig.height=8, fig.width=10}
spr0Par = llply(oms, function(x) {
  res = spr0Yr(x)
  FLPar(array(c(res), dim=c(1, dim(res)[2]),
              dimnames=list(params="spr0", iter=seq(dim(res)[2]))))
})
spr0 = ldply(oms, function(x) {
  d = as.data.frame(spr0Yr(x), drop=TRUE)
  data.frame(year=d$year, spr0=d$data)
}, .id="sid")
ggplot_stock_ts(spr0, y="spr0", ylab=expression(SPR[0]))
save(spr0, spr0Par, file=file.path(root, "data/interim/03.0_om", "spr0.RData"))
```

# Year-varying SRR and virgin biomass

Beverton–Holt with a FishLife steepness prior and year-specific SPR0
(`icesdata::ftmb`).

```{r sr}
# icesdata:::ftmb is not exported and aborts this R build (NA/NaN gradient,
# then process exit 5). Skip year-varying SPR0 SRR; production
# nonStationarity below uses eqs from 02.0.
sr = NULL
srPar = NULL
sr_sids = character()
save(sr, srPar, file=file.path(root, "data/interim/03.0_om", "sr.RData"))
```

**Figure `r Fig=Fig+1; Fig`**. Year-varying virgin biomass implied by the
dynamic-SPR0 Beverton–Holt fit.

```{r virgin, fig.height=8, fig.width=10}
if (!length(sr_sids)) {
  virgin = data.frame(sid=character(), year=numeric(), v=numeric())
  cat("Skipping virgin biomass: no year-varying ftmb fits.\n")
} else {
  virgin = mdply(sr_sids, function(id) {
    cbind(sid=id, model.frame(srPar[[id]], drop=TRUE),
          year=unlist(dimnames(spr0Yr(oms[[id]]))$year))
  })[, -c(1, 6)]
  virgin$year = as.numeric(virgin$year)
  print(ggplot_stock_ts(virgin, y="v", ylab="Biomass"))
}
save(virgin, file=file.path(root, "data/interim/03.0_om", "virgin.RData"))
```

# Time-varying MSY (production SRR)

**Figure `r Fig=Fig+1; Fig`**. Year-specific \(B_{\mathrm{MSY}}\) from
`nonStationarity` on the production SRR.

```{r rf-ssb, fig.height=8, fig.width=10}
rf = mdply(sids, function(id) {
  out = tryCatch(
    cbind(sid = id, nonStationarity(oms[[id]], sr = eqs[[id]])),
    error = function(e)
      stop("nonStationarity(", id, "): ", conditionMessage(e), call. = FALSE))
  out
})
rf = reshape2::dcast(rf, sid+year+refpt ~ quant, value.var="data")
save(rf, file=file.path(root, "data/interim/03.0_om", "rf.RData"))
ggplot_stock_ts(subset(rf, refpt=="msy"), y="ssb", ylab="SSB")
```

**Figure `r Fig=Fig+1; Fig`**. Year-specific MSY yield on the production SRR.

```{r rf-yield, fig.height=8, fig.width=10}
ggplot_stock_ts(subset(rf, refpt=="msy"), y="yield", ylab="Yield")
```

**Figure `r Fig=Fig+1; Fig`**. Year-specific \(F_{\mathrm{MSY}}\) on the
production SRR.

```{r rf-f, fig.height=8, fig.width=10}
ggplot_stock_ts(subset(rf, refpt=="msy"), y="harvest", ylab="F")
```

# ABI (production Operating Model)

**Figure `r Fig=Fig+1; Fig`**. Age-based indicator \(\mathrm{ABI}_{\mathrm{MSY}}\)
on the production Operating Model (`eqs`), scaled to each stock's mean.

```{r abi, fig.height=8, fig.width=10}
abi = ldply(sids, function(id) {
  rtn = tryCatch(icesdata::abi(oms[[id]], eqs[[id]]), error = function(e)
    stop("abi(", id, "): ", conditionMessage(e), call. = FALSE))
  cbind(sid = id, as.data.frame(rtn, drop = TRUE))
}, .id = NULL)
names(abi)[names(abi) == "data"] = "abi"
save(abi, file=file.path(root, "data/interim/03.0_om", "abi.RData"))
abi2 = ddply(abi, .(sid), transform, abi=abi/mean(abi))
ggplot_stock_ts(abi2, y="abi", ylab="ABI / mean ABI")
```

# Equilibrium yield and recruitment (scenario Operating Models)

**Figure `r Fig=Fig+1; Fig`**. Equilibrium yield against SSB for the six SRR
branches (colour = SRR, linetype = prior). Grey path and points are the
observed assessment trajectory.

```{r pf-yield, fig.height=11, fig.width=10}
pf = rbind(
  cbind(id=1, ldply(bh1, function(x) model.frame(FLQuants(x, SSB=ssb, Yield=catch, Recruits=rec), drop=TRUE))),
  cbind(id=2, ldply(bh2, function(x) model.frame(FLQuants(x, SSB=ssb, Yield=catch, Recruits=rec), drop=TRUE))),
  cbind(id=3, ldply(bh3, function(x) model.frame(FLQuants(x, SSB=ssb, Yield=catch, Recruits=rec), drop=TRUE))),
  cbind(id=4, ldply(rk1, function(x) model.frame(FLQuants(x, SSB=ssb, Yield=catch, Recruits=rec), drop=TRUE))),
  cbind(id=5, ldply(rk2, function(x) model.frame(FLQuants(x, SSB=ssb, Yield=catch, Recruits=rec), drop=TRUE))),
  cbind(id=6, ldply(rk3, function(x) model.frame(FLQuants(x, SSB=ssb, Yield=catch, Recruits=rec), drop=TRUE))))
names(pf)[2] = "sid"
pf = merge(pf, stocks[, c("sid","name")])
scen = expand.grid(Scenario=c("h & R0 Estimated","Steepness Prior","R0 Fixed"),
                   SRR=c("Beverton & Holt","Ricker"))
pf = cbind(scen[pf$id, ], pf)
obs = ldply(bh1, function(x)
  model.frame(FLQuants(x, SSB=ssb.obs, Yield=catch.obs, Recruits=rec.obs), drop=TRUE),
  .id="sid")
obs = merge(obs, stocks[, c("sid","name")])
save(pf, obs, file=file.path(root, "data/interim/03.0_om", "pf.RData"))

ggplot(pf, aes(SSB, Yield, colour=SRR, linetype=Scenario)) +
  facet_wrap(~ name, scales="free", ncol=2) +
  geom_path(data=obs, aes(SSB, Yield), inherit.aes=FALSE,
            colour=bm_col("muted"), linewidth=0.35, alpha=0.75, na.rm=TRUE) +
  geom_point(data=obs, aes(SSB, Yield), inherit.aes=FALSE,
             colour=bm_col("ink"), size=1.1, alpha=0.45, stroke=0, na.rm=TRUE) +
  geom_line(linewidth=0.75, na.rm=TRUE) +
  scale_colour_manual(values=c("Beverton & Holt"=bm_col("advice"),
                               "Ricker"=bm_col("catch"))) +
  scale_linetype_manual(values=c("h & R0 Estimated"="solid",
                                 "Steepness Prior"="dashed",
                                 "R0 Fixed"="longdash")) +
  labs(x="SSB (t)", y="Yield (t)", colour=NULL, linetype=NULL) +
  theme_bluemarine_html()
```

**Figure `r Fig=Fig+1; Fig`**. Equilibrium recruitment against SSB for the
same six SRR scenarios, with the observed stock–recruit pairs in grey.

```{r pf-rec, fig.height=11, fig.width=10}
ggplot(pf, aes(SSB, Recruits, colour=SRR, linetype=Scenario)) +
  facet_wrap(~ name, scales="free", ncol=2) +
  geom_path(data=obs, aes(SSB, Recruits), inherit.aes=FALSE,
            colour=bm_col("muted"), linewidth=0.35, alpha=0.75, na.rm=TRUE) +
  geom_point(data=obs, aes(SSB, Recruits), inherit.aes=FALSE,
             colour=bm_col("ink"), size=1.1, alpha=0.45, stroke=0, na.rm=TRUE) +
  geom_line(linewidth=0.75, na.rm=TRUE) +
  scale_colour_manual(values=c("Beverton & Holt"=bm_col("advice"),
                               "Ricker"=bm_col("catch"))) +
  scale_linetype_manual(values=c("h & R0 Estimated"="solid",
                                 "Steepness Prior"="dashed",
                                 "R0 Fixed"="longdash")) +
  labs(x="SSB (t)", y="Recruits", colour=NULL, linetype=NULL) +
  theme_bluemarine_html()
```

# Scenario diagnostics

Life-history covariates, time series, process error and ABI on the six SRR
branches (not the production Operating Model).

```{r covars}
covars = rbind.fill(
  scen_rows("Beverton & Holt", 1, map_stock(bh1, function(x) covarFn(x))),
  scen_rows("Beverton & Holt", 2, map_stock(bh2, function(x) covarFn(x))),
  scen_rows("Beverton & Holt", 3, map_stock(bh3, function(x) covarFn(x))),
  scen_rows("Ricker", 1, map_stock(rk1, function(x) covarFn(x, model="ricker"))),
  scen_rows("Ricker", 2, map_stock(rk2, function(x) covarFn(x, model="ricker"))),
  scen_rows("Ricker", 3, map_stock(rk3, function(x) covarFn(x, model="ricker"))))
if (is.null(covars) || !NROW(covars))
  covars = data.frame(SRR=character(), Scenario=integer(), stringsAsFactors=FALSE)
if (NROW(covars) && "Scenario" %in% names(covars))
  covars$Scenario = factor(covars$Scenario, levels=1:3,
    labels=c("h & R0 Estimated","Steepness Prior","R0 Fixed"))
if (".id" %in% names(covars)) names(covars)[names(covars) == ".id"] = "sid"
save(covars, file=file.path(root, "data/interim/03.0_om", "covars.RData"))
```

```{r ts-scen}
tsFn = function(x) {
  model.frame(FLQuants(f=fbar(x), hrate=catch(x)%/%ebiomass(x),
                       eb=ebiomass(x), ssb=ssb(x), rec=rec(x), catch=catch(x)),
              drop=TRUE)
}
ts = rbind.fill(
  scen_rows("Beverton & Holt", 1, map_stock(bh1, tsFn)),
  scen_rows("Beverton & Holt", 2, map_stock(bh2, tsFn)),
  scen_rows("Beverton & Holt", 3, map_stock(bh3, tsFn)),
  scen_rows("Ricker", 1, map_stock(rk1, tsFn)),
  scen_rows("Ricker", 2, map_stock(rk2, tsFn)),
  scen_rows("Ricker", 3, map_stock(rk3, tsFn)))
if (is.null(ts) || !NROW(ts))
  ts = data.frame(SRR=character(), Scenario=integer(), stringsAsFactors=FALSE)
if (NROW(ts) && "Scenario" %in% names(ts))
  ts$Scenario = factor(ts$Scenario, levels=1:3,
    labels=c("h & R0 Estimated","Steepness Prior","R0 Fixed"))
save(ts, file=file.path(root, "data/interim/03.0_om", "ts.RData"))
```

```{r pe-scen}
peScen = rbind.fill(
  scen_rows("Beverton & Holt", 1, map_stock(bh1, function(x)
    model.frame(processError(x), drop=TRUE))),
  scen_rows("Beverton & Holt", 2, map_stock(bh2, function(x)
    model.frame(processError(x), drop=TRUE))),
  scen_rows("Beverton & Holt", 3, map_stock(bh3, function(x)
    model.frame(processError(x), drop=TRUE))),
  scen_rows("Ricker", 1, map_stock(rk1, function(x)
    model.frame(processError(x), drop=TRUE))),
  scen_rows("Ricker", 2, map_stock(rk2, function(x)
    model.frame(processError(x), drop=TRUE))),
  scen_rows("Ricker", 3, map_stock(rk3, function(x)
    model.frame(processError(x), drop=TRUE))))
if (is.null(peScen) || !NROW(peScen))
  peScen = data.frame(SRR=character(), Scenario=integer(), stringsAsFactors=FALSE)
if (NROW(peScen) && "Scenario" %in% names(peScen))
  peScen$Scenario = factor(peScen$Scenario, levels=1:3,
    labels=c("h & R0 Estimated","Steepness Prior","R0 Fixed"))
save(peScen, file=file.path(root, "data/interim/03.0_om", "peScen.RData"))
```

```{r abi-scen}
abiFn = function(id, eq) {
  rtn = as.data.frame(abi(oms[[id]], eq[[id]]), drop=TRUE)
  if ("try-error" %in% is(rtn)) return(NULL)
  rtn
}
abiScen = rbind(
  cbind(SRR="Beverton & Holt", rbind(
    cbind(Scenario=1, mdply(data.frame(id=sids), abiFn, bh1)),
    cbind(Scenario=2, mdply(data.frame(id=sids), abiFn, bh2)),
    cbind(Scenario=3, mdply(data.frame(id=sids), abiFn, bh3)))),
  cbind(SRR="Ricker", rbind(
    cbind(Scenario=1, mdply(data.frame(id=sids), abiFn, rk1)),
    cbind(Scenario=2, mdply(data.frame(id=sids), abiFn, rk2)),
    cbind(Scenario=3, mdply(data.frame(id=sids), abiFn, rk3)))))
abiScen$Scenario = factor(abiScen$Scenario, levels=1:3,
  labels=c("h & R0 Estimated","Steepness Prior","R0 Fixed"))
names(abiScen)[5] = "abi"
save(abiScen, file=file.path(root, "data/interim/03.0_om", "abiScen.RData"))
abiTerminal = ddply(abiScen, .(id, SRR, Scenario), with,
                    data.frame(abi=mean(abi[year %in% (max(year)-0:2)])))
save(abiTerminal, file=file.path(root, "data/interim/03.0_om", "abiTerminal.RData"))
```

# Time-varying MSY (scenario Operating Models)

```{r nonstat}
nsFn = function(id, eq, matSel=FALSE) {
  out = cbind(sid=id, nonStationarity(oms[[id]], sr=eq[[id]], matSel=matSel))
  if ("try-error" %in% is(out)) return(NULL)
  out
}
rf2 = rbind(
  cbind(SRR="Beverton & Holt", rbind(
    cbind(Scenario=1, mdply(sids, nsFn, bh1)),
    cbind(Scenario=2, mdply(sids, nsFn, bh2)),
    cbind(Scenario=3, mdply(sids, nsFn, bh3)))),
  cbind(SRR="Ricker", rbind(
    cbind(Scenario=1, mdply(sids, nsFn, rk1)),
    cbind(Scenario=2, mdply(sids, nsFn, rk2)),
    cbind(Scenario=3, mdply(sids, nsFn, rk3)))))
rf2$Scenario = factor(rf2$Scenario, levels=1:3,
  labels=c("h & R0 Estimated","Steepness Prior","R0 Fixed"))
save(rf2, file=file.path(root, "data/interim/03.0_om", "rf2.RData"))
msy = reshape2::dcast(subset(rf2, refpt=="msy"), sid+year+SRR+Scenario ~ quant, value.var="data")
names(msy)[5:9] = paste("msy", names(msy)[5:9], sep=".")
names(msy) = gsub("\\.", "_", names(msy))
save(msy, file=file.path(root, "data/interim/03.0_om", "msy.RData"))

rfMatSel = rbind.fill(
  cbind(SRR="Beverton & Holt", rbind.fill(
    cbind(Scenario=1, mdply(sids, nsFn, bh1, TRUE)),
    cbind(Scenario=2, mdply(sids, nsFn, bh2, TRUE)),
    cbind(Scenario=3, mdply(sids, nsFn, bh3, TRUE)))),
  cbind(SRR="Ricker", rbind.fill(
    cbind(Scenario=1, mdply(sids, nsFn, rk1, TRUE)),
    cbind(Scenario=2, mdply(sids, nsFn, rk2, TRUE)),
    cbind(Scenario=3, mdply(sids, nsFn, rk3, TRUE)))))
rfMatSel$Scenario = factor(rfMatSel$Scenario, levels=1:3,
  labels=c("h & R0 Estimated","Steepness Prior","R0 Fixed"))
save(rfMatSel, file=file.path(root, "data/interim/03.0_om", "rfMatSel.RData"))
msyMatSel = reshape2::dcast(subset(rfMatSel, refpt=="msy"),
                 sid+year+SRR+Scenario ~ quant, value.var="data")
names(msyMatSel)[5:9] = paste("msy", names(msyMatSel)[5:9], sep=".")
names(msyMatSel) = gsub("\\.", "_", names(msyMatSel))
save(msyMatSel, file=file.path(root, "data/interim/03.0_om", "msyMatSel.RData"))
```

**Figure `r Fig=Fig+1; Fig`**. Year-specific MSY yield by SRR and prior.

```{r plot-msy-scenarios, fig.height=8, fig.width=10}
msy_plot = merge(msy, msyMatSel, by=c("sid","year","SRR","Scenario"))
names(msy_plot) = gsub("msy_", "", names(msy_plot))
ggplot_stock_ts_cmp(msy_plot, y="yield.x", colour="Scenario", linetype="SRR",
                    ylab="Yield")
```

**Figure `r Fig=Fig+1; Fig`**. Year-specific MSY when selectivity is set equal
to maturity.

```{r plot-msy-matsel, fig.height=8, fig.width=10}
ggplot_stock_ts_cmp(msy_plot, y="yield.y", colour="Scenario", linetype="SRR",
                    ylab="Yield")
```

**Figure `r Fig=Fig+1; Fig`**. Production-like branch (Beverton–Holt, \(R_0\)
fixed): MSY versus MSY with selectivity equal to maturity.

```{r plot-msy-compare, fig.height=8, fig.width=10}
msy_sel = subset(msy_plot, SRR=="Beverton & Holt" & Scenario=="R0 Fixed")
msy_sel = reshape2::melt(msy_sel, id.vars=c("sid","year"),
                         measure.vars=c("yield.x","yield.y"),
                         variable.name="series", value.name="yield")
msy_sel$series = factor(msy_sel$series, levels=c("yield.x","yield.y"),
                        labels=c("MSY","MSY (Sel = Mat)"))
ggplot_stock_ts_cmp(msy_sel, y="yield", colour="series", ylab="Yield",
                    colour_values=c("MSY"=bm_col("catch"),
                                    "MSY (Sel = Mat)"=bm_col("advice")))
```

# Realised status from assessments

Summaries from SAG and the conditioned `FLStock`s **before** any backtest.

```{r advice-setup}
scr_file = file.path(root, "data/results/01_screening.RData")
if (file.exists(scr_file)) {
  load(scr_file)
  advice = screening$advice
} else {
  sagAdvice = function(sid) {
    sid = unique(as.character(sid))
    parts = lapply(sid, function(s) {
      url = paste0("https://asd.ices.dk/API/getAdviceViewRecord?StockCode=",
                   URLencode(s, reserved=TRUE))
      d = tryCatch(jsonlite::fromJSON(url), error=function(e) NULL)
      if (is.null(d) || !NROW(d)) return(NULL)
      d = as.data.frame(d, stringsAsFactors=FALSE)
      data.frame(sid=s,
                 assYear=as.integer(d$assessmentYear),
                 adviceYear=as.integer(substr(as.character(d$adviceApplicableFrom),1,4)),
                 advice=suppressWarnings(as.numeric(d$adviceValue)),
                 stringsAsFactors=FALSE)
    })
    out = do.call(rbind, Filter(Negate(is.null), parts))
    if (is.null(out))
      return(data.frame(sid=character(), assYear=integer(),
                        adviceYear=integer(), advice=double(),
                        stringsAsFactors=FALSE))
    out[order(out$sid, out$adviceYear, out$assYear), ]
  }
  advice = sagAdvice(sids)
}

ts_sag = as.data.table(sag$ts)
rp = as.data.table(sag$rfpts)
rp_current = rp[order(sid, -assYear), .SD[1], by=sid]
dt = merge(ts_sag, rp_current[, .(sid, Blim, Btrigger=MSYBtrigger,
                                  Fmsy=FMSY, Fpa)],
           by="sid", all.x=TRUE)
dt = merge(dt, stocks, by="sid", all.x=TRUE)
num_cols = intersect(c("rec","ssb","catch","landings","discards","fbar","f"), names(dt))
if (length(num_cols))
  dt[, (num_cols) := lapply(.SD, as.numeric), .SDcols=num_cols]
if (!"fbar" %in% names(dt) && "f" %in% names(dt))
  dt$fbar = dt$f
```

**Figure `r Fig=Fig+1; Fig`**. ICES advised catch (Advice and Scenarios
Database) and reported catch from the latest SAG assessment.

```{r asd-advice, fig.height=10, fig.width=10}
if (!"name" %in% names(advice))
  advice = merge(advice, stocks, by="sid")
advice = subset(advice, is.finite(adviceYear) & is.finite(advice))
latest_ass = aggregate(assYear ~ sid, data=sag$ts, FUN=max)
reported = merge(sag$ts, latest_ass, by=c("sid","assYear"))
reported = merge(reported, stocks, by="sid")
reported = subset(reported, is.finite(year) & is.finite(catch))
advice_plot = rbind(
  data.frame(sid=advice$sid, name=advice$name, year=advice$adviceYear,
             series="ICES advice", data=advice$advice, stringsAsFactors=FALSE),
  data.frame(sid=reported$sid, name=reported$name, year=reported$year,
             series="Reported catch", data=reported$catch, stringsAsFactors=FALSE))
advice_plot$series = factor(advice_plot$series,
                            levels=c("Reported catch","ICES advice"))
advice_plot = subset(advice_plot, year >= 2010L & is.finite(data))
advice_plot$name = factor(advice_plot$name, levels=levels(stocks$name))
ggplot(advice_plot, aes(year, data, colour=series, linetype=series)) +
  geom_line(linewidth=0.85) +
  geom_point(size=1.2, alpha=0.9) +
  scale_catch_advice_colour() +
  scale_catch_advice_linetype() +
  facets_stock_row() +
  labs(x=NULL, y="Catch (t)", colour=NULL, linetype=NULL)
```

**Table `r Tab=Tab+1; Tab`**. Terminal-year SSB, \(F\) and catch (also the
mean of the last three years).

```{r advice-current}
Current = dt[order(sid, year),
  .(termYear = last(year),
    ssbT     = last(ssb),
    ssbL3    = mean(tail(ssb, 3), na.rm=TRUE),
    fbarT    = last(fbar),
    fbarL3   = mean(tail(fbar, 3), na.rm=TRUE),
    catchT   = last(catch),
    catchL3  = mean(tail(catch, 3), na.rm=TRUE),
    recT     = last(rec),
    recL3    = mean(tail(rec, 3), na.rm=TRUE),
    ssb_btrigT  = last(ssb) / first(Btrigger),
    ssb_btrigL3 = mean(tail(ssb, 3)) / first(Btrigger),
    ssb_blimT   = last(ssb) / first(Blim),
    f_fmsyT     = last(fbar) / first(Fmsy),
    f_fmsyL3    = mean(tail(fbar, 3)) / first(Fmsy)),
  by=.(sid, name)]
knitr::kable(Current[, c(1:3, seq(5, 9, 2))], digits=2,
             col.names=c("SID","Stock","Year","SSB","F","Catch"),
             row.names=FALSE)
```

**Table `r Tab=Tab+1; Tab`**. Terminal-year status relative to
\(MSYB_{\mathrm{trigger}}\), \(B_{\lim}\) and \(F_{\mathrm{MSY}}\).

```{r advice-current-rel}
knitr::kable(Current[, c(1:3, 12, 14, 15)], digits=2,
             col.names=c("SID","Stock","Year","SSB/Btrig","SSB/Blim","F/FMSY"),
             row.names=FALSE)
```

**Table `r Tab=Tab+1; Tab`**. Historical since 2010: share of years below
\(B_{\lim}\) and \(B_{\mathrm{trigger}}\), share with \(F>F_{\mathrm{MSY}}\),
mean catch and AAV.

```{r advice-path}
histICES = subset(dt, year >= 2010)[order(sid, year),
  .(n_years       = .N,
    year_from     = first(year),
    year_to       = last(year),
    p_below_blim  = mean(ssb < Blim, na.rm=TRUE),
    p_below_btrig = mean(ssb < Btrigger, na.rm=TRUE),
    p_f_over_fmsy = mean(fbar > Fmsy, na.rm=TRUE),
    mean_catch    = mean(catch, na.rm=TRUE),
    med_catch     = median(catch, na.rm=TRUE),
    aav           = mean(abs(diff(catch)) / head(catch, -1), na.rm=TRUE),
    max_drawdown  = min(catch, na.rm=TRUE) / max(catch, na.rm=TRUE)),
  by=.(sid, name)]
knitr::kable(histICES, digits=3, row.names=FALSE)
```

**Table `r Tab=Tab+1; Tab`**. Realised summary (Current joined to Historical) for
comparison with the HCR counterfactual in `05.1_metrics.Rmd`.

```{r advice-realised}
realised = merge(Current, histICES, by=c("sid","name"))
knitr::kable(realised, digits=3, row.names=FALSE)
```

**Figure `r Fig=Fig+1; Fig`**. SAG SSB with \(B_{\lim}\) (dashed) and
\(MSYB_{\mathrm{trigger}}\) (dotted).

```{r advice-fig-ssb, fig.width=10, fig.height=6}
ggplot(dt, aes(year, ssb)) +
  geom_line(colour=bm_col("advice"), linewidth=0.7) +
  geom_hline(aes(yintercept=Blim), linetype="dashed", colour=bm_col("catch")) +
  geom_hline(aes(yintercept=Btrigger), linetype="dotted", colour=bm_col("trigger")) +
  facet_wrap(~ name, scales="free_y") +
  labs(x=NULL, y="SSB (t)")
```

**Figure `r Fig=Fig+1; Fig`**. Mean catch in the last ten years versus
production-Operating Model MSY (dashed).

```{r advice-fig-catch, fig.width=10, fig.height=6}
msy_rf = ldply(eqs, function(x)
  model.frame(FLBRP::refpts(x)["msy","yield"], drop=TRUE), .id="sid")
ggplot(merge(ddply(dt, .(sid), with,
                   data.frame(cumCatch=cumsum(tail(catch, 10))/seq(10),
                              year=tail(year, 10))),
             stocks, by="sid"),
       aes(year, cumCatch)) +
  geom_line(colour=bm_col("advice"), linewidth=0.7) +
  geom_hline(data=merge(msy_rf, stocks, by="sid"),
             aes(yintercept=msy), colour=bm_col("catch"), linetype=2) +
  facet_wrap(~ name, scales="free_y", ncol=2) +
  labs(x=NULL, y="Mean catch (t)") +
  scale_y_continuous(limits=c(0, NA))
```

**Figure `r Fig=Fig+1; Fig`**. Kobe-style path of \(SSB/MSYB_{\mathrm{trigger}}\)
against \(F/F_{\mathrm{MSY}}\); the terminal year is the large point.

```{r advice-fig-kobe, fig.width=10, fig.height=6}
ggplot(dt, aes(ssb/Btrigger, fbar/Fmsy, colour=year)) +
  geom_path(alpha=0.7) +
  geom_point(data=dt[, .SD[.N], by=sid], size=2) +
  geom_hline(yintercept=1, linetype="dashed") +
  geom_vline(xintercept=1, linetype="dashed") +
  scale_colour_viridis_c() +
  facet_wrap(~ name, scales="free") +
  labs(x="SSB / MSY Btrigger", y="F / FMSY")
```

```{r save}
to_save = c("sag","oms","eqs","rsdl","rods","devs","stocks","sids",
            "dynB0","Current","histICES","realised","dt","rp_current",
            "lterm","lterm_pass")
to_save = intersect(to_save, ls())
save(list=to_save, file=file.path(root, "data/results/03.0_om.RData"))
```

`data/results/03.0_om.RData` is updated with `dynB0`, the realised-status
tables, and the `fwd` vs `FLBRP` check (`lterm`). Scenario diagnostics remain
under `data/interim/03.0_om/`.
