---
title: "Open-loop benchmark projections"
subtitle: "Methods — historical fwd at ICES FMSY, six SRRs, three uncertainty treatments"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

**Methods — Benchmark projections** (`tex/project-markup.tex`): historical
open-loop `fwd` at ICES \(F_{\mathrm{MSY}}\) from 1990 (or two years after the first assessment year if later) to the terminal year,
compared with the realised assessment.

Same Operating Models as [`04.2_closedLoop.Rmd`](04.2_closedLoop.Rmd),
built in [`02.0_condition_om.Rmd`](02.0_condition_om.Rmd) (`oms`, `eqls`).
The only difference is the harvest policy: constant ICES \(F_{\mathrm{MSY}}\)
from 1990 here, ICES HCR from 2015 there. Both start from the historical
assessment (`oms`).

| Uncertainty | Operating Model | `fwd` residuals |
|-------------|:---------------:|-----------------|
| `rec` | `oms` | historical rec / geometric mean |
| `srr` | `oms` | `srResiduals(eq)` |
| `M` | as `srr`, plus autocorrelated \(M\) on 100 iterations | `srResiduals(eq)` |

ICES \(F_{\mathrm{MSY}}\) is held fixed; each SRR supplies its own biology /
stock–recruitment. Historical projections keep the assessment's year-specific
\(M\), mass and maturity and replay recruitment residuals, so the productivity
that actually occurred is held common across policies.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
suppressPackageStartupMessages(suppressWarnings(library(knitr)))
root <- "C:/active/blueMarine"
dir.create(file.path(root, "tex/figs"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "data/results"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "Rmd/cache", "04.1_openLoop"), recursive = TRUE, showWarnings = FALSE)
knitr::opts_chunk$set(
  echo      = FALSE, 
  warning   = FALSE, 
  message   = FALSE, 
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(root, "Rmd/cache", "04.1_openLoop/"),
  fig.path  = file.path(root, "tex/figs", "04.1_openLoop-"),
  fig.width = 9, 
  fig.height= 5, 
  dev       ="png")
Fig=0
Tab=0
```


```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(ggplotFL)
  library(ggplot2)
  library(FLBacktest)
  library(FLasher)
  library(doParallel)
  library(parallel)
  library(foreach)
})
source(file.path(root, "R/plot_theme.R"))
source(file.path(root, "R/loadStocks.R"))
source(file.path(root, "R/om.R"))
theme_set(theme_bluemarine_html())
```

```{r load-om}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))
if (!exists("eqs")) {
  if (!"bh1" %in% names(eqls))
    stop("eqs missing and eqls[['bh1']] not found — knit 02.0 first.",
         call. = FALSE)
  eqs = eqls[["bh1"]]
}
sids = as.character(stocks$sid)
source(file.path(root, "R/om.R"))
require_om_gate(root, sids)
om   = oms
srDev = lapply(eqls, function(eq) FLQuants(lapply(eq, srResiduals)))
recDev = FLQuants(lapply(oms, function(x) {
  r = rec(x)
  r %/% exp(mean(log(c(r)), na.rm = TRUE))
}))
startYr = setNames(vapply(sids, function(id)
  openloop_start(om[[id]]), integer(1L)), sids)
endYr   = setNames(sapply(om, function(x) dims(x)$maxyear), names(om))
```

# Open-loop `fwd`

ICES \(F_{\mathrm{MSY}}\) on the historical assessment Operating Models (`oms`),
all six SRR scenarios. Each `fwd` is `FLStocks(foreach(id = sids))`.
Unconstrained BH (`eqs` / `bh1`) is the example below. Simulations start in 1990 from `oms` where the assessment allows, or in
the second year after `minyear` if the series begins after 1990.

```{r scenario-fwd}
flr_pkgs = c("methods", "FLCore", "FLBRP", "icesdata", "FLBacktest", "FLasher")
ncores = max(1L, detectCores() - 4L)
cl = makeCluster(ncores)
registerDoParallel(cl)
tryCatch({
  clusterExport(cl, "flr_pkgs")
  clusterEvalQ(cl, {
    suppressPackageStartupMessages(
      lapply(flr_pkgs, library, character.only = TRUE))
    NULL
  })
  clusterExport(cl, c("om", "eqls", "eqs", "recDev", "srDev",
                      "startYr", "endYr", "sids"), envir = environment())
  scen  = list()
  scen2 = list()
  for (nm in names(eqls)) {
    clusterExport(cl, "nm", envir = environment())
    scen[[nm]] = FLStocks(foreach(id = sids, .packages = flr_pkgs,
                                  .export = c("om", "eqls", "eqs", "recDev",
                                              "startYr", "endYr", "nm")) %dopar% {
      fices = c(refpts(eqs[[id]])["FMSY", "harvest"])
      ftrg  = window(fbar(om[[id]]), start = startYr[id], end = endYr[id]) %=% fices
      FLasher::fwd(om[[id]], fbar = ftrg, sr = eqls[[nm]][[id]], residuals = recDev[[id]])
    })
    names(scen[[nm]]) = sids

    scen2[[nm]] = FLStocks(foreach(id = sids, .packages = flr_pkgs,
                                   .export = c("om", "eqls", "eqs", "srDev",
                                               "startYr", "endYr", "nm")) %dopar% {
      fices = c(refpts(eqs[[id]])["FMSY", "harvest"])
      ftrg  = window(fbar(om[[id]]), start = startYr[id], end = endYr[id]) %=% fices
      FLasher::fwd(om[[id]], fbar = ftrg, sr = eqls[[nm]][[id]],
          residuals = srDev[[nm]][[id]])
    })
    names(scen2[[nm]]) = sids
  }
}, finally = stopCluster(cl))
ftar  = scen[["bh1"]]
ftar2 = scen2[["bh1"]]
openLoop = list(Recruits = scen, SRR = scen2)
save(openLoop, startYr, endYr, file = file.path(root, "data/results/03.1_openLoop.RData"))
```

