---
title: "Closed-loop counterfactual simulation"
subtitle: "Methods — historical hcrICES, six SRRs, three uncertainty treatments"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

**Methods — Counterfactual simulation** (`tex/project-markup.tex`): historical
closed-loop `hcrICES` from 2015 to the terminal year, starting from the
realised assessment (`oms`), compared with the realised assessment.

Same Operating Models as [`04.1_openLoop.Rmd`](04.1_openLoop.Rmd),
built in [`02.0_condition_om.Rmd`](02.0_condition_om.Rmd) (`oms`, `eqls`).
The only difference is the harvest policy: ICES HCR here, constant ICES
\(F_{\mathrm{MSY}}\) there. The Advice rule does **not** start from the
open-loop \(F_{\mathrm{target}}\) trajectory.

| Uncertainty | Operating Model | `sr_deviances` |
|-------------|:---------------:|----------------|
| `rec` | `oms` | historical rec / geometric mean |
| `srr` | `oms` | `srResiduals(eq)` |
| `M` | as `srr`, plus autocorrelated \(M\) on 100 iterations | `srResiduals(eq)` |

ICES \(F_{\mathrm{tar}}\) / \(B_{\mathrm{trigger}}\) / \(B_{\lim}\) are held
fixed (`hcrParams` from `eqs`). Each SRR supplies its own biology /
stock–recruitment. Historical projections keep the assessment's year-specific
\(M\), mass and maturity and replay recruitment residuals, so the productivity
that actually occurred is held common across policies.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- bm_root()
root <- bm_root()

dir.create(file.path(root, "tex/figs"),                     recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "data/results"),                 recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "Rmd/cache", "04.2_closedLoop"), recursive=TRUE, showWarnings=FALSE)

suppressPackageStartupMessages(suppressWarnings(library(knitr)))
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(root, "Rmd/cache", "04.2_closedLoop/"),
  fig.path  = file.path(root, "tex/figs", "04.2_closedLoop-"),
  fig.width = 9,
  fig.height= 5,
  dev       ="png")

Fig=0
Tab=0
```


```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(ggplotFL)
  library(ggplot2)
  library(FLasher)
  library(FLBacktest)
  library(doParallel)
  library(parallel)
  library(foreach)})

source(file.path(root, "R/plot_theme.R"))

theme_set(theme_bluemarine_html())
```

```{r load-om}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))

sids = as.character(stocks$sid)
om   = oms
if (!exists("eqs")) {
  if (!"bh1" %in% names(eqls))
    stop("eqs missing and eqls[['bh1']] not found — knit 02.0 first.",
         call. = FALSE)
  eqs = eqls[["bh1"]]
}
source(file.path(root, "R/om.R"))
require_om_gate(root, sids)

srDev = lapply(eqls, function(eq) FLQuants(lapply(eq, srResiduals)))
startYr = setNames(rep(2015, length(sids)), sids)
endYr   = setNames(sapply(om, function(x) dims(x)$maxyear), names(om))
```

# Closed-loop `hcrICES`

ICES HCR on the historical assessment (`oms`) from **2015**, all six SRR
scenarios. Each `hcrICES` is `FLStocks(foreach(id = sids))`. The 2015 stock
state is the assessment, not the open-loop \(F_{\mathrm{target}}\) run.

```{r scenario-hcr}
flr_pkgs = c("methods", "FLCore", "FLBRP", "icesdata", "FLBacktest", "FLasher")
ncores = max(1L, detectCores() - 4L)
cl = makeCluster(ncores)
registerDoParallel(cl)
tryCatch({
  clusterExport(cl, c("flr_pkgs", "om", "eqls", "eqs", "srDev",
                      "startYr", "endYr"), envir = environment())
  clusterEvalQ(cl, {
    suppressPackageStartupMessages(
      lapply(flr_pkgs, library, character.only = TRUE))
    NULL
  })
  scen2 = list()
  for (nm in names(eqls)) {
    clusterExport(cl, "nm", envir = environment())
    scen2[[nm]] = FLStocks(foreach(id = sids,
                                   .packages = flr_pkgs,
                                   .export = c("om", "eqls", "eqs", "srDev",
                                               "startYr", "endYr", "nm")) %dopar% {
      hcrICES(om[[id]], eql = eqls[[nm]][[id]],
              sr_deviances = srDev[[nm]][[id]],
              params = hcrParams(eqs[[id]]),
              start = startYr[id], end = endYr[id],
              err = NULL, implErr = 0)[[1]]
    })
    names(scen2[[nm]]) = sids
  }
}, finally = stopCluster(cl))
closedLoop = list("SRR" = scen2)

save(closedLoop, file=file.path(root, "data/results/03.2_closedLoop.RData"))
```

Comparison of historical, \(F_{\mathrm{tar}}\) and the HCR, time below
\(B_{\lim}\) / \(MSYB_{\mathrm{trigger}}\), current status and rebuild
times are in [`06.0_report.Rmd`](06.0_report.Rmd).
