---
title: "Rebuild under the ICES HCR"
subtitle: "20-year projections from the historical and closed-loop terminal states"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

Take each stock in its **terminal year**, extend it 20 years with `fwdWindow`
(FLBRP mean biology), and apply the ICES hockey-stick (`hcrICES`) with mean
recruitment. After the terminal year the observed recruitment-residual history
is switched off and biology is held at recent means, so this is mean
productivity, not a continuation of the historical productivity series.
Two starting states:

- **Historical** — latest WGCSE `FLStock` (`oms`)
- **Advice rule** — closed-loop `hcrICES` from
  [`04.2_closedLoop.Rmd`](04.2_closedLoop.Rmd)

The question is whether a stock that is still below \(MSYB_{\mathrm{trigger}}\)
recovers, how long that takes, and whether SSB reaches \(B_{\mathrm{MSY}}\).

HCR parameters (`ftar`, `btrig`, `blim`) are ICES values on `eqs`
(`eqls[["bh1"]]`, held fixed). The Operating Model SRR is **bh3**
(Beverton–Holt with an \(R_0\) prior from the hockey-stick).

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
root="C:/active/blueMarine"

dir.create(file.path(root, "tex/figs"),                  recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "data/results"),              recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "Rmd/cache", "04.3_rebuild"), recursive=TRUE, showWarnings=FALSE)

suppressPackageStartupMessages(suppressWarnings(library(knitr)))
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(root, "Rmd/cache", "04.3_rebuild/"),
  fig.path  = file.path(root, "tex/figs", "04.3_rebuild-"),
  fig.width = 9,
  fig.height= 5,
  dev       ="png")

Fig=0
Tab=0
```

```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(ggplotFL)
  library(ggplot2)
  library(FLasher)
  library(FLBacktest)
})

if (!exists("project_hcr", mode = "function"))
  suppressMessages(devtools::load_all("C:/active/flr/backtest", quiet = TRUE))

source(file.path(root, "R/plot_theme.R"))
theme_set(theme_bluemarine_html())
```

```{r load}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))
load(file.path(root, "data/results/03.2_closedLoop.RData"))
source(file.path(root, "R/om.R"))

sids = as.character(stocks$sid)
require_om_gate(root, sids)
eqs  = eqls[["bh1"]]
srr  = "bh3"
if (!srr %in% names(closedLoop[["SRR"]]) || !srr %in% names(eqls))
  stop("SRR branch '", srr, "' missing from closedLoop or eqls")
ar   = closedLoop[["SRR"]][[srr]]
nYears = 20L

ices_rp = function(id) {
  rp_hcr = refpts(eqs[[id]])
  rp_om  = refpts(eqls[[srr]][[id]])
  c(Blim  = c(rp_hcr["Blim",        "ssb"]),
    Btrig = c(rp_hcr["MSYBtrigger", "ssb"]),
    BMSY  = c(rp_om["msy",          "ssb"]),
    FMSY  = c(rp_hcr["FMSY",        "harvest"]))
}
```

# Terminal status

SSB and \(F\) in the last assessment year, before the 20-year projection.

```{r terminal, results='hide'}
starts = list(Historical = oms, "Advice rule" = ar)

terminal = ldply(names(starts), function(pol) {
  ldply(sids, function(id) {
    rp  = ices_rp(id)
    stk = starts[[pol]][[id]]
    yr  = dims(stk)$maxyear
    SB  = c(iterMedians(ssb(stk)[, ac(yr)]))
    Fb  = c(iterMedians(fbar(stk)[, ac(yr)]))
    data.frame(
      policy = pol, sid = id, year = yr,
      SSB_Blim  = SB / rp[["Blim"]],
      SSB_Btrig = SB / rp[["Btrig"]],
      SSB_BMSY  = SB / rp[["BMSY"]],
      F_FMSY    = Fb / rp[["FMSY"]],
      below_Btrig = SB < rp[["Btrig"]],
      below_BMSY  = SB < rp[["BMSY"]])
  })
})
terminal = merge(terminal, stocks[, c("sid", "name")], by = "sid")
terminal$name   = factor(terminal$name,   levels = levels(stocks$name))
terminal$policy = factor(terminal$policy, levels = names(starts))
terminal = terminal[order(terminal$name, terminal$policy), ]
```

# Projection

`fwdWindow` to terminal year \(+\) `r nYears`, then `hcrICES` with
deviations \(= 1\) (`FLBacktest::project_hcr`). Rebuild time is the first
projected year in which median SSB reaches the threshold
(`FLBacktest::years_to`). Already above is 0 years; `NA` means not
reached within `r nYears` years.

```{r project, results='hide'}
future = lapply(starts, function(stks) {
  FLStocks(setNames(llply(sids, function(id) {
    project_hcr(stks[[id]], eqls[[srr]][[id]], hcrParams(eqs[[id]]), nYears)
  }), sids))
})

rebuild = ldply(names(starts), function(pol) {
  ldply(sids, function(id) {
    rp    = ices_rp(id)
    maxyr = subset(terminal, sid == id & policy == pol)$year[1]
    s     = ssb(future[[pol]][[id]])
    term  = subset(terminal, sid == id & policy == pol)
    data.frame(
      policy = pol, sid = id,
      years_to_Btrig = if (isTRUE(term$below_Btrig[1]))
        years_to(s, rp[["Btrig"]], maxyr) else 0L,
      years_to_BMSY = if (isTRUE(term$below_BMSY[1]))
        years_to(s, rp[["BMSY"]], maxyr) else 0L)
  })
})
rebuild = merge(rebuild, terminal[
  , c("sid", "policy", "name", "SSB_Btrig", "SSB_BMSY")],
  by = c("sid", "policy"))
rebuild$name   = factor(rebuild$name,   levels = levels(stocks$name))
rebuild$policy = factor(rebuild$policy, levels = names(starts))
rebuild = rebuild[order(rebuild$name, rebuild$policy), ]

save(future, terminal, rebuild, nYears, sids, stocks, srr,
     file = file.path(root, "data/results/04.3_rebuild.RData"))
```

Trajectories of SSB, \(F\), recruitment and catch are shown in
[`06.0_report.Rmd`](06.0_report.Rmd).
