---
title: "SAM closed-loop counterfactual"
subtitle: "Methods — historical hcrICES with a simulated SAM observation model"
author: "Laurence Kell, Sea++"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

SAM closed-loop counterpart of [`04.2_closedLoop.Rmd`](04.2_closedLoop.Rmd).
This file is `04.4` because [`04.3_rebuild.Rmd`](04.3_rebuild.Rmd) is already
the 20-year rebuild from historical and perfect-information terminal states.

Same Operating Models as `04.2`, built in
[`02.0_condition_om.Rmd`](02.0_condition_om.Rmd). Same ICES hockey-stick
(`FLBacktest::hcrICES`). The difference is the observation model:

| Notebook | `err` | What the HCR sees |
|----------|-------|-------------------|
| `04.2` | `NULL` | Perfect information (true Operating Model status) |
| `04.4` | `FLIndex` | Yearly SAM fit on an Operating Model-derived survey |

This is **simulated full feedback**, not historical SAM peels. Each advice year
`hcrICES` updates the index from Operating Model numbers, calls
`FLfse::FLR_SAM` → `SAM2FLStock`, short-term forecasts, then `hcrFn`, and
implements the TAC on the Operating Model. Equivalent driver:
`backtest(..., method = "full")`. There is no separate SAM HCR.

ICES \(F_{\mathrm{tar}}\) / \(B_{\mathrm{trigger}}\) / \(B_{\lim}\) stay on
`eqs` (`eqls[["bh1"]]`). The Operating Model SRR is **bh3**, as in `04.3`.
`sr_deviances` are `srResiduals` from that bh3 `FLBRP`.

Stocks: Celtic Sea `cod.27.7e-k`, `whg.27.7b-ce-k`, `had.27.7b-k`, and Irish
Sea whiting `whg.27.7a` (SAM age-structured in `stocks.csv`). Irish Sea cod
and NEA mackerel are not SAM here.

Default `nits = 1` is a smoke iteration count. Increase `nits` for a
stochastic OEM; `indexCv` is 0.1 as in `FLBacktest::backtest`.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- bm_root()
root <- bm_root()

dir.create(file.path(root, "tex/figs"),                          recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "data/results"),                      recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "Rmd/cache", "04.4_closedLoop_sam"), recursive=TRUE, showWarnings=FALSE)

suppressPackageStartupMessages(suppressWarnings(library(knitr)))
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(root, "Rmd/cache", "04.4_closedLoop_sam/"),
  fig.path  = file.path(root, "tex/figs", "04.4_closedLoop_sam-"),
  fig.width = 9,
  fig.height= 5,
  dev       ="png")

Fig=0
Tab=0
```

```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(ggplotFL)
  library(ggplot2)
  library(FLasher)
  if (requireNamespace("devtools", quietly = TRUE) &&
  else
    library(FLBacktest)
  library(FLfse)
  library(stockassessment)
  library(doParallel)
  library(parallel)
  library(foreach)})

source(file.path(root, "R/plot_theme.R"))

theme_set(theme_bluemarine_html())
```

```{r load-om}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))
recDevs_file = file.path(root, "data/om/recDevs.RData")
if (file.exists(recDevs_file))
  load(recDevs_file)

om = oms
if (!exists("eqs")) eqs = eqls[["bh1"]]
srr = "bh3"
if (!srr %in% names(eqls))
  stop("SRR branch '", srr, "' missing from eqls — knit 02.0 first.",
       call. = FALSE)

sam_sids = c("cod.27.7e-k", "whg.27.7b-ce-k", "had.27.7b-k", "whg.27.7a")
sam_csv  = as.character(stocks$sid[grepl("SAM", stocks$assessment)])
sids     = sam_sids[sam_sids %in% names(om) & sam_sids %in% sam_csv]

nits    = 1L
indexCv = 0.1
smoke   = FALSE

if (!length(sids))
  stop("No SAM stocks in oms / stocks.csv.", call. = FALSE)

srDev   = FLQuants(setNames(lapply(sids, function(id)
  srResiduals(eqls[[srr]][[id]])), sids))
startYr = setNames(rep(2005, length(sids)), sids)
endYr   = setNames(sapply(sids, function(id) dims(om[[id]])$maxyear), sids)

if (isTRUE(smoke)) {
  sids_run = sids[1]
  endYr[sids_run] = min(startYr[sids_run] + 2L, endYr[sids_run])
} else {
  sids_run = sids
}
```

# Observation model

`samIndex` on ages excluding recruitment and the plus group (prototype
`_sam.Rmd`; `hcrICES` then updates `index[, year]` from Operating Model
`stock.n`). The index slot is catchability × lognormal noise (`indexCv`),
not a stored catch-at-age series — `hcrICES` overwrites it each year from
the Operating Model. FLSAM is not used; the assessment is `FLfse` +
`stockassessment`.

```{r oem}
samOem = function(stk, start, end, lag = 1, indexCv = 0.1, nits = 1L) {
  if (nits > 1L && dims(stk)$iter == 1)
    stk = propagate(stk, nits)
  stock.n(stk)[is.na(stock.n(stk))] = 1e3
  stock.n(stk) = qmax(stock.n(stk), 1e-2)

  amin = range(stk)["min"] + 1
  amax = range(stk)["plusgroup"] - 1
  idx  = samIndex(trim(stk, age = ac(amin:amax)))

  yrs = intersect(ac((start - lag):end), dimnames(index(idx))$year)
  index(idx)[] = 1
  if (length(yrs))
    index(idx)[, yrs] = 1
  index(idx) = index(idx) %*%
    rlnorm(nits, index(idx) %=% 0, indexCv) %*%
    rlnorm(nits, index(idx)[1] %=% 0, indexCv)
  idx
}
```

# Closed-loop `hcrICES`

One stock failure does not stop the rest (`try`). Workers load **FLfse**,
**stockassessment**, and **FLBacktest**. Timing: `start = 2005`,
`end = dims(om)$maxyear`, `lag = 1`, `interval = 1`, `implErr = 0`.

```{r scenario-hcr}
flr_pkgs = c("methods", "FLCore", "FLBRP", "icesdata", "FLasher",
             "FLfse", "stockassessment")
ncores = max(1L, min(length(sids_run), detectCores() - 4L))
cl = makeCluster(ncores)
registerDoParallel(cl)
closedLoopSam = list()
failed = character()
tryCatch({
  clusterExport(cl, c("flr_pkgs", "om", "eqls", "eqs", "srDev", "samOem",
                      "startYr", "endYr", "srr", "nits", "indexCv"),
                envir = environment())
  clusterEvalQ(cl, {
    suppressPackageStartupMessages(
      lapply(flr_pkgs, library, character.only = TRUE))
    if (requireNamespace("devtools", quietly = TRUE) &&
    else
      library(FLBacktest)
    NULL
  })
  runs = foreach(id = sids_run,
                 .packages = flr_pkgs,
                 .errorhandling = "pass",
                 .export = c("om", "eqls", "eqs", "srDev", "samOem",
                             "startYr", "endYr", "srr", "nits", "indexCv")) %dopar% {
    tryCatch({
      set.seed(1000L + match(id, names(om)))
      stk = om[[id]]
      if (nits > 1L && dims(stk)$iter == 1)
        stk = propagate(stk, nits)
      idx = samOem(stk, start = startYr[id], end = endYr[id],
                   lag = 1, indexCv = indexCv, nits = nits)
      hcrICES(stk, eql = eqls[[srr]][[id]],
              sr_deviances = srDev[[id]],
              params = hcrParams(eqs[[id]]),
              start = startYr[id], end = endYr[id],
              lag = 1, interval = 1,
              err = idx, implErr = 0)[[1]]
    }, error = function(e) e)
  }
  names(runs) = sids_run
  ok = vapply(runs, function(x) is(x, "FLStock"), logical(1))
  failed = sids_run[!ok]
  closedLoopSam = if (any(ok)) do.call(FLStocks, runs[ok]) else list()
}, finally = stopCluster(cl))

save(closedLoopSam, sids, sids_run, failed, nits, indexCv, srr, startYr, endYr,
     file = file.path(root, "data/results/04.4_closedLoop_sam.RData"))
```

# Comparison (optional)

SSB, \(F\) and catch against the historical assessment and, if present, the
perfect-information closed loop from `03.2_closedLoop.RData` (`04.2`).

```{r smoke-plot, fig.height=7, fig.width=9}
perfect_info = function(id, file) {
  if (!file.exists(file)) return(NULL)
  e = new.env()
  load(file, envir = e)
  if (!exists("closedLoop", envir = e, inherits = FALSE)) return(NULL)
  cl = e$closedLoop
  if (is.list(cl) && !is.null(cl[["SRR"]])) {
    nm = intersect(c(srr, "bh1"), names(cl[["SRR"]]))
    if (!length(nm)) nm = names(cl[["SRR"]])[1]
    if (length(nm) && !is.null(cl[["SRR"]][[nm[1]]][[id]]))
      return(cl[["SRR"]][[nm[1]]][[id]])
  }
  if (is(cl, "FLStocks") && id %in% names(cl))
    return(cl[[id]])
  NULL
}

pi_file = file.path(root, "data/results/03.2_closedLoop.RData")
if (is(closedLoopSam, "FLStocks") && length(closedLoopSam)) {
  for (id in names(closedLoopSam)) {
    stks = list(Historical = om[[id]], SAM = closedLoopSam[[id]])
    pi = perfect_info(id, pi_file)
    if (!is.null(pi))
      stks[["Perfect"]] = pi
    nm = as.character(subset(stocks, sid == id)$name)
    print(plot(do.call(FLStocks, stks),
               metrics = list(F = fbar, SB = ssb, C = catch),
               probs = c(0.75, 0.5, 0.25)) +
            labs(title = nm) +
            geom_vline(xintercept = startYr[id], linetype = 3,
                       colour = bm_col("muted")))
  }
}
```

**Figure `r Fig = Fig + 1; Fig`**. SAM closed loop versus historical (and
perfect information when `04.2` results are a `list(SRR = …)`). Trajectories
are the **bh3** Operating Model. Failures in `failed` did not produce a
stock.
