---
title: "All-stock digest"
subtitle: "Pick case studies after the generic backtest — not before"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

One page per knit of **every** stock that passed the Operating Model gate
and has open- and closed-loop results. Use this to choose which stocks
need a `06.1`-style case study. Do not put SAM, EqSim essays or
presentation layout here.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- bm_root()
root <- bm_root()
suppressPackageStartupMessages(suppressWarnings(library(knitr)))
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      error = FALSE)
```

```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(FLBRP)
  library(plyr)
  library(FLBacktest)
})
source(file.path(root, "R/om.R"))
load_om(root)
require_om_gate(root, as.character(stocks$sid))
load(file.path(root, "data/results/03.1_openLoop.RData"))
load(file.path(root, "data/results/03.2_closedLoop.RData"))
load(file.path(root, "data/results/03.3_lterm_eq.RData"))
eqs = eqls[["bh1"]]
srr = "bh1"
if (!srr %in% names(openLoop[["SRR"]]) || !srr %in% names(closedLoop[["SRR"]]))
  stop("openLoop/closedLoop SRR has no '", srr, "' — knit 04.1 and 04.2.",
       call. = FALSE)
```

```{r digest}
term = function(stk, yr) {
  c(SSB = c(ssb(stk)[, ac(yr)]), F = c(fbar(stk)[, ac(yr)]))
}

tab = ldply(as.character(stocks$sid), function(id) {
  if (is.null(openLoop[["SRR"]][[srr]][[id]]))
    stop("Missing openLoop$SRR$", srr, "$", id, " — knit 04.1.", call. = FALSE)
  if (is.null(closedLoop[["SRR"]][[srr]][[id]]))
    stop("Missing closedLoop$SRR$", srr, "$", id, " — knit 04.2.", call. = FALSE)
  yr = dims(oms[[id]])$maxyear
  rp = refpts(eqs[[id]])
  h  = term(oms[[id]], yr)
  o  = term(openLoop[["SRR"]][[srr]][[id]], yr)
  c_ = term(closedLoop[["SRR"]][[srr]][[id]], yr)
  data.frame(
    sid = id,
    name = as.character(stocks$name[match(id, stocks$sid)]),
    year = yr,
    gate = {
      p = lterm_eq$pass
      isTRUE(p$pass[match(id, as.character(p$sid))])
    },
    hist_SSB_Btrig = h["SSB"] / c(rp["MSYBtrigger", "ssb"]),
    ftar_SSB_Btrig = o["SSB"] / c(rp["MSYBtrigger", "ssb"]),
    ar_SSB_Btrig   = c_["SSB"] / c(rp["MSYBtrigger", "ssb"]),
    hist_F_Ftar    = h["F"] / c(rp["FMSY", "harvest"]),
    ftar_F_Ftar    = o["F"] / c(rp["FMSY", "harvest"]),
    ar_F_Ftar      = c_["F"] / c(rp["FMSY", "harvest"]),
    stringsAsFactors = FALSE)
})
```

**Table 1**. Terminal year versus ICES $B_{\mathrm{trigger}}$ and
$F_{\mathrm{target}}$. `bh1` branch. Case studies currently:
Irish Sea whiting and Celtic Sea cod (`06.1_TwoStocks.Rmd`).

```{r table}
knitr::kable(
  tab[, c("name", "sid", "year", "gate",
          "hist_SSB_Btrig", "ftar_SSB_Btrig", "ar_SSB_Btrig",
          "hist_F_Ftar")],
  digits = 2, row.names = FALSE,
  col.names = c("Stock", "SID", "Year", "Gate",
                "Hist SSB/$B_{trig}$", "$F_{tar}$ SSB/$B_{trig}$",
                "AR SSB/$B_{trig}$", "Hist $F/F_{tar}$"))
```
