---
title: "What would have happened if ICES advice had been followed?"
subtitle: "Generic Operating Model"
author: "Laurence Kell, Sea++"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: flatly
bibliography: refs.bib
---

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- bm_root()
root <- bm_root()

dir.create(file.path(root, "tex/figs"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "Rmd/cache", "06.2_generic"), recursive = TRUE,
           showWarnings = FALSE)

suppressPackageStartupMessages(suppressWarnings(library(knitr)))
knitr::opts_chunk$set(
  echo       = FALSE,
  warning    = FALSE,
  message    = FALSE,
  error      = FALSE,
  cache      = FALSE,
  cache.path = file.path(root, "Rmd/cache", "06.2_generic/"),
  fig.path   = file.path(root, "tex/figs", "06.2_generic-"),
  fig.width  = 9,
  fig.height = 5,
  dev        = "png")

Fig = 0
Tab = 0
```

```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(FLBRP)
  library(icesdata)
  library(FLRebuild)
  library(plyr)
  library(ggplot2)
  library(patchwork)
  library(openxlsx)
  library(FLasher)
  library(FLBacktest)
  library(ggplotFL)
  library(kableExtra)
})


source(file.path(root, "R/plot_theme.R"))
source(file.path(root, "R/om.R"))
source(file.path(root, "R/advice_sheets.R"))
theme_set(theme_bluemarine_html())
update_geom_defaults("line", list(linewidth = 1.1))

cols = c(Historical = "#000000", Ftarget = "#0072B2", "Advice rule" = "#D55E00")
metrics = list(F = fbar, SSB = ssb, Rec = rec, Catch = catch)
```

```{r load}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))
load(file.path(root, "data/results/03.1_openLoop.RData"))
load(file.path(root, "data/results/03.2_closedLoop.RData"))

sids = setdiff(as.character(stocks$sid), c("cod.27.7e-k", "whg.27.7a"))
if (!length(sids))
  stop("No stocks left after dropping the 06.1 case studies.", call. = FALSE)
stocks = stocks[match(sids, stocks$sid), , drop = FALSE]
stocks$name = factor(as.character(stocks$name),
                     levels = as.character(stocks$name))
require_om_gate(root, sids)
om_nm = "scn2"
# ICES HCR inputs (SAG Ftarget / Blim / Btrigger on bh1). Not the OM SRR.
eqs = eqls[["bh1"]]

om_eq = function(id) {
  eq = eqls[[om_nm]][[id]]
  if (is.null(eq))
    stop("Missing Operating Model eqls$", om_nm, "$", id, call. = FALSE)
  eq
}
om_sr = function(id) {
  eq = om_eq(id)
  sr_att = attributes(eq)$sr
  if (is.list(sr_att) && !is(sr_att, "FLSR")) sr_att = sr_att[[1]]
  if (!is(sr_att, "FLSR"))
    stop("No attached FLSR on eqls$", om_nm, "$", id, call. = FALSE)
  # as(FLBRP, "FLSR") uses this FLBRP's model and params.
  # The attached FLSR can carry one param iter per year, which breaks ffwd.
  as(eq, "FLSR")
}

nYears = 20L
startYr_ftar = setNames(vapply(sids, function(id)
  openloop_start(oms[[id]]), integer(1L)), sids)
startYr_ar = 2015L
endYr = setNames(as.integer(stocks$end[match(sids, stocks$sid)]), sids)

missing = character()
if (is.null(eqls[[om_nm]]))
  missing = c(missing, paste0("eqls$", om_nm))
for (id in sids) {
  if (is.null(oms[[id]])) missing = c(missing, paste0("oms$", id))
  if (is.null(openLoop[["SRR"]][[om_nm]][[id]]))
    missing = c(missing, paste0("openLoop$SRR$", om_nm, "$", id))
  if (is.null(closedLoop[["SRR"]][[om_nm]][[id]]))
    missing = c(missing, paste0("closedLoop$SRR$", om_nm, "$", id))
  sr_err = tryCatch({ om_sr(id); NULL }, error = function(e) conditionMessage(e))
  if (!is.null(sr_err)) missing = c(missing, sr_err)
}
if (length(missing))
  stop("Missing backtest objects: ", paste(missing, collapse = ", "), call. = FALSE)

advice = load_advice_sheets(root)
advice_vs_hcr = check_advice_vs_eqls(advice, eqs, sids)
```

# Introduction

> **Caveat.** This analysis is based on a generic Operating Model conditioned on a single stock–recruitment relationship. $B_{\mathrm{MSY}}$, $F_{\mathrm{MSY}}$ and the counterfactual trajectories are therefore conditional on that one hypothesis; the results may not be supported as stock-specific conclusions. Stock-specific SRRs for Celtic Sea cod and Irish Sea whiting are in [`06.1_TwoStocks.Rmd`](06.1_TwoStocks.Rmd).

What would historical and current status have looked like for these ICES Category 1 stocks if ICES advice had been implemented?

To answer this hypothetical question we conduct a **backtest**, where we took the most recent stock assessment, went back several years and then simulated i) fishing at the ICES $F_{\mathrm{target}}$ and ii) the ICES advice rule. This provides two counterfactual evolutions of each stock. The first is a projection or open-loop simulation, while the second is a closed-loop (i.e. with feed-back) simulation of the ICES hockey-stick advice rule.

Historical productivity, variability and regime shifts are preserved by reusing historical biological (natural mortality, maturity and mass-at-age) and fishery (selectivity) assumptions, and recruitment deviations. For the projections a stock–recruitment relationship, fitted to the historical stock assessment estimates, was assumed so that if spawning stock biomass increased so would the expected level of recruitment, but year-class variability was preserved. The same Operating Model hypothesis is used for each stock, i.e. a Beverton--Holt functional form with virgin recruitment based on the average level for SSB greater than $MSY_{Btrigger}$.

In the backtest the ICES reference points, i.e. ICES $B_{\lim}$, $B_{\mathrm{trigger}}$ and $F_{\mathrm{target}}$, are **operational HCR control points**. In the simulations (i.e. in the Operating Model) in contrast $B_{\mathrm{MSY}}$ and $F_{\mathrm{MSY}}$ are **performance metrics** conditional on the assumed SRR.

## ICES reference points

ICES derives $B_{\lim}$, $B_{\mathrm{trigger}}$, $F_{\lim}$, $F_{\mathrm{pa}}$ and
$F_{\mathrm{MSY}}$ using the \texttt{EqSim} software, together with estimates of SSB and
recruitment from stock assessments. Usually a hockey-stick stock recruitment relationship
is assumed, with constant recruitment above a breakpoint, below which recruitment declines
linearly to the origin [@mesnil2010continuous]; the backtest instead uses a biologically
plausible SRR for the Operating Model.

ICES commonly sets $B_{\lim}$ based on a two-line hockey-stick stock recruitment relationship; where there is a wide range of SSB and impaired recruitment is observed $B_{\lim}$ is set at the breakpoint of a segmented (hockey-stick) regression. The $F$ reference points are then conditioned on it: $F_{\lim}$ is the fishing mortality that drives equilibrium SSB to $B_{\lim}$ under deterministic dynamics,
while $F_{\mathrm{pa}}$ is taken as $F_{p.05}$, the largest $F$ for which the probability of
SSB falling below $B_{\lim}$ does not exceed 5\% in stochastic projections that include
assessment and advice error. $F_{\mathrm{MSY}}$ is obtained by maximising stochastic
long-term yield, but is then capped so that $F_{\mathrm{MSY}} \le F_{p.05}$, and
$F_{\mathrm{target}}$ is reduced linearly once SSB falls below $B_{\mathrm{trigger}}$
[@ices2021technical]. For many ICES stocks the hockey-stick breakpoint sits near or below
$B_{\mathrm{MSY}}$ under equilibrium assumptions, so that a limit can look like a target.
The precautionary cap and the advice rule are both based on $B_{\lim}$, therefore the
resulting $F$ reference points are not the same as those based on equilibrium or long-term projections since $\texttt{EqSim}$ derives its reference points by constraining them jointly. ICES $B_{\lim}$, $B_{\mathrm{trigger}}$ and $F_{\mathrm{target}}$ may not be consistent with the Operating Model values of $B_{\mathrm{MSY}}$.

The SSB–recruit “data” are assessment outputs, not independent observations [@Winker2022InsideOut], so if assessment assumptions change so will the reference points even if there has been no change in the actual stock.

Current ICES status is copied from the advice sheets in `docs/advice`. It is not scored on Operating Model $B_{\mathrm{MSY}}$.

```{r ices-status, results='asis', eval=FALSE}
for (id in sids) {
  sh = advice_row(advice, id)
  Tab = Tab + 1
  cat(sprintf(
    paste0("\n**Table %d**. %s (`%s`). ICES Advice %s (`%s`). %s Advice: %s.\n\n"),
    Tab,
    as.character(stocks$name[match(id, stocks$sid)]),
    id, sh$ices_advice_year, sh$sheet_file, sh$stock_dev, advice_catch_lab(sh)))
  print(advice_status_block(advice, id, stocks))
  cat("\n")
}
```

SAG $F_{\mathrm{MSY}}$, $B_{\lim}$ and $MSYB_{\mathrm{trigger}}$ on `eqls[["bh1"]]` (the HCR inputs used in the backtest) match these sheets, so the open- and closed-loop runs do not need to be re-knit for reference points.

ICES does not report $B_{\mathrm{MSY}}$. What it reports is a $B_{\lim}$ typology plus an EqSim (or equivalent) scan for $F_{\mathrm{MSY}}$, then $MSYB_{\mathrm{trigger}}=B_{\mathrm{pa}}$ unless a management plan says otherwise.

```{r ices-srr-tab}
rp_tab = ldply(sids, function(id) {
  sh = advice_row(advice, id)
  om = refpts(om_eq(id))
  data.frame(
    sid         = id,
    stock       = as.character(stocks$name[match(id, stocks$sid)]),
    sheet       = sh$ices_advice_year,
    Ftarget     = sh$FMSY,
    Blim        = sh$Blim,
    Bpa         = sh$Bpa,
    Fpa         = sh$Fpa,
    MSYBtrigger = sh$MSYBtrigger,
    BMSY        = c(om["msy", "ssb"]),
    FMSY_om     = c(om["msy", "harvest"]),
    Btrig_Bmsy  = sh$MSYBtrigger / c(om["msy", "ssb"]),
    blim_basis  = sh$blim_basis,
    fmsy_basis  = sh$fmsy_basis,
    stringsAsFactors = FALSE)
})
rp_tab = rp_tab[match(sids, rp_tab$sid), ]
hcr_ok = all(advice_vs_hcr$match)
if (!isTRUE(hcr_ok))
  stop("Sheet vs HCR mismatch slipped past the load check", call. = FALSE)
```

**Table `r Tab = Tab + 1; Tab`**. Technical basis of ICES $B_{\lim}$ and $F_{\mathrm{MSY}}$ as printed on the advice sheets in `docs/advice` (not the Operating Model). $F_{\lim}$ is not defined on these sheets. $MSYB_{\mathrm{trigger}}=B_{\mathrm{pa}}$ in every case.

```{r ices-srr-kable}
knitr::kable(
  rp_tab[, c("stock", "sid", "sheet", "blim_basis", "fmsy_basis")],
  row.names = FALSE,
  col.names = c("Stock", "SID", "Sheet", "$B_{lim}$", "$F_{MSY}$ (ICES)"))
```

Sources: the PDFs in `docs/advice`.

**Table `r Tab = Tab + 1; Tab`**. ICES operational points versus Operating Model $B_{\mathrm{MSY}}$ and $F_{\mathrm{MSY}}$ (scn2). $B_{\mathrm{trigger}}/B_{\mathrm{MSY}}$ well below one is the expected gap when the trigger is a $B_{\mathrm{pa}}$ proxy, not MSY [@winker2026rebuilding].

```{r rp-kable}
knitr::kable(
  rp_tab[, c("stock", "Blim", "MSYBtrigger", "Ftarget", "BMSY", "FMSY_om",
             "Btrig_Bmsy")],
  digits = c(0, 0, 0, 3, 0, 3, 2),
  row.names = FALSE,
  col.names = c("Stock", "$B_{lim}$", "$B_{trig}$", "$F_{target}$",
                "$B_{MSY}$", "$F_{MSY}$", "$B_{trig}/B_{MSY}$"),
  align = c("l", rep("r", 6)),
  escape = FALSE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed")) %>%
  add_header_above(c(" " = 1, "ICES" = 3, "OM" = 2, "Ratio" = 1),
                   escape = FALSE)
```

# What if advice had been followed?

The backtest poses the question that for a stock, assuming the observed and estimated biology, selection pattern, and recruitment, what would historical and current status have been if ICES advice had been implemented?

Black is the historical assessment, Blue is the $F_{\mathrm{target}}$ (open-loop) projection from 1990, Orange is applying the ICES advice rule from 2015. Horizontal lines are ICES $B_{\lim}$, $B_{\mathrm{trigger}}$ and $F_{\mathrm{target}}$. Red bands are Operating Model $SSB < B_{\mathrm{MSY}}$ and $F > F_{\mathrm{MSY}}$.

```{r ts-refpts, fig.height=8, fig.width=10}
for (id in sids) {
  nm   = as.character(stocks$name[match(id, stocks$sid)])
  ices = refpts(eqs[[id]])
  rp   = refpts(om_eq(id))
  x0   = dims(oms[[id]])$minyear
  yT   = endYr[id]

  p = plot(FLStocks(
        Historical    = window(oms[[id]], end = yT),
        Ftarget       = window(openLoop[["SRR"]][[om_nm]][[id]],
                               start = startYr_ftar[id], end = yT),
        "Advice rule" = window(closedLoop[["SRR"]][[om_nm]][[id]],
                               start = startYr_ar, end = yT)),
      metrics = metrics) +
      scale_colour_manual(values = cols) +
      scale_fill_manual(values = cols, guide = "none") +
      geom_flpar(data = FLPars(SSB = FLPar(Blim = c(ices["Blim", "ssb"]))),
                 x = x0 + 10, colour = "red", linetype = 2) +
      geom_flpar(data = FLPars(SSB = FLPar(Btrig = c(ices["MSYBtrigger", "ssb"]))),
                 x = x0 + 5, colour = bm_col("trigger"), linetype = 2) +
      geom_flpar(data = FLPars(F = FLPar(Ftarget = c(ices["FMSY", "harvest"]))),
                 x = x0 + 10, colour = bm_col("f_target"), linetype = 2) +
      labs(title = nm, x = "Year") +
      theme_bluemarine_html()
  print(shade_msy_status(p, xmin = x0, xmax = yT,
                         bmsy = c(rp["msy", "ssb"]),
                         fmsy = c(rp["msy", "harvest"])))
}
```

**Figure `r Fig = Fig + 1; Fig`**. Assessment (black), $F_{\mathrm{target}}$ from 1990 (blue), ICES advice rule from 2015 (orange). Coloured lines are ICES $B_{\lim}$, $B_{\mathrm{trigger}}$ and $F_{\mathrm{target}}$. Red bands are Operating Model $SSB < B_{\mathrm{MSY}}$ and $F > F_{\mathrm{MSY}}$.

# Historical, current and future comparisons

**Historical** — open loop, 1990 to the terminal year, $F=F_{\mathrm{target}}$, observed residuals.

**Current** — closed loop, 2015 to the terminal year, ICES advice rule, same residuals (productivity known).

**Future** — twenty-year rebuild from the terminal state with **mean** recruitment. That is not the same productivity history.

Terminal years differ by stock where the assessment series end in different years.

Grey dashed lines are Operating Model $R_0$, $B_{\mathrm{MSY}}$, MSY and $F_{\mathrm{MSY}}$. Coloured dashed lines are ICES $B_{\lim}$, $B_{\mathrm{trigger}}$ and $F_{\mathrm{target}}$. Red bands are Operating Model $SSB < B_{\mathrm{MSY}}$ and $F > F_{\mathrm{MSY}}$.

```{r backtest-closed-hcr, fig.height=8, fig.width=18}
for (id in sids) {
  nm   = as.character(stocks$name[match(id, stocks$sid)])
  om   = window(oms[[id]], end = endYr[id])
  eq   = om_eq(id)
  rp   = refpts(eq)
  ices = refpts(eqs[[id]])
  yT   = endYr[id]
  x0   = max(1990L, as.integer(dims(om)$minyear))

  om_par = FLPars(
    Rec   = FLPar(R0   = c(rp["virgin", "rec"])),
    SSB   = FLPar(Bmsy = c(rp["msy", "ssb"])),
    Catch = FLPar(MSY  = c(rp["msy", "yield"])),
    F     = FLPar(Fmsy = c(rp["msy", "harvest"])))
  ices_blim  = FLPars(SSB = FLPar(Blim  = c(ices["Blim", "ssb"])))
  ices_btrig = FLPars(SSB = FLPar(Btrig = c(ices["MSYBtrigger", "ssb"])))
  ices_ftar  = FLPars(F   = FLPar(Ftarget = c(ices["FMSY", "harvest"])))

  hist_fut = project_hcr(om, eq, hcrParams(eqs[[id]]), nYears)
  ar_fut   = project_hcr(window(closedLoop[["SRR"]][[om_nm]][[id]], end = yT), eq,
                         hcrParams(eqs[[id]]), nYears)
  yF = dims(hist_fut)$maxyear

  p_h = plot(FLStocks(
      Historical = window(om, start = x0, end = yT),
      Ftarget    = window(openLoop[["SRR"]][[om_nm]][[id]], start = x0, end = yT)),
    metrics = metrics) +
    scale_colour_manual(values = cols) +
    scale_fill_manual(values = cols, guide = "none") +
    geom_vline(xintercept = c(startYr_ftar[id], startYr_ar), linetype = 3) +
    geom_flpar(data = om_par, x = x0, colour = bm_col("muted"), linetype = 2) +
    geom_flpar(data = ices_blim,  x = 2010, colour = "red", linetype = 2) +
    geom_flpar(data = ices_btrig, x = 2010, colour = bm_col("trigger"), linetype = 2) +
    geom_flpar(data = ices_ftar,  x = 2010, colour = bm_col("f_target"), linetype = 2) +
    labs(title = paste(nm, "— Historical")) +
    theme_bluemarine_html()
  p_h = shade_msy_status(p_h, xmin = x0, xmax = yT,
                         bmsy = c(rp["msy", "ssb"]),
                         fmsy = c(rp["msy", "harvest"]))

  p_c = plot(FLStocks(
      Historical    = window(hist_fut, start = startYr_ar, end = yT),
      "Advice rule" = window(ar_fut, start = startYr_ar, end = yT)),
    metrics = metrics) +
    scale_colour_manual(values = cols) +
    scale_fill_manual(values = cols, guide = "none") +
    geom_vline(xintercept = c(startYr_ar, yT), linetype = 3) +
    geom_flpar(data = om_par, x = startYr_ar, colour = bm_col("muted"), linetype = 2) +
    geom_flpar(data = ices_blim,  x = startYr_ar + 2, colour = "red", linetype = 2) +
    geom_flpar(data = ices_btrig, x = startYr_ar + 2, colour = bm_col("trigger"), linetype = 2) +
    geom_flpar(data = ices_ftar,  x = startYr_ar + 2, colour = bm_col("f_target"), linetype = 2) +
    labs(title = "Current") +
    theme_bluemarine_html()
  p_c = shade_msy_status(p_c, xmin = startYr_ar, xmax = yT,
                         bmsy = c(rp["msy", "ssb"]),
                         fmsy = c(rp["msy", "harvest"]))

  p_f = plot(FLStocks(
      Historical    = window(hist_fut, start = 2000, end = yF),
      "Advice rule" = window(ar_fut, start = 2000, end = yF)),
    metrics = metrics) +
    scale_colour_manual(values = cols) +
    scale_fill_manual(values = cols, guide = "none") +
    geom_vline(xintercept = c(startYr_ftar[id], startYr_ar, yT), linetype = 3) +
    geom_flpar(data = om_par, x = 2000, colour = bm_col("muted"), linetype = 2) +
    geom_flpar(data = ices_blim,  x = 2010, colour = "red", linetype = 2) +
    geom_flpar(data = ices_btrig, x = 2010, colour = bm_col("trigger"), linetype = 2) +
    geom_flpar(data = ices_ftar,  x = 2010, colour = bm_col("f_target"), linetype = 2) +
    labs(title = "Future") +
    theme_bluemarine_html()
  p_f = shade_msy_status(p_f, xmin = 2000, xmax = yF,
                         bmsy = c(rp["msy", "ssb"]),
                         fmsy = c(rp["msy", "harvest"]))

  print((p_h | p_c | p_f) +
          plot_layout(widths = c(max(1, yT - x0), max(1, yT - startYr_ar),
                                 max(1, yF - 2000)),
                      guides = "collect") &
          theme(legend.position = "bottom"))
}
```

**Figure `r Fig = Fig + 1; Fig`**. Historical (left), Current (middle), Future (right). Middle panel is a zoom of the right-hand series. Red bands are Operating Model $SSB < B_{\mathrm{MSY}}$ and $F > F_{\mathrm{MSY}}$.

# Productivity Regimes

Catch is not the only driver, there have been changes in recruitment level independent of biomass and in growth. The Operating Model used a stock–recruitment relationship and residuals based on the estimated recruitment. Those residuals, with time-varying $M$, maturity and mass-at-age, determine the observed productivity history [@Lindmark2026productivity]. If $F$ falls, SSB can rise, which changes expected recruitment; while the residual history stays the same.

```{r rec-devs, fig.height=8, fig.width=9}
if (!exists("rod", mode = "function"))
  stop("STARS rod() is missing; load FLRebuild", call. = FALSE)

rsdl = FLQuants(lapply(sids, function(id) {
  r = residuals(om_sr(id))
  if (!any(is.finite(c(r))))
    stop("SRR residuals are not finite for ", id, call. = FALSE)
  window(r, end = endYr[id])
}))
names(rsdl) = sids

rods = ldply(sids, function(id) {
  rd = rod(rsdl[[id]])
  if (!is.data.frame(rd) || !nrow(rd) || !("regime" %in% names(rd)))
    stop("STARS/regime polygons missing for ", id, call. = FALSE)
  cbind(sid = id, rd)
})

stock_levels = as.character(stocks$name[match(sids, stocks$sid)])

print(
  ggplot(transform(
           merge(transform(as.data.frame(rsdl, drop = TRUE), sid = qname),
                 stocks, by = "sid"),
           name = factor(as.character(name), levels = stock_levels))) +
    geom_polygon(aes(year, data, group = regime),
                 fill = bm_col("strip_fill"), colour = bm_col("advice"),
                 linewidth = 0.25, alpha = 0.55,
                 data = transform(
                   merge(rods, stocks, by = "sid"),
                   name = factor(as.character(name), levels = stock_levels))) +
    geom_hline(yintercept = 0, colour = bm_col("ref_strong"),
               linewidth = 0.35) +
    geom_line(aes(year, data), colour = bm_col("ink"), linewidth = 0.45) +
    geom_point(aes(year, data), colour = bm_col("ink"), size = 0.9) +
    facet_grid(name ~ ., scales = "free_y") +
    labs(x = "Year", y = "log residual") +
    theme_bluemarine_html() +
    theme_facet_rows_rotated())
```

**Figure `r Fig = Fig + 1; Fig`**. Log recruitment residuals from the Operating Model SRR (scn2), with STARS regime means (shaded). These are the deviations replayed in the historical backtest.

`b0dyn` replays the assessment at $F=0$ with the same SRR, life history and residuals: dynamic unfished SSB given the productivity that occurred [@Lindmark2026productivity].

```{r b0dyn, fig.height=10, fig.width=16}
ps = vector("list", length(sids))
names(ps) = sids
for (id in sids) {
  stk  = window(oms[[id]], end = endYr[id])
  sr   = om_sr(id)
  stk0 = b0dyn(stk, sr)
  ps[[id]] =
    ggplot(FLQuants(SSB = ssb(stk), b0dyn = ssb(stk0)),
           aes(year, data, colour = qname)) +
      geom_line() +
      scale_colour_manual(values = c(SSB = "#000000", b0dyn = bm_col("catch"))) +
      labs(title = as.character(stocks$name[match(id, stocks$sid)]),
           x = "Year", y = "Biomass", colour = NULL) +
      theme_bluemarine_html()
}
print(wrap_plots(ps, ncol = 2, guides = "collect") &
        theme(legend.position = "bottom"))
```

**Figure `r Fig = Fig + 1; Fig`**. Assessment SSB (black) and dynamic unfished SSB (red). The gap is depletion against a productivity-varying unfished stock.

```{r pf-rec, fig.height=8, fig.width=14}
eq_om = lapply(sids, om_eq)
names(eq_om) = sids

pf = ldply(sids, function(id) {
  cbind(sid = id, om = om_nm,
        model.frame(FLQuants(eq_om[[id]], SSB = ssb, Yield = catch, Recruits = rec),
                    drop = TRUE))
})
pf = merge(pf, stocks[, c("sid", "name")], by = "sid")
pf$name = factor(pf$name, levels = levels(stocks$name))

obs = ldply(sids, function(id) {
  cbind(sid = id,
        model.frame(FLQuants(eq_om[[id]],
                             SSB = ssb.obs, Yield = catch.obs, Recruits = rec.obs),
                    drop = TRUE))
})
obs = merge(obs, stocks[, c("sid", "name")], by = "sid")
obs$name = factor(obs$name, levels = levels(stocks$name))

ggplot(pf, aes(SSB, Recruits)) +
  facet_wrap(~ name, scales = "free", ncol = 2) +
  geom_path(data = obs, aes(SSB, Recruits), inherit.aes = FALSE,
            colour = bm_col("muted"), linewidth = 0.35, alpha = 0.75) +
  geom_point(data = obs, aes(SSB, Recruits), inherit.aes = FALSE,
             colour = bm_col("ink"), size = 1.1, alpha = 0.45, stroke = 0) +
  geom_line(colour = bm_col("catch"), linewidth = 0.75) +
  labs(x = "SSB (t)", y = "Recruits") +
  theme_bluemarine_html()
```

**Figure `r Fig = Fig + 1; Fig`**. SRRs: Operating Model stock–recruitment relationships (scn2) with observed pairs.

```{r pf-yield, fig.height=8, fig.width=14}
ggplot(pf, aes(SSB, Yield)) +
  facet_wrap(~ name, scales = "free", ncol = 2) +
  geom_path(data = obs, aes(SSB, Yield), inherit.aes = FALSE,
            colour = bm_col("muted"), linewidth = 0.35, alpha = 0.75) +
  geom_point(data = obs, aes(SSB, Yield), inherit.aes = FALSE,
             colour = bm_col("ink"), size = 1.1, alpha = 0.45, stroke = 0) +
  geom_line(colour = bm_col("catch"), linewidth = 0.75) +
  labs(x = "SSB (t)", y = "Yield (t)") +
  theme_bluemarine_html()
```

**Figure `r Fig = Fig + 1; Fig`**. Production functions: equilibrium yield versus SSB (scn2). $B_{\mathrm{MSY}}$ and $F_{\mathrm{MSY}}$ on these curves are Operating Model properties, not ICES control points.

```{r export, results="hide"}
dir.create(file.path(root, "data/results"), recursive = TRUE, showWarnings = FALSE)

dat = NULL
for (id in sids) {
  om = oms[[id]]
  ft = openLoop[["SRR"]][[om_nm]][[id]]
  ar = closedLoop[["SRR"]][[om_nm]][[id]]
  if (is.null(om) || is.null(ft) || is.null(ar))
    stop("Missing time series for Excel export: ", id, call. = FALSE)
  dat = rbind(dat,
    cbind(What = "ICES",    sid = id,
          model.frame(FLCore::metrics(om, list(SSB = ssb, Catch = catch)),
                      drop = TRUE)),
    cbind(What = "Ftarget", sid = id,
          model.frame(FLCore::metrics(ft, list(SSB = ssb, Catch = catch)),
                      drop = TRUE)),
    cbind(What = "AR",      sid = id,
          model.frame(FLCore::metrics(ar, list(SSB = ssb, Catch = catch)),
                      drop = TRUE)))
}
if (is.null(dat) || !nrow(dat))
  stop("Excel time series sheet is empty", call. = FALSE)

ref = ldply(sids, function(id) {
  ices = refpts(eqs[[id]])
  om   = refpts(om_eq(id))
  data.frame(
    sid         = id,
    stock       = as.character(stocks$name[match(id, stocks$sid)]),
    om          = om_nm,
    Ftarget     = c(ices["FMSY",        "harvest"]),
    Blim        = c(ices["Blim",        "ssb"]),
    MSYBtrigger = c(ices["MSYBtrigger", "ssb"]),
    BMSY        = c(om["msy",           "ssb"]),
    FMSY        = c(om["msy",           "harvest"]),
    stringsAsFactors = FALSE)
})
if (!nrow(ref))
  stop("Excel refpts sheet is empty", call. = FALSE)

xlsx = file.path(root, "data/results/06.2.xlsx")
wb = createWorkbook()
addWorksheet(wb, "all")
writeData(wb, "all", dat)
freezePane(wb, "all", firstRow = TRUE)
addWorksheet(wb, "refpts")
writeData(wb, "refpts", ref)
freezePane(wb, "refpts", firstRow = TRUE)
saveWorkbook(wb, xlsx, overwrite = TRUE)
```
