---
title: "Open-loop benchmark projections"
subtitle: "Methods — historical fwd at ICES Ftar and OM FMSY, six SRRs"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

**Methods — Benchmark projections** (`tex/project-markup.tex`): historical
open-loop `fwd` from 2015 (`startYr_ar`, the closed-loop HCR start) to the
terminal year, compared with the realised assessment.

Two constant-$F$ arms are run for each stock and each OM / SRR scenario in
`eqls`:

1. **ICES $F_{\mathrm{tar}}$** — published ICES $F_{\mathrm{MSY}}$ from `eqs` /
   `bh1` (SAG), held fixed across all SRR branches. Saved under
   `openLoop$SRR` / `openLoop$Recruits`.
2. **OM $F_{\mathrm{MSY}}$** — that scenario's own equilibrium $F_{\mathrm{MSY}}$
   from `refpts(eqls[[srr]][[sid]])["msy", "harvest"]`. Saved under
   `openLoop$OM_FMSY` / `openLoop$OM_FMSY_Recruits`.

Same Operating Models as [`02.2_closedLoop.Rmd`](02.2_closedLoop.Rmd),
built in [`01.0_condition_om.Rmd`](01.0_condition_om.Rmd) (`oms`, `eqls`).
The only difference from the closed loop is the harvest policy: constant $F$
here, ICES HCR there, both from 2015 and both starting from the historical
assessment (`oms`).

| Uncertainty | Operating Model | `fwd` residuals | Object |
|-------------|:---------------:|-----------------|--------|
| `rec` | `oms` | historical rec / geometric mean | `Recruits` / `OM_FMSY_Recruits` |
| `srr` | `oms` | `srResiduals(eq)` | `SRR` / `OM_FMSY` |

ICES $F_{\mathrm{tar}}$ is held fixed; each SRR supplies its own biology /
stock–recruitment and (for the OM arm) its own $F_{\mathrm{MSY}}$. Historical
projections keep the assessment's year-specific $M$, mass and maturity and
replay recruitment residuals, so the productivity that actually occurred is
held common across policies.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
# knitr is already loaded by rmarkdown::render; do not library() a second copy.
if (!"knitr" %in% loadedNamespaces())
  suppressPackageStartupMessages(library(knitr))
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- data_root()
here <- clean_root()
dir.create(file.path(here, "tex/figs"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(root, "data/results"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(here, "Rmd/cache", "02.1_openLoop"), recursive = TRUE, showWarnings = FALSE)
knitr::opts_chunk$set(
  echo      = FALSE, 
  warning   = FALSE, 
  message   = FALSE, 
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(here, "Rmd/cache", "02.1_openLoop/"),
  fig.path  = file.path(here, "tex/figs", "02.1_openLoop-"),
  fig.width = 9, 
  fig.height= 5, 
  dev       ="png")
Fig=0
Tab=0
```


```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(ggplot2)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(FLBacktest)
  library(FLasher)
  library(doParallel)
  library(parallel)
  library(foreach)
})
source(file.path(here, "R/constants.R"))
source(file.path(here, "R/loadStocks.R"))
source(file.path(here, "R/om.R"))
source(file.path(here, "R/plot_traj.R"))
theme_set(theme_paper())
```

```{r load-om}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))
if (!exists("eqs")) {
  if (!"bh1" %in% names(eqls))
    stop("eqs missing and eqls[['bh1']] not found — knit 01.0 first.",
         call. = FALSE)
  eqs = eqls[["bh1"]]
}
sids = as.character(stocks$sid)
source(file.path(here, "R/om.R"))
require_om_gate(root, sids)
om   = oms
srDev = lapply(eqls, function(eq) FLQuants(lapply(eq, srResiduals)))
recDev = FLQuants(lapply(oms, function(x) {
  r = rec(x)
  r %/% exp(mean(log(c(r)), na.rm = TRUE))
}))
startYr = setNames(rep(startYr_ar, length(sids)), sids)
endYr   = setNames(sapply(om, function(x) dims(x)$maxyear), names(om))
too_short = sids[sapply(om[sids], function(x)
  dims(x)$minyear >= startYr_ar || dims(x)$maxyear < startYr_ar)]
if (length(too_short))
  stop("Open-loop start ", startYr_ar, " is outside the assessment years for: ",
       paste(too_short, collapse = ", "), call. = FALSE)
```

# Open-loop `fwd`

Two constant-$F$ benchmarks on the historical assessment Operating Models
(`oms`), all SRR scenarios in `eqls`. Each `fwd` is
`FLStocks(foreach(id = sids))`. Simulations start in 2015 (`startYr_ar`),
the same year as the closed-loop HCR runs.

**Saved object** (`data/results/02.1_openLoop.RData`):

- `openLoop$Recruits[[srr]][[sid]]` — ICES $F_{\mathrm{tar}}$, `recDev`
- `openLoop$SRR[[srr]][[sid]]` — ICES $F_{\mathrm{tar}}$, `srDev` (paper default)
- `openLoop$OM_FMSY_Recruits[[srr]][[sid]]` — OM $F_{\mathrm{MSY}}$, `recDev`
- `openLoop$OM_FMSY[[srr]][[sid]]` — OM $F_{\mathrm{MSY}}$, `srDev`

```{r scenario-fwd}
flr_pkgs = c("methods", "FLCore", "FLBRP", "icesdata", "FLBacktest", "FLasher")
ncores = max(1L, detectCores() - 4L)
libs = .libPaths()
cl = makeCluster(ncores)
registerDoParallel(cl)
tryCatch({
  clusterExport(cl, c("flr_pkgs", "libs"), envir = environment())
  clusterEvalQ(cl, {
    .libPaths(libs)
    suppressPackageStartupMessages(
      lapply(flr_pkgs, library, character.only = TRUE))
    NULL
  })
  clusterExport(cl, c("om", "eqls", "eqs", "recDev", "srDev",
                      "startYr", "endYr", "sids"), envir = environment())
  scen       = list()
  scen2      = list()
  scen_om    = list()
  scen2_om   = list()
  # Reference OM (srr_om) plus the six SRR robustness branches (not scn0–scn2).
  srr_run = intersect(unique(c(srr_om, "bh1", "bh2", "bh3", "rk1", "rk2", "rk3")),
                      names(eqls))
  if (!srr_om %in% srr_run)
    stop("eqls has no reference OM '", srr_om, "' — knit 01.0 first.",
         call. = FALSE)
  for (nm in srr_run) {
    clusterExport(cl, "nm", envir = environment())

    # --- ICES Ftar (published SAG FMSY on eqs / bh1), fixed across SRRs ---
    scen[[nm]] = FLStocks(foreach(id = sids, .packages = flr_pkgs,
                                  .export = c("om", "eqls", "eqs", "recDev",
                                              "startYr", "endYr", "nm")) %dopar% {
      fices = c(refpts(eqs[[id]])["FMSY", "harvest"])
      ftrg  = window(fbar(om[[id]]), start = startYr[id], end = endYr[id]) %=% fices
      FLasher::fwd(om[[id]], fbar = ftrg, sr = eqls[[nm]][[id]], residuals = recDev[[id]])
    })
    names(scen[[nm]]) = sids

    scen2[[nm]] = FLStocks(foreach(id = sids, .packages = flr_pkgs,
                                   .export = c("om", "eqls", "eqs", "srDev",
                                               "startYr", "endYr", "nm")) %dopar% {
      fices = c(refpts(eqs[[id]])["FMSY", "harvest"])
      ftrg  = window(fbar(om[[id]]), start = startYr[id], end = endYr[id]) %=% fices
      FLasher::fwd(om[[id]], fbar = ftrg, sr = eqls[[nm]][[id]],
          residuals = srDev[[nm]][[id]])
    })
    names(scen2[[nm]]) = sids

    # --- OM FMSY from this scenario's FLBRP (msy harvest) ---
    scen_om[[nm]] = FLStocks(foreach(id = sids, .packages = flr_pkgs,
                                     .export = c("om", "eqls", "recDev",
                                                 "startYr", "endYr", "nm")) %dopar% {
      fom  = c(refpts(eqls[[nm]][[id]])["msy", "harvest"])
      ftrg = window(fbar(om[[id]]), start = startYr[id], end = endYr[id]) %=% fom
      FLasher::fwd(om[[id]], fbar = ftrg, sr = eqls[[nm]][[id]], residuals = recDev[[id]])
    })
    names(scen_om[[nm]]) = sids

    scen2_om[[nm]] = FLStocks(foreach(id = sids, .packages = flr_pkgs,
                                      .export = c("om", "eqls", "srDev",
                                                  "startYr", "endYr", "nm")) %dopar% {
      fom  = c(refpts(eqls[[nm]][[id]])["msy", "harvest"])
      ftrg = window(fbar(om[[id]]), start = startYr[id], end = endYr[id]) %=% fom
      FLasher::fwd(om[[id]], fbar = ftrg, sr = eqls[[nm]][[id]],
          residuals = srDev[[nm]][[id]])
    })
    names(scen2_om[[nm]]) = sids
  }
}, finally = stopCluster(cl))
ftar  = scen[["bh1"]]
ftar2 = scen2[["bh1"]]
openLoop = list(
  Recruits         = scen,
  SRR              = scen2,
  OM_FMSY_Recruits = scen_om,
  OM_FMSY          = scen2_om)
save(openLoop, startYr, endYr,
     file = file.path(root, "data/results/02.1_openLoop.RData"))
```

# Trajectories

Historical vs the two open-loop arms for \(F\), recruitment, SSB and catch,
laid out as in [`02.2_closedLoop.Rmd`](02.2_closedLoop.Rmd). Each series is
divided by its Historical mean (plotted window). Facets: quantity \(\times\)
stock. Horizontal line at 1; ICES / OM reference points on the same scale.
Vertical dotted line: open-loop start (`r startYr_ar`).

```{r traj-plot, fig.height=9, fig.width=11}
labs = setNames(as.character(stocks$name), as.character(stocks$sid))
srr_plot = if (srr_om %in% names(openLoop[["SRR"]])) srr_om else
  names(openLoop[["SRR"]])[1]
print(plot_hcr_vs_hist(
  om, list("ICES Ftar" = openLoop[["SRR"]][[srr_plot]]), eqs,
  list("ICES Ftar" = eqls[[srr_plot]]),
  sids = sids, labels = labs, start_ar = startYr_ar, arm_fmt = "%s"))
```

**Figure `r Fig = Fig + 1; Fig`**. Open-loop constant \(F\) at ICES
\(F_{\mathrm{tar}}\) (`r srr_plot`) vs historical `FLStock`. On the reference
OM (`bhw`) OM \(F_{\mathrm{MSY}}\) is the ICES \(F_{\mathrm{MSY}}\), so the
OM \(F_{\mathrm{MSY}}\) arm is the same run and is not drawn.
