---
title: "Closed-loop counterfactual simulation"
subtitle: "Methods — historical hcrICES, perfect information, funder OM bh3"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

**Methods — Counterfactual simulation** (`tex/project-markup.tex`): historical
closed-loop `hcrICES` from 2015 to the terminal year, starting from the
realised assessment (`oms`), compared with the realised assessment.

Same Operating Models as [`02.1_openLoop.Rmd`](02.1_openLoop.Rmd),
built in [`01.0_condition_om.Rmd`](01.0_condition_om.Rmd) (`oms`, `eqls`).
The only difference is the harvest policy: ICES HCR here, constant ICES
\(F_{\mathrm{MSY}}\) there. The Advice rule does **not** start from the
open-loop \(F_{\mathrm{target}}\) trajectory.

## Uncertainty treatment

Process / structural uncertainty is carried by **recruitment deviations** and,
for peer review, by **alternative SRRs**. The funder draft runs the Operating
Model SRR **bh3** only (`R/constants.R::srr_om`), replaying `srResiduals` from
that equilibrium. Set `BM_CLOSED_SRR=all` to expand the same HCR loop across
`bh1`–`bh3` / `rk1`–`rk3` (still reusing each branch's recruitment residuals).
Short-cut assessment error and SAM sit downstream:
[`02.3_closedLoop_shortcut.Rmd`](02.3_closedLoop_shortcut.Rmd),
[`02.4.1_sam.Rmd`](02.4.1_sam.Rmd) / [`02.4.2_closedLoop_sam.Rmd`](02.4.2_closedLoop_sam.Rmd).

| Layer | What varies | Default funder run |
|-------|-------------|--------------------|
| Recdevs | `srResiduals(eqls[[srr]][[id]])` | **bh3** residuals |
| Alternative SRRs | OM biology / stock–recruitment | **bh3** only (`BM_CLOSED_SRR=all` for six) |
| Assessment OEM | `err` in `hcrICES` | `NULL` (perfect information) |

ICES \(F_{\mathrm{tar}}\) / \(B_{\mathrm{trigger}}\) / \(B_{\lim}\) are held
fixed (`hcrParams` from `eqs`). The advised catch carries the ICES TAC
stability clause (`R/constants.R::hcr_tac_bounds`): year-on-year change limited to
\(\pm 20\%\) of the previous catch when SSB is at or above
\(MSYB_{\mathrm{trigger}}\), and unconstrained below it
(`hcrICES(bndTac, bndWhen = "btrig")`). Historical projections keep the
assessment's year-specific \(M\), mass and maturity and replay recruitment
residuals, so the productivity that actually occurred is held common across
policies.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- data_root()
here <- clean_root()

dir.create(file.path(here, "tex/figs"),                     recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "data/results"),                 recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(here, "Rmd/cache", "02.2_closedLoop"), recursive=TRUE, showWarnings=FALSE)

# knitr is already loaded by rmarkdown::render; do not library() a second copy.
if (!"knitr" %in% loadedNamespaces())
  suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(here, "Rmd/cache", "02.2_closedLoop/"),
  fig.path  = file.path(here, "tex/figs", "02.2_closedLoop-"),
  fig.width = 9,
  fig.height= 5,
  dev       ="png")

Fig=0
Tab=0
```


```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(ggplot2)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(FLasher)
  library(FLBacktest)
  library(doParallel)
  library(parallel)
  library(foreach)})

source(file.path(here, "R/constants.R"))
source(file.path(here, "R/om.R"))
source(file.path(here, "R/plot_traj.R"))
theme_set(theme_paper())
```

```{r load-om}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))

sids = as.character(stocks$sid)
om   = oms
if (!exists("eqs")) {
  if (!"bh1" %in% names(eqls))
    stop("eqs missing and eqls[['bh1']] not found — knit 01.0 first.",
         call. = FALSE)
  eqs = eqls[["bh1"]]
}
source(file.path(here, "R/om.R"))
require_om_gate(root, sids)

srDev = lapply(eqls, function(eq) FLQuants(lapply(eq, srResiduals)))
startYr = setNames(rep(2015, length(sids)), sids)
endYr   = setNames(sapply(om, function(x) dims(x)$maxyear), names(om))
# TAC stability clause (R/om.R): +/-20% year-on-year when SSB >= MSY Btrigger
bndTac  = hcr_tac_bounds()
bndWhen = hcr_tac_bnd_when()

# Funder default: one SRR (bh3). BM_CLOSED_SRR=all → six branches.
srr6 = c("bh1", "bh2", "bh3", "rk1", "rk2", "rk3")
srr_env = Sys.getenv("BM_CLOSED_SRR", unset = srr_om)
if (identical(tolower(trimws(srr_env)), "all")) {
  srr_run = intersect(srr6, names(eqls))
} else {
  srr_run = intersect(trimws(strsplit(srr_env, ",", fixed = TRUE)[[1]]),
                      names(eqls))
}
if (!length(srr_run))
  stop("No SRR from BM_CLOSED_SRR=", srr_env, " found in eqls.", call. = FALSE)
```

# Closed-loop `hcrICES`

ICES HCR on the historical assessment (`oms`) from **2015**, Operating Model
SRR(s): `r paste(srr_run, collapse = ", ")`. Perfect information (`err = NULL`).
Each `hcrICES` is `FLStocks(foreach(id = sids))`. The 2015 stock state is the
assessment, not the open-loop \(F_{\mathrm{target}}\) run.

```{r scenario-hcr}
flr_pkgs = c("methods", "FLCore", "FLBRP", "icesdata", "FLBacktest", "FLasher")
ncores = max(1L, detectCores() - 4L)
libs = .libPaths()
cl = makeCluster(ncores)
registerDoParallel(cl)
tryCatch({
  clusterExport(cl, c("flr_pkgs", "libs", "om", "eqls", "eqs", "srDev",
                      "startYr", "endYr", "bndTac", "bndWhen"),
                envir = environment())
  clusterEvalQ(cl, {
    .libPaths(libs)
    suppressPackageStartupMessages(
      lapply(flr_pkgs, library, character.only = TRUE))
    NULL
  })
  scen2 = list()
  for (nm in srr_run) {
    clusterExport(cl, "nm", envir = environment())
    scen2[[nm]] = FLStocks(foreach(id = sids,
                                   .packages = flr_pkgs,
                                   .export = c("om", "eqls", "eqs", "srDev",
                                               "startYr", "endYr", "nm",
                                               "bndTac", "bndWhen")) %dopar% {
      hcrICES(om[[id]], eql = eqls[[nm]][[id]],
              sr_deviances = srDev[[nm]][[id]],
              params = hcrParams(eqs[[id]]),
              start = startYr[id], end = endYr[id],
              err = NULL, implErr = 0,
              bndTac = bndTac, bndWhen = bndWhen)[[1]]
    })
    names(scen2[[nm]]) = sids
  }
}, finally = stopCluster(cl))
closedLoop = list("SRR" = scen2, bndTac = bndTac, bndWhen = bndWhen,
                  srr_run = srr_run)

save(closedLoop, file=file.path(root, "data/results/02.2_closedLoop.RData"))
```

# Trajectories

Historical vs Advice rule for \(F\), recruitment, SSB and catch. Each
series is divided by its Historical mean (plotted window); panels share
a relative \(y\)-scale. Facets: quantity \(\times\) stock. Horizontal
line at 1; ICES / OM reference points on the same scale. Vertical dotted
line: HCR start (2015).

```{r traj-plot, fig.height=9, fig.width=11}
labs = setNames(as.character(stocks$name), as.character(stocks$sid))
srr_plot = if (srr_om %in% names(closedLoop[["SRR"]])) srr_om else
  names(closedLoop[["SRR"]])[1]
print(plot_hcr_vs_hist(
  om, closedLoop[["SRR"]][[srr_plot]], eqs, eqls[[srr_plot]],
  sids = sids, labels = labs, start_ar = 2015L,
  advice_lab = "Advice rule"))
```

**Figure `r Fig = Fig + 1; Fig`**. Perfect-information closed loop
(`r srr_plot`) vs historical `FLStock`.

```{r traj-plot-srr, fig.height=9, fig.width=11, eval = length(closedLoop[["SRR"]]) > 1L}
print(plot_hcr_vs_hist(
  om, closedLoop[["SRR"]], eqs, eqls[names(closedLoop[["SRR"]])],
  sids = sids, labels = labs, start_ar = 2015L))
```

```{r traj-srr-caption, results='asis', eval = length(closedLoop[["SRR"]]) > 1L}
Fig <- Fig + 1
cat("\n\n**Figure ", Fig,
    "**. Same contrast with one Advice-rule line per SRR.\n\n", sep = "")
```

Comparison of historical, \(F_{\mathrm{tar}}\) and the HCR, time below
\(B_{\lim}\) / \(MSYB_{\mathrm{trigger}}\), current status and rebuild
times are in [`04.0_paper_figures.Rmd`](04.0_paper_figures.Rmd).
