---
title: "Closed-loop short-cut OEM"
subtitle: "Methods — hcrICES with EqSim-style assessment error, between perfect and SAM"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

Same historical closed-loop ICES HCR as
[`02.2_closedLoop.Rmd`](02.2_closedLoop.Rmd) (perfect information), but with a
**short-cut** observation-error model: multiplicative SSB assessment error from
`FLBacktest::eqSimErr`. Full SAM assessment OEM is
[`02.4.1_sam.Rmd`](02.4.1_sam.Rmd) / [`02.4.2_closedLoop_sam.Rmd`](02.4.2_closedLoop_sam.Rmd).

Default Operating Model SRR is **bh3** (`BM_CLOSED_SRR`, same as `02.2`).
HCR parameters stay on `eqs` (`eqls[["bh1"]]`). TAC stability clause matches
`02.2`. This notebook is off the default funder knit
(`Rscript scripts/run_pipeline.R --only shortcut`).

| Control | Env var | Default |
|---------|---------|---------|
| SRR(s) | `BM_CLOSED_SRR` | `bh3` (`all` → six) |
| \(\sigma\) (log SSB) | `BM_SHORTCUT_SIGMA` | `0.2` |
| \(\rho\) (lag-1) | `BM_SHORTCUT_RHO` | `0` (uncorrelated SSB) |
| Iterations | `BM_SHORTCUT_NITS` | `100` |

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- data_root()
here <- clean_root()

dir.create(file.path(here, "tex/figs"),                              recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "data/results"),                          recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(here, "Rmd/cache", "02.3_closedLoop_shortcut"), recursive=TRUE, showWarnings=FALSE)

if (!"knitr" %in% loadedNamespaces())
  suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(here, "Rmd/cache", "02.3_closedLoop_shortcut/"),
  fig.path  = file.path(here, "tex/figs", "02.3_closedLoop_shortcut-"),
  fig.width = 9,
  fig.height= 5,
  dev       ="png")

Fig=0
Tab=0
```

```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(ggplot2)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(FLasher)
  library(FLBacktest)
  library(doParallel)
  library(parallel)
  library(foreach)})

source(file.path(here, "R/constants.R"))
source(file.path(here, "R/om.R"))
source(file.path(here, "R/plot_traj.R"))
theme_set(theme_paper())
```

```{r load-om}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))

sids = as.character(stocks$sid)
om   = oms
if (!exists("eqs")) {
  if (!"bh1" %in% names(eqls))
    stop("eqs missing and eqls[['bh1']] not found — knit 01.0 first.",
         call. = FALSE)
  eqs = eqls[["bh1"]]
}
require_om_gate(root, sids)

srDev = lapply(eqls, function(eq) FLQuants(lapply(eq, srResiduals)))
startYr = setNames(rep(2015L, length(sids)), sids)
endYr   = setNames(sapply(om, function(x) dims(x)$maxyear), names(om))
bndTac  = hcr_tac_bounds()
bndWhen = hcr_tac_bnd_when()

srr6 = c("bh1", "bh2", "bh3", "rk1", "rk2", "rk3")
srr_env = Sys.getenv("BM_CLOSED_SRR", unset = srr_om)
if (identical(tolower(trimws(srr_env)), "all")) {
  srr_run = intersect(srr6, names(eqls))
} else {
  srr_run = intersect(trimws(strsplit(srr_env, ",", fixed = TRUE)[[1]]),
                      names(eqls))
}
if (!length(srr_run))
  stop("No SRR from BM_CLOSED_SRR=", srr_env, " found in eqls.", call. = FALSE)

sigma = as.numeric(Sys.getenv("BM_SHORTCUT_SIGMA", unset = "0.2"))
rho   = as.numeric(Sys.getenv("BM_SHORTCUT_RHO",   unset = "0"))
nits  = as.integer(Sys.getenv("BM_SHORTCUT_NITS",  unset = "100"))
if (!is.finite(sigma) || sigma < 0)
  stop("BM_SHORTCUT_SIGMA must be a finite non-negative number.", call. = FALSE)
if (!is.finite(nits) || nits < 1L)
  stop("BM_SHORTCUT_NITS must be a positive integer.", call. = FALSE)
```

# Short-cut `hcrICES`

EqSim Type 1 SSB error (`eqSimErr`) multiplies perceived numbers before the
HCR. Operating Model SRR(s): `r paste(srr_run, collapse = ", ")`.
\(\sigma = `r sigma`\), \(\rho = `r rho`\), `r nits` iterations.

```{r scenario-shortcut}
flr_pkgs = c("methods", "FLCore", "FLBRP", "icesdata", "FLBacktest", "FLasher")
ncores = max(1L, detectCores() - 4L)
libs = .libPaths()
cl = makeCluster(ncores)
registerDoParallel(cl)
tryCatch({
  clusterExport(cl, c("flr_pkgs", "libs", "om", "eqls", "eqs", "srDev",
                      "startYr", "endYr", "bndTac", "bndWhen",
                      "sigma", "rho", "nits"),
                envir = environment())
  clusterEvalQ(cl, {
    .libPaths(libs)
    suppressPackageStartupMessages(
      lapply(flr_pkgs, library, character.only = TRUE))
    NULL
  })
  scen = list()
  for (nm in srr_run) {
    clusterExport(cl, "nm", envir = environment())
    scen[[nm]] = FLStocks(foreach(id = sids,
                                  .packages = flr_pkgs,
                                  .export = c("om", "eqls", "eqs", "srDev",
                                              "startYr", "endYr", "nm",
                                              "bndTac", "bndWhen",
                                              "sigma", "rho", "nits")) %dopar% {
      stk = om[[id]]
      if (dims(stk)$iter < nits)
        stk = propagate(stk, nits)
      rec = srDev[[nm]][[id]]
      if (dims(rec)$iter < nits)
        rec = propagate(rec, nits)
      err = eqSimErr(sigma, rho = rho,
                     start = startYr[id], end = endYr[id],
                     nits = nits, seed = NULL)
      hcrICES(stk, eql = eqls[[nm]][[id]],
              sr_deviances = rec,
              params = hcrParams(eqs[[id]]),
              start = startYr[id], end = endYr[id],
              err = err, implErr = 0,
              bndTac = bndTac, bndWhen = bndWhen)[[1]]
    })
    names(scen[[nm]]) = sids
  }
}, finally = stopCluster(cl))

closedLoopShortcut = list(
  SRR = scen, bndTac = bndTac, bndWhen = bndWhen,
  srr_run = srr_run, sigma = sigma, rho = rho, nits = nits)

save(closedLoopShortcut,
     file = file.path(root, "data/results/02.3_closedLoop_shortcut.RData"))
```

# Trajectories

Historical vs Advice rule (EqSim short-cut; median over iterations). Each
metric is scaled by its Historical mean; facet by quantity \(\times\) stock
(shared relative \(y\)-scale).

```{r traj-plot, fig.height=9, fig.width=11}
labs = setNames(as.character(stocks$name), as.character(stocks$sid))
srr_plot = if (srr_om %in% names(closedLoopShortcut[["SRR"]])) srr_om else
  names(closedLoopShortcut[["SRR"]])[1]
print(plot_hcr_vs_hist(
  om, closedLoopShortcut[["SRR"]][[srr_plot]], eqs, eqls[[srr_plot]],
  sids = sids, labels = labs, start_ar = 2015L,
  advice_lab = "Advice rule"))
```

**Figure `r Fig = Fig + 1; Fig`**. Short-cut closed loop (`r srr_plot`,
median of `r nits` iterations) vs historical `FLStock`.

```{r traj-plot-srr, fig.height=9, fig.width=11, eval = length(closedLoopShortcut[["SRR"]]) > 1L}
print(plot_hcr_vs_hist(
  om, closedLoopShortcut[["SRR"]], eqs,
  eqls[names(closedLoopShortcut[["SRR"]])],
  sids = sids, labels = labs, start_ar = 2015L))
```

```{r traj-srr-caption, results='asis', eval = length(closedLoopShortcut[["SRR"]]) > 1L}
Fig <- Fig + 1
cat("\n\n**Figure ", Fig,
    "**. Short-cut Advice rule by SRR (medians).\n\n", sep = "")
```

Perfect-information counterpart: [`02.2_closedLoop.Rmd`](02.2_closedLoop.Rmd).
SAM OEM: [`02.4.1_sam.Rmd`](02.4.1_sam.Rmd) / [`02.4.2_closedLoop_sam.Rmd`](02.4.2_closedLoop_sam.Rmd).
