---
title: "SAM closed loop: observation-error iterations"
subtitle: "hcrICES with SAM, scn0–scn2, 100 iterations, survey CV 30%"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

The SAM closed loop of [`02.4.2_closedLoop_sam.Rmd`](02.4.2_closedLoop_sam.Rmd)
repeated over many survey realisations, so the Advice-rule outcome carries
observation error. Check [`02.4.1_sam.Rmd`](02.4.1_sam.Rmd) first.

| Setting | Value | Override |
|---------|-------|----------|
| Stocks | SAM stocks + Irish Sea cod (`sam_sids`) | — |
| Operating Model SRRs | `scn0`, `scn1`, `scn2` | — |
| Iterations | 100 | `BM_SAM_MC_NITS` |
| Survey CV (`indexCv`) | 0.3 | `BM_SAM_MC_CV` |
| Workers | cores − 2 | `BM_SAM_MC_CORES` |

`scn0`: near-linear BH (\(h = 0.999\)); `scn1`: unconstrained BH (as `bh1`);
`scn2`: BH with steepness prior \(h = 0.7\) (see `01.0_condition_om.Rmd`).

`FLR_SAM` fits one iteration at a time, so each (SRR, stock, iteration) is a
separate single-iteration `hcrICES` run, from 2015, `implErr = 0`, the TAC
clause of `02.2`, and recruitment residuals replayed from each SRR (the same
in every iteration). Iterations differ only in the survey noise. The survey
seed depends on stock and iteration, not SRR, so the three SRRs see the same
surveys. Catch-at-age placeholders are passed to SAM as `NA` as in `02.4.1`.

Runs are spread over a PSOCK cluster. Workers load the Operating Model from
disk and build their own surveys; nothing but job ids is shipped. Each run is
cached under `data/interim/02.4.3_closedLoop_sam_mc/`, so a re-knit resumes.
Failed runs are listed and stop the knit after all others finish.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- data_root()
here <- clean_root()

dir.create(file.path(here, "tex/figs"),                               recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "data/results"),                           recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(here, "Rmd/cache", "02.4.3_closedLoop_sam_mc"),  recursive=TRUE, showWarnings=FALSE)

# knitr is already loaded by rmarkdown::render; do not library() a second copy.
if (!"knitr" %in% loadedNamespaces())
  suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(here, "Rmd/cache", "02.4.3_closedLoop_sam_mc/"),
  fig.path  = file.path(here, "tex/figs", "02.4.3_closedLoop_sam_mc-"),
  fig.width = 9,
  fig.height= 5,
  dev       ="png")

Fig=0
Tab=0

nits <- as.integer(Sys.getenv("BM_SAM_MC_NITS", unset = "100"))
indexCv <- as.numeric(Sys.getenv("BM_SAM_MC_CV", unset = "0.3"))
ncores <- as.integer(Sys.getenv("BM_SAM_MC_CORES",
                                unset = max(1L, parallel::detectCores() - 2L)))
if (!is.finite(nits) || nits < 1L) stop("BM_SAM_MC_NITS must be >= 1.", call. = FALSE)
if (!is.finite(indexCv) || indexCv <= 0) stop("BM_SAM_MC_CV must be > 0.", call. = FALSE)
if (!is.finite(ncores) || ncores < 1L) stop("BM_SAM_MC_CORES must be >= 1.", call. = FALSE)
```

```{r libs}
source(file.path(here, "R/sam_oem.R"))
source(file.path(here, "R/sam_mc.R"))
sam_libs()
suppressPackageStartupMessages({
  library(FLCore)
  library(ggplot2)
  library(FLBRP)
  library(icesdata)
  library(FLasher)
  library(FLBacktest)
  library(FLfse)
  library(stockassessment)
  library(parallel)
})
source(file.path(here, "R/constants.R"))
source(file.path(here, "R/om.R"))
source(file.path(here, "R/plot_traj.R"))
theme_set(theme_paper())
```

```{r load-om}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))
if (!exists("eqs"))
  eqs <- eqls[["bh1"]]

srr_run <- c("scn0", "scn1", "scn2")
miss <- setdiff(srr_run, names(eqls))
if (length(miss))
  stop("eqls is missing ", paste(miss, collapse = ", "),
       " — knit 01.0_condition_om.Rmd first.", call. = FALSE)

sids <- sam_sids(oms, stocks)
require_om_gate(root, sids)
om <- oms[sids]
for (s in srr_run) {
  miss <- setdiff(sids, names(eqls[[s]]))
  if (length(miss))
    stop("eqls$", s, " is missing ", paste(miss, collapse = ", "), call. = FALSE)
}

startYr <- 2015L
too_short <- names(om)[sapply(om, function(x) dims(x)$maxyear) < startYr]
if (length(too_short))
  stop("Terminal year is before 2015 for: ",
       paste(too_short, collapse = ", "), call. = FALSE)

out <- file.path(root, "data/interim/02.4.3_closedLoop_sam_mc",
                 sprintf("cv%03d", round(100 * indexCv)))
dir.create(out, recursive = TRUE, showWarnings = FALSE)
```

SAM stocks: `r paste(sids, collapse = ", ")`. SRRs:
`r paste(srr_run, collapse = ", ")`. Iterations: **`r nits`**. Survey CV:
**`r indexCv`**. Workers: **`r ncores`**. Jobs:
**`r length(srr_run) * length(sids) * nits`**.

# Closed loop

```{r scenario-hcr}
jobs <- expand.grid(iter = seq_len(nits), id = sids, srr = srr_run,
                    stringsAsFactors = FALSE)
jobs <- jobs[, c("srr", "id", "iter")]
todo <- !file.exists(mapply(sam_mc_file, out, jobs$srr, jobs$id, jobs$iter))
message(sum(todo), " of ", nrow(jobs), " runs to do; ",
        sum(!todo), " cached in ", out)

bndTac <- hcr_tac_bounds()
bndWhen <- hcr_tac_bnd_when()
# Global environment so the closure does not carry the knit env to workers.
worker <- function(k, jobs, sids, indexCv, start, bndTac, bndWhen, out)
  sam_mc_job(jobs$srr[k], jobs$id[k], jobs$iter[k], sids = sids,
             indexCv = indexCv, start = start,
             bndTac = bndTac, bndWhen = bndWhen, out = out)
environment(worker) <- globalenv()
t0 <- Sys.time()
if (any(todo)) {
  cl <- makeCluster(min(ncores, sum(todo)))
  res <- tryCatch({
    clusterCall(cl, sam_mc_init, here = here, root = root, libs = .libPaths())
    parLapplyLB(cl, which(todo), worker,
      jobs = jobs, sids = sids, indexCv = indexCv, start = startYr,
      bndTac = bndTac, bndWhen = bndWhen, out = out)
  }, finally = stopCluster(cl))
  fail <- Filter(function(r) !isTRUE(r$ok), res)
  if (length(fail)) {
    msg <- vapply(fail, function(r)
      sprintf("%s / %s / iter %d: %s", r$srr, r$id, as.integer(r$iter), r$msg),
      character(1))
    stop(length(fail), " of ", length(res), " SAM closed-loop runs failed ",
         "(successful runs are cached; re-knit to retry):\n",
         paste(head(msg, 20), collapse = "\n"), call. = FALSE)
  }
}
runTime <- difftime(Sys.time(), t0, units = "mins")

collected <- lapply(setNames(srr_run, srr_run), function(s)
  setNames(lapply(sids, function(id)
    sam_mc_collect(out, s, id, seq_len(nits))), sids))
retried <- do.call(rbind, lapply(srr_run, function(s)
  do.call(rbind, lapply(sids, function(id) {
    a <- attr(collected[[s]][[id]], "seed_attempt")
    k <- which(a > 1L)
    if (!length(k)) return(NULL)
    data.frame(SRR = s, Stock = id, Iter = as.integer(names(a)[k]),
               Attempt = a[k], stringsAsFactors = FALSE)
  }))))
closedLoopSamMC <- lapply(collected, FLStocks)
rm(collected)

save(closedLoopSamMC, retried, sids, srr_run, nits, indexCv, startYr,
     bndTac, bndWhen,
     file = file.path(root, "data/results/02.4.3_closedLoop_sam_mc.RData"))
```

Run time this knit: `r round(as.numeric(runTime), 1)` min.

A run whose SAM fit fails is retried with a new survey seed (up to three
attempts). Those runs no longer share the survey with the other SRRs.

**Table `r Tab = Tab + 1; Tab`**. Runs that needed a new survey seed.

```{r retried}
if (is.null(retried)) {
  cat("None.\n")
} else knitr::kable(retried, row.names = FALSE)
```

# Trajectories

Same layout as `02.2` / `02.3`: metrics relative to the Historical mean,
facet quantity \(\times\) stock. Line = median over iterations; band = 5–95%.

```{r traj-ribbon, fig.height=9, fig.width=11}
labs <- setNames(as.character(stocks$name[match(sids, stocks$sid)]), sids)
print(plot_hcr_ribbon(om, closedLoopSamMC, sids = sids, labels = labs,
                      start_ar = startYr))
```

**Figure `r Fig = Fig + 1; Fig`**. SAM closed loop over `r nits` survey
realisations (CV `r indexCv`), one colour per Operating Model SRR, vs
historical `FLStock`.

```{r traj-median, fig.height=9, fig.width=11}
print(plot_hcr_vs_hist(om, closedLoopSamMC, eqs, eqls[srr_run],
                       sids = sids, labels = labs, start_ar = startYr))
```

**Figure `r Fig = Fig + 1; Fig`**. Medians with ICES and Operating Model
(`scn0`) reference points.

# Terminal year

**Table `r Tab = Tab + 1; Tab`**. Terminal year: median (5–95%) of SSB /
\(B_{\mathrm{MSY}}\) and \(F\) / \(F_{\mathrm{MSY}}\) for the Operating Model
SRR, and probability SSB < ICES \(B_{\lim}\).

```{r terminal-table}
q3 <- function(x) {
  p <- stats::quantile(x, c(0.05, 0.5, 0.95), na.rm = TRUE)
  sprintf("%.2f (%.2f–%.2f)", p[2], p[1], p[3])
}
term <- do.call(rbind, lapply(srr_run, function(s) {
  do.call(rbind, lapply(sids, function(id) {
    stk <- closedLoopSamMC[[s]][[id]]
    y <- ac(dims(stk)$maxyear)
    rp <- refpts(eqls[[s]][[id]])
    rpi <- refpts(eqs[[id]])
    b <- c(ssb(stk)[, y])
    f <- c(fbar(stk)[, y])
    data.frame(
      Stock = labs[[id]], SRR = s,
      Year = y,
      `SSB/Bmsy` = q3(b / c(rp["msy", "ssb"])),
      `F/Fmsy` = q3(f / c(rp["msy", "harvest"])),
      `P(SSB<Blim)` = round(mean(b < c(rpi["Blim", "ssb"])), 2),
      check.names = FALSE, stringsAsFactors = FALSE)
  }))
}))
knitr::kable(term, row.names = FALSE)
```
