---
title: "SAM closed loop"
subtitle: "OEM fit, then hcrICES with a SAM assessment"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

SAM observation model for the closed-loop ICES HCR. This is not the funder
draft (`02.2` is perfect information; `02.3` is the EqSim short-cut OEM).
Three stages, selected by `BM_SAM_STAGE` (`fit`, `loop`, `srr`; default `fit`):

1. **fit** — build an age-structured survey from Operating Model numbers and
   fit SAM once per SAM stock. The knit stops if a fit fails.
2. **loop** — same OEM inside `hcrICES` (`err` is an `FLIndex`) for **bh3**
   only, from 2015, `implErr = 0`.
3. **srr** — the same closed loop for `bh1`, `bh2`, `bh3`, `rk1`, `rk2`, `rk3`.

Irish Sea cod and NEA mackerel are not SAM assessments and are not included.
HCR parameters stay on `eqs` (`eqls[["bh1"]]`). The TAC stability clause
matches `02.2`.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- data_root()
here <- clean_root()

dir.create(file.path(here, "tex/figs"),                         recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "data/results"),                     recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(here, "Rmd/cache", "02.4_closedLoop_sam"), recursive=TRUE, showWarnings=FALSE)

# knitr is already loaded by rmarkdown::render; do not library() a second copy.
if (!"knitr" %in% loadedNamespaces())
  suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(here, "Rmd/cache", "02.4_closedLoop_sam/"),
  fig.path  = file.path(here, "tex/figs", "02.4_closedLoop_sam-"),
  fig.width = 9,
  fig.height= 5,
  dev       ="png")

Fig=0
Tab=0

stage <- Sys.getenv("BM_SAM_STAGE", unset = "fit")
stage <- match.arg(stage, c("fit", "loop", "srr"))
```

```{r libs}
# renv sandbox does not see the user win-library (FLfse, stockassessment, FLR).
# Append only — prepending shadows renv tidyverse binaries (breaks ggplot2/vctrs).
user_lib <- file.path(Sys.getenv("LOCALAPPDATA", unset = ""),
                      "R", "win-library",
                      paste(R.version$major,
                            strsplit(R.version$minor, ".", fixed = TRUE)[[1]][1],
                            sep = "."))
if (!nzchar(user_lib) || !dir.exists(user_lib))
  user_lib <- "C:/Users/lauri/AppData/Local/R/win-library/4.4"
if (dir.exists(user_lib) && !user_lib %in% .libPaths())
  .libPaths(c(.libPaths(), user_lib))

if (!requireNamespace("FLfse", quietly = TRUE) ||
    !requireNamespace("stockassessment", quietly = TRUE))
  stop("SAM OEM needs installed packages FLfse and stockassessment ",
       "(user library: ", user_lib, ").", call. = FALSE)
suppressPackageStartupMessages({
  library(FLCore)
  library(ggplot2)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(reshape2)
  library(FLasher)
  library(FLBacktest)
  library(FLfse)
  library(stockassessment)
})
source(file.path(here, "R/constants.R"))
source(file.path(here, "R/om.R"))
options(FLfse.sam.q = 1)
theme_set(theme_paper())
```

```{r load-om}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))

srr6 <- c("bh1", "bh2", "bh3", "rk1", "rk2", "rk3")
miss <- setdiff(srr6, names(eqls))
if (length(miss))
  stop("eqls is missing ", paste(miss, collapse = ", "),
       " — knit 01.0_condition_om.Rmd first.", call. = FALSE)
if (!exists("eqs"))
  eqs <- eqls[["bh1"]]

sam_csv <- as.character(stocks$sid[grepl("SAM", stocks$assessment)])
sids <- sam_csv[sam_csv %in% names(oms)]
if (!length(sids))
  stop("No SAM stocks in oms / stocks.csv.", call. = FALSE)
require_om_gate(root, sids)

om <- oms[sids]
nits <- 1L
indexCv <- 0.1
startYr <- setNames(rep(2015L, length(sids)), sids)
endYr <- setNames(sapply(om, function(x) dims(x)$maxyear), names(om))
too_short <- names(endYr)[endYr < startYr]
if (length(too_short))
  stop("Terminal year is before 2015 for: ",
       paste(too_short, collapse = ", "), call. = FALSE)

srr_run <- switch(stage,
  fit  = character(),
  loop = "bh3",
  srr  = srr6)
```

Stage: **`r stage`**. SAM stocks: `r paste(sids, collapse = ", ")`.

# Observation model

Ages exclude recruitment and the plus group. The stored index is a
catchability of 1 times lognormal noise (`indexCv`). SAM is started at that
catchability (`options(FLfse.sam.q = 1)`). `hcrICES` multiplies the stored
index by Operating Model `stock.n` each advice year. The one-shot fit below
applies that product once, over the whole assessment, and calls
`FLfse::FLR_SAM`.

```{r oem}
sam_oem <- function(stk, indexCv = 0.1, nits = 1L) {
  if (nits > 1L && dims(stk)$iter == 1L)
    stk <- propagate(stk, nits)
  stock.n(stk)[is.na(stock.n(stk))] <- 1e3
  stock.n(stk) <- qmax(stock.n(stk), 1e-2)
  rng <- range(stk)
  amin <- unname(rng["min"]) + 1
  pg <- unname(rng["plusgroup"])
  if (!is.finite(pg))
    pg <- unname(rng["max"])
  amax <- pg - 1
  if (!is.finite(amin) || !is.finite(amax) || amax < amin)
    stop("SAM survey ages are empty.", call. = FALSE)
  idx <- samIndex(trim(stk, age = ac(amin:amax)))
  index(idx)[] <- 1
  index(idx) <- index(idx) %*%
    rlnorm(nits, index(idx) %=% 0, indexCv) %*%
    rlnorm(nits, index(idx)[1] %=% 0, indexCv)
  idx
}

sam_observe <- function(stk, idx) {
  ages <- dimnames(index(idx))$age
  yrs <- dimnames(index(idx))$year
  obs <- stock.n(stk)[ages, yrs]
  obs[!is.finite(obs) | obs <= 0] <- 1e-2
  index(idx) <- obs %*% index(idx)
  idx
}

# Years before the first advice year stay as survey observations.
# hcrICES multiplies the current year by Operating Model numbers, so those
# years must still be the catchability multiplier, not numbers.
sam_prime <- function(stk, idx, start, lag = 1L) {
  yrs <- as.integer(dimnames(index(idx))$year)
  hist <- yrs < (as.integer(start) - as.integer(lag))
  if (!any(hist))
    return(idx)
  ages <- dimnames(index(idx))$age
  y <- ac(yrs[hist])
  obs <- stock.n(stk)[ages, y]
  obs[!is.finite(obs) | obs <= 0] <- 1e-2
  index(idx)[, y] <- obs %*% index(idx)[, y]
  idx
}
```

# SAM fit to the Operating Model

```{r sam-fit}
set.seed(2404L)
oem <- FLIndices(lapply(om, sam_oem, indexCv = indexCv, nits = nits))
names(oem) <- sids

fit_one <- function(id) {
  stk <- om[[id]]
  cn <- catch.n(stk)
  yrs_all <- dimnames(cn)$year
  ok <- vapply(yrs_all, function(y) {
    x <- c(cn[, y])
    any(is.finite(x) & x > 0)
  }, logical(1))
  if (!any(ok))
    stop("No catch-at-age for ", id, call. = FALSE)
  yrs <- dimnames(cn)$year[ok]
  stk <- window(stk, start = as.numeric(yrs[1]), end = as.numeric(yrs[length(yrs)]))
  idx <- sam_observe(stk, oem[[id]])
  idx <- window(idx, start = dims(stk)$minyear, end = dims(stk)$maxyear)
  fit <- FLfse::FLR_SAM(stk, idx, NA_rm = TRUE)
  if (inherits(fit, "error") || inherits(fit, "try-error"))
    stop("FLR_SAM failed for ", id, call. = FALSE)
  hat <- FLfse::SAM2FLStock(fit)
  if (!is(hat, "FLStock"))
    stop("SAM2FLStock did not return an FLStock for ", id, call. = FALSE)
  both <- as.character(intersect(dimnames(ssb(stk))$year, dimnames(ssb(hat))$year))
  if (length(both) < 5L)
    stop("SAM fit for ", id, " overlaps the Operating Model in fewer than 5 years.",
         call. = FALSE)
  tail_yrs <- tail(both, 10)
  ssb_r <- median(c(ssb(hat)[, tail_yrs] / ssb(stk)[, tail_yrs]), na.rm = TRUE)
  f_r <- median(c(fbar(hat)[, tail_yrs] / fbar(stk)[, tail_yrs]), na.rm = TRUE)
  conv <- if (is.null(fit$opt)) NA_integer_ else unname(fit$opt$convergence)
  if (!is.finite(ssb_r) || !is.finite(f_r) ||
      ssb_r < 0.5 || ssb_r > 2 || f_r < 0.5 || f_r > 2)
    stop("SAM fit for ", id, " is off the Operating Model scale (SSB ",
         signif(ssb_r, 3), ", F ", signif(f_r, 3),
         "). convergence=", conv, ".", call. = FALSE)
  list(stock = hat, ratio = c(ssb = ssb_r, f = f_r, conv = conv))
}

fits <- lapply(sids, function(id) {
  message("SAM fit: ", id)
  tryCatch(
    fit_one(id),
    error = function(e)
      stop("SAM OEM fit failed for ", id, ": ", conditionMessage(e),
           call. = FALSE))
})
names(fits) <- sids
samFit <- FLStocks(lapply(fits, `[[`, "stock"))
samRatio <- do.call(rbind, lapply(sids, function(id) {
  data.frame(sid = id,
             ssb = unname(fits[[id]]$ratio["ssb"]),
             f = unname(fits[[id]]$ratio["f"]),
             conv = unname(fits[[id]]$ratio["conv"]),
             stringsAsFactors = FALSE)
}))
save(oem, samFit, samRatio, sids, indexCv, nits, stage,
     file = file.path(root, "data/results/02.4_sam_oem.RData"))
```

**Table `r Tab=Tab+1; Tab`**. Median SAM / Operating Model over the last 10
overlapping years. The fit stops if either ratio is outside 0.5–2.
Convergence 0 is a completed `sam.fit`.

```{r sam-fit-table}
knitr::kable(
  cbind(name = as.character(stocks$name[match(samRatio$sid, stocks$sid)]),
        samRatio),
  digits = 3, row.names = FALSE,
  col.names = c("Stock", "SID", "SSB ratio", "F ratio", "Convergence"))
```

```{r sam-fit-plot, fig.height=8, fig.width=10}
cmp <- lapply(sids, function(id) {
  stk <- om[[id]]
  hat <- samFit[[id]]
  yrs <- as.character(intersect(dimnames(ssb(stk))$year, dimnames(ssb(hat))$year))
  data.frame(
    sid = id,
    name = as.character(stocks$name[match(id, stocks$sid)]),
    year = as.integer(yrs),
    SSB = c(ssb(hat)[, yrs]),
    F = c(fbar(hat)[, yrs]),
    Catch = c(catch(hat)[, yrs]),
    stringsAsFactors = FALSE)
})
cmp <- do.call(rbind, cmp)
cmp_om <- lapply(sids, function(id) {
  stk <- om[[id]]
  hat <- samFit[[id]]
  yrs <- as.character(intersect(dimnames(ssb(stk))$year, dimnames(ssb(hat))$year))
  data.frame(
    sid = id,
    name = as.character(stocks$name[match(id, stocks$sid)]),
    year = as.integer(yrs),
    SSB = c(ssb(stk)[, yrs]),
    F = c(fbar(stk)[, yrs]),
    Catch = c(catch(stk)[, yrs]),
    stringsAsFactors = FALSE)
})
cmp_om <- do.call(rbind, cmp_om)
cmp$series <- "SAM"
cmp_om$series <- "Operating model"
d <- rbind(cmp_om, cmp)
d <- reshape2::melt(d, id.vars = c("sid", "name", "year", "series"),
                    variable.name = "metric", value.name = "value")
ggplot(d, aes(year, value, colour = series)) +
  geom_line() +
  facet_grid(name ~ metric, scales = "free_y") +
  scale_colour_manual(values = c("Operating model" = paper_col("Historical"),
                                 "SAM" = paper_col("advice"))) +
  labs(x = NULL, y = NULL, colour = NULL) +
  theme_paper()
```

**Figure `r Fig=Fig+1; Fig`**. One-shot SAM fit (colour) against the
Operating Model (black). The survey is Operating Model numbers times
lognormal noise.

# Closed loop

Runs only when `BM_SAM_STAGE` is `loop` (bh3) or `srr` (six SRRs). Each
advice year refits SAM on the OEM index.

```{r scenario-hcr, eval = stage %in% c("loop", "srr")}
srDev <- lapply(eqls[srr_run], function(eq) FLQuants(lapply(eq[sids], srResiduals)))
bndTac <- hcr_tac_bounds()
bndWhen <- hcr_tac_bnd_when()
# sam.fit is run here, not on a cluster: shipping FLIndex to a worker
# has produced non-finite optim failures that do not occur in-process.
closedLoopSam <- list()
for (nm in srr_run) {
  runs <- setNames(vector("list", length(sids)), sids)
  for (id in sids) {
    message("SAM loop ", nm, " ", id)
    set.seed(1000L + match(id, sids))
    stk <- om[[id]]
    if (nits > 1L && dims(stk)$iter == 1L)
      stk <- propagate(stk, nits)
    idx <- sam_prime(stk, oem[[id]], start = startYr[id], lag = 1L)
    runs[[id]] <- tryCatch(
      hcrICES(stk, eql = eqls[[nm]][[id]],
              sr_deviances = srDev[[nm]][[id]],
              params = hcrParams(eqs[[id]]),
              start = startYr[id], end = endYr[id],
              lag = 1, interval = 1,
              err = idx, implErr = 0,
              bndTac = bndTac, bndWhen = bndWhen,
              verbose = TRUE)[[1]],
      error = function(e)
        stop(nm, " / ", id, ": ", conditionMessage(e), call. = FALSE))
  }
  closedLoopSam[[nm]] <- FLStocks(runs)
}

save(closedLoopSam, sids, srr_run, nits, indexCv, startYr, endYr, stage,
     file = file.path(root, "data/results/02.4_closedLoop_sam.RData"))
```

```{r sam-loop-plot, fig.height=8, fig.width=10, eval = stage %in% c("loop", "srr")}
rows <- lapply(names(closedLoopSam), function(nm) {
  lapply(sids, function(id) {
    hat <- closedLoopSam[[nm]][[id]]
    stk <- om[[id]]
    yrs <- as.character(intersect(dimnames(ssb(stk))$year, dimnames(ssb(hat))$year))
    yrs <- yrs[as.integer(yrs) >= startYr[id]]
    data.frame(
      sid = id,
      name = as.character(stocks$name[match(id, stocks$sid)]),
      srr = nm,
      year = as.integer(yrs),
      SSB = c(ssb(hat)[, yrs]),
      F = c(fbar(hat)[, yrs]),
      Catch = c(catch(hat)[, yrs]),
      stringsAsFactors = FALSE)
  })
})
d <- do.call(rbind, unlist(rows, recursive = FALSE))
d <- reshape2::melt(d, id.vars = c("sid", "name", "srr", "year"),
                    variable.name = "metric", value.name = "value")
ggplot(d, aes(year, value, colour = srr)) +
  geom_line() +
  facet_grid(name ~ metric, scales = "free_y") +
  labs(x = NULL, y = NULL, colour = NULL) +
  theme_paper()
```

```{r sam-loop-caption, results="asis", eval=stage %in% c("loop", "srr")}
Fig <- Fig + 1
cat("\n\n**Figure ", Fig,
    "**. SAM closed loop from 2015. One line per SRR branch.\n\n", sep = "")
```
