---
title: "SAM fit to the Operating Model"
subtitle: "Observation error model check before the SAM closed loop"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

Builds the age-structured survey from Operating Model numbers and fits SAM
once per SAM stock, then compares the fit with the Operating Model. Check
this notebook before running the HCR in
[`02.4.2_closedLoop_sam.Rmd`](02.4.2_closedLoop_sam.Rmd), which reuses the
survey index saved here. The knit stops if a fit fails or is off the
Operating Model scale.

Stocks are the SAM assessments in `stocks.csv` plus `sam_extra` (Irish Sea
cod), whose ICES assessment is not SAM but which uses the same management
procedure so all four case studies are comparable. NEA mackerel is not
included.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- data_root()
here <- clean_root()

dir.create(file.path(here, "tex/figs"),                 recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "data/results"),             recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(here, "Rmd/cache", "02.4.1_sam"),  recursive=TRUE, showWarnings=FALSE)

# knitr is already loaded by rmarkdown::render; do not library() a second copy.
if (!"knitr" %in% loadedNamespaces())
  suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(here, "Rmd/cache", "02.4.1_sam/"),
  fig.path  = file.path(here, "tex/figs", "02.4.1_sam-"),
  fig.width = 9,
  fig.height= 5,
  dev       ="png")

Fig=0
Tab=0
```

```{r libs}
source(file.path(here, "R/sam_oem.R"))
sam_libs()
suppressPackageStartupMessages({
  library(FLCore)
  library(ggplot2)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(reshape2)
  library(FLasher)
  library(FLBacktest)
  library(FLfse)
  library(stockassessment)
})
source(file.path(here, "R/constants.R"))
source(file.path(here, "R/om.R"))
source(file.path(here, "R/plot_traj.R"))
options(FLfse.sam.q = 1)
theme_set(theme_paper())
```

```{r load-om}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))
if (!exists("eqs"))
  eqs <- eqls[["bh1"]]

sids <- sam_sids(oms, stocks)
require_om_gate(root, sids)

om <- oms[sids]
nits <- 1L
indexCv <- 0.1
```

SAM stocks: `r paste(sids, collapse = ", ")`.

# Observation model

Ages exclude recruitment and the plus group. The stored index is a
catchability of 1 times lognormal noise (`indexCv` = `r indexCv`). SAM is
started at that catchability (`options(FLfse.sam.q = 1)`). In the closed loop
`hcrICES` multiplies the stored index by Operating Model `stock.n` each
advice year; the one-shot fit below applies that product once, over the
whole assessment, and calls `FLfse::FLR_SAM`. Helpers are in
`R/sam_oem.R`.

Catch-at-age at or below `r sam_catch_tol` is a placeholder for a missing
age (e.g. Celtic Sea cod ages 1–2 before 2004) and is passed to SAM as `NA`.
Read as a real catch, SAM estimates \(F \approx 0\) at those ages and biases
Fbar and SSB. Ages with no catch in any year (haddock age 0, where the
Operating Model F is 0) keep the placeholder: near-zero catch is correct
there, and the forecast inside `hcrICES` needs finite catches. The Operating
Model is not changed.

**Table `r Tab=Tab+1; Tab`**. Catch-at-age cells passed to SAM as missing.

```{r catch-na-table}
naTab <- do.call(rbind, lapply(sids, function(id) {
  cn <- catch.n(om[[id]])
  bad <- sam_catch_bad(om[[id]])
  if (!any(bad))
    return(data.frame(sid = id, ages = "", years = "", cells = 0L))
  ages <- dimnames(cn)$age[apply(bad, 1, any)]
  yrs <- as.integer(dimnames(cn)$year[apply(bad, 2, any)])
  data.frame(sid = id, ages = paste(ages, collapse = ", "),
             years = paste(range(yrs), collapse = "–"), cells = sum(bad))
}))
knitr::kable(naTab, row.names = FALSE,
             col.names = c("SID", "Ages", "Years", "Cells"))
```

# SAM fit to the Operating Model

```{r sam-fit}
set.seed(2404L)
oem <- FLIndices(lapply(om, sam_oem, indexCv = indexCv, nits = nits))
names(oem) <- sids

fit_one <- function(id) {
  stk <- om[[id]]
  cn <- catch.n(stk)
  yrs_all <- dimnames(cn)$year
  ok <- vapply(yrs_all, function(y) {
    x <- c(cn[, y])
    any(is.finite(x) & x > sam_catch_tol)
  }, logical(1))
  if (!any(ok))
    stop("No catch-at-age for ", id, call. = FALSE)
  yrs <- dimnames(cn)$year[ok]
  stk <- window(stk, start = as.numeric(yrs[1]), end = as.numeric(yrs[length(yrs)]))
  idx <- sam_observe(stk, oem[[id]])
  idx <- window(idx, start = dims(stk)$minyear, end = dims(stk)$maxyear)
  fit <- FLfse::FLR_SAM(sam_catch_na(stk), idx, NA_rm = TRUE)
  if (inherits(fit, "error") || inherits(fit, "try-error"))
    stop("FLR_SAM failed for ", id, call. = FALSE)
  hat <- FLfse::SAM2FLStock(fit)
  if (!is(hat, "FLStock"))
    stop("SAM2FLStock did not return an FLStock for ", id, call. = FALSE)
  both <- as.character(intersect(dimnames(ssb(stk))$year, dimnames(ssb(hat))$year))
  if (length(both) < 5L)
    stop("SAM fit for ", id, " overlaps the Operating Model in fewer than 5 years.",
         call. = FALSE)
  tail_yrs <- tail(both, 10)
  ssb_r <- median(c(ssb(hat)[, tail_yrs] / ssb(stk)[, tail_yrs]), na.rm = TRUE)
  f_r <- median(c(fbar(hat)[, tail_yrs] / fbar(stk)[, tail_yrs]), na.rm = TRUE)
  conv <- if (is.null(fit$opt)) NA_integer_ else unname(fit$opt$convergence)
  if (!is.finite(ssb_r) || !is.finite(f_r) ||
      ssb_r < 0.5 || ssb_r > 2 || f_r < 0.5 || f_r > 2)
    stop("SAM fit for ", id, " is off the Operating Model scale (SSB ",
         signif(ssb_r, 3), ", F ", signif(f_r, 3),
         "). convergence=", conv, ".", call. = FALSE)
  list(stock = hat, ratio = c(ssb = ssb_r, f = f_r, conv = conv))
}

fits <- lapply(sids, function(id) {
  message("SAM fit: ", id)
  tryCatch(
    fit_one(id),
    error = function(e)
      stop("SAM OEM fit failed for ", id, ": ", conditionMessage(e),
           call. = FALSE))
})
names(fits) <- sids
samFit <- FLStocks(lapply(fits, `[[`, "stock"))
samRatio <- do.call(rbind, lapply(sids, function(id) {
  data.frame(sid = id,
             ssb = unname(fits[[id]]$ratio["ssb"]),
             f = unname(fits[[id]]$ratio["f"]),
             conv = unname(fits[[id]]$ratio["conv"]),
             stringsAsFactors = FALSE)
}))
save(oem, samFit, samRatio, sids, indexCv, nits,
     file = file.path(root, "data/results/02.4.1_sam.RData"))
```

**Table `r Tab=Tab+1; Tab`**. Median SAM / Operating Model over the last 10
overlapping years. The fit stops if either ratio is outside 0.5–2.
Convergence 0 is a completed `sam.fit`.

```{r sam-fit-table}
knitr::kable(
  cbind(name = as.character(stocks$name[match(samRatio$sid, stocks$sid)]),
        samRatio),
  digits = 3, row.names = FALSE,
  col.names = c("Stock", "SID", "SSB ratio", "F ratio", "Convergence"))
```

# Trajectories

Same layout as [`02.2_closedLoop.Rmd`](02.2_closedLoop.Rmd): \(F\), SSB, Rec,
Catch relative to the Historical mean; facet quantity \(\times\) stock;
ICES / OM reference points; colour = Historical vs one-shot SAM fit.

```{r traj-fit-plot, fig.height=9, fig.width=11}
labs = setNames(as.character(stocks$name[match(sids, stocks$sid)]), sids)
eql_ref = if (srr_om %in% names(eqls)) eqls[[srr_om]] else eqls[["bh3"]]
print(plot_hcr_vs_hist(
  om, samFit, eqs, eql_ref,
  sids = sids, labels = labs, start_ar = 2015L,
  advice_lab = "SAM fit"))
```

**Figure `r Fig = Fig + 1; Fig`**. One-shot SAM fit vs historical Operating
Model `FLStock` (survey = OM numbers \(\times\) lognormal noise).
