---
title: "SAM closed loop"
subtitle: "hcrICES with a SAM assessment"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

Closed-loop ICES HCR with SAM as the assessment. This is not the funder
draft (`02.2` is perfect information; `02.3` is the EqSim short-cut OEM).
The survey index and SAM stocks come from
[`02.4.1_sam.Rmd`](02.4.1_sam.Rmd); knit and check that first.

The OEM enters `hcrICES` as `err` (an `FLIndex`); each advice year refits SAM
on the index, from 2015, `implErr = 0`. SRRs are selected by
`BM_SAM_STAGE`:

1. **loop** (default) — reference OM only (`srr_om`, **bhw**).
2. **srr** — `bh1`, `bh2`, `bh3`, `rk1`, `rk2`, `rk3`.

HCR parameters stay on `eqs` (`eqls[["bh1"]]`). The TAC stability clause
matches `02.2`. As in `02.4.1`, catch-at-age placeholders (at or below
`sam_catch_tol`) are passed to SAM as `NA`; the Operating Model's historical
values are restored in the saved runs.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- data_root()
here <- clean_root()

dir.create(file.path(here, "tex/figs"),                           recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(root, "data/results"),                       recursive=TRUE, showWarnings=FALSE)
dir.create(file.path(here, "Rmd/cache", "02.4.2_closedLoop_sam"), recursive=TRUE, showWarnings=FALSE)

# knitr is already loaded by rmarkdown::render; do not library() a second copy.
if (!"knitr" %in% loadedNamespaces())
  suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  cache.path= file.path(here, "Rmd/cache", "02.4.2_closedLoop_sam/"),
  fig.path  = file.path(here, "tex/figs", "02.4.2_closedLoop_sam-"),
  fig.width = 9,
  fig.height= 5,
  dev       ="png")

Fig=0
Tab=0

stage <- Sys.getenv("BM_SAM_STAGE", unset = "loop")
stage <- match.arg(stage, c("loop", "srr"))
```

```{r libs}
source(file.path(here, "R/sam_oem.R"))
sam_libs()
suppressPackageStartupMessages({
  library(FLCore)
  library(ggplot2)
  library(FLBRP)
  library(icesdata)
  library(plyr)
  library(reshape2)
  library(FLasher)
  library(FLBacktest)
  library(FLfse)
  library(stockassessment)
})
source(file.path(here, "R/constants.R"))
source(file.path(here, "R/om.R"))
source(file.path(here, "R/plot_traj.R"))
options(FLfse.sam.q = 1)
theme_set(theme_paper())
```

```{r load-om}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))

srr6 <- c("bh1", "bh2", "bh3", "rk1", "rk2", "rk3")
miss <- setdiff(srr6, names(eqls))
if (length(miss))
  stop("eqls is missing ", paste(miss, collapse = ", "),
       " — knit 01.0_condition_om.Rmd first.", call. = FALSE)
if (!exists("eqs"))
  eqs <- eqls[["bh1"]]

samFile <- file.path(root, "data/results/02.4.1_sam.RData")
if (!file.exists(samFile))
  stop("Missing ", samFile, " — knit 02.4.1_sam.Rmd first.", call. = FALSE)
samEnv <- new.env()
load(samFile, envir = samEnv)
for (v in c("oem", "sids", "indexCv", "nits"))
  if (!exists(v, envir = samEnv, inherits = FALSE))
    stop(samFile, " has no '", v, "' — re-knit 02.4.1_sam.Rmd.", call. = FALSE)
oem <- samEnv$oem
sids <- samEnv$sids
indexCv <- samEnv$indexCv
nits <- samEnv$nits

miss <- setdiff(sids, names(oms))
if (length(miss))
  stop("SAM stocks from 02.4.1 are not in oms: ",
       paste(miss, collapse = ", "), call. = FALSE)
if (!identical(sort(names(oem)), sort(sids)))
  stop("02.4.1_sam.RData: oem stocks do not match sids — re-knit 02.4.1.",
       call. = FALSE)
require_om_gate(root, sids)

om <- oms[sids]
startYr <- setNames(rep(2015L, length(sids)), sids)
endYr <- setNames(sapply(om, function(x) dims(x)$maxyear), names(om))
too_short <- names(endYr)[endYr < startYr]
if (length(too_short))
  stop("Terminal year is before 2015 for: ",
       paste(too_short, collapse = ", "), call. = FALSE)

srr_run <- switch(stage,
  loop = srr_om,
  srr  = srr6)
```

Stage: **`r stage`** (`r paste(srr_run, collapse = ", ")`). SAM stocks:
`r paste(sids, collapse = ", ")`. Survey `indexCv` = `r indexCv`.

# Closed loop

```{r scenario-hcr}
srDev <- lapply(eqls[srr_run], function(eq) FLQuants(lapply(eq[sids], srResiduals)))
bndTac <- hcr_tac_bounds()
bndWhen <- hcr_tac_bnd_when()
# sam.fit is run here, not on a cluster: shipping FLIndex to a worker
# has produced non-finite optim failures that do not occur in-process.
closedLoopSam <- list()
for (nm in srr_run) {
  runs <- setNames(vector("list", length(sids)), sids)
  for (id in sids) {
    message("SAM loop ", nm, " ", id)
    set.seed(1000L + match(id, sids))
    stk <- om[[id]]
    if (nits > 1L && dims(stk)$iter == 1L)
      stk <- propagate(stk, nits)
    idx <- sam_prime(stk, oem[[id]], start = startYr[id], lag = 1L)
    # hcrICES fits SAM on the stock it projects, so placeholders are set to NA
    # there and the historical OM values put back after the loop. Projection
    # years must keep finite catch.n: fwd derives the discard ratio from it.
    res <- tryCatch(
      hcrICES(sam_catch_na(stk, end = startYr[id]), eql = eqls[[nm]][[id]],
              sr_deviances = srDev[[nm]][[id]],
              params = hcrParams(eqs[[id]]),
              start = startYr[id], end = endYr[id],
              lag = 1, interval = 1,
              err = idx, implErr = 0,
              bndTac = bndTac, bndWhen = bndWhen,
              verbose = TRUE)[[1]],
      error = function(e)
        stop(nm, " / ", id, ": ", conditionMessage(e), call. = FALSE))
    pre <- ac(dims(stk)$minyear:startYr[id])
    catch.n(res)[, pre] <- catch.n(stk)[, pre]
    if (any(is.na(catch.n(res))))
      stop(nm, " / ", id, ": NA catch.n left in projected years.", call. = FALSE)
    runs[[id]] <- res
  }
  closedLoopSam[[nm]] <- FLStocks(runs)
}

save(closedLoopSam, sids, srr_run, nits, indexCv, startYr, endYr, stage,
     file = file.path(root, "data/results/02.4.2_closedLoop_sam.RData"))
```

# Trajectories

Same layout as `02.2` / `02.3`: Historical vs Advice rule (SAM MP), metrics
relative to Historical mean, facet quantity \(\times\) stock.

```{r traj-plot, fig.height=9, fig.width=11}
labs = setNames(as.character(stocks$name[match(sids, stocks$sid)]), sids)
srr_plot = if (srr_om %in% names(closedLoopSam)) srr_om else names(closedLoopSam)[1]
print(plot_hcr_vs_hist(
  om, closedLoopSam[[srr_plot]], eqs, eqls[[srr_plot]],
  sids = sids, labels = labs, start_ar = 2015L,
  advice_lab = "Advice rule"))
```

**Figure `r Fig = Fig + 1; Fig`**. SAM closed loop (`r srr_plot`) vs
historical `FLStock`.

```{r traj-plot-srr, fig.height=9, fig.width=11, eval = length(closedLoopSam) > 1L}
print(plot_hcr_vs_hist(
  om, closedLoopSam, eqs, eqls[names(closedLoopSam)],
  sids = sids, labels = labs, start_ar = 2015L))
```

```{r traj-srr-caption, results="asis", eval = length(closedLoopSam) > 1L}
Fig <- Fig + 1
cat("\n\n**Figure ", Fig,
    "**. Same contrast with one Advice-rule line per SRR.\n\n", sep = "")
```
