---
title: "All-stock digest"
subtitle: "Pick case studies after the generic backtest — not before"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

One page per knit of **every** stock that passed the Operating Model gate
and has open- and closed-loop results. Use this to choose which stocks
need a `06.1`-style case study. Do not put SAM, EqSim essays or
presentation layout here.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- data_root()
here <- clean_root()
# knitr is already loaded by rmarkdown::render; do not library() a second copy.
if (!"knitr" %in% loadedNamespaces())
  suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      error     = FALSE)
```

```{r libs}
suppressPackageStartupMessages({
  library(ggplot2)
  library(FLCore)
  library(FLBRP)
  library(plyr)
  library(FLBacktest)
})
source(file.path(here, "R/constants.R"))
source(file.path(here, "R/om.R"))
load_om(root)
require_om_gate(root, as.character(stocks$sid))
load(file.path(root, "data/results/02.1_openLoop.RData"))
load(file.path(root, "data/results/02.2_closedLoop.RData"))
load(file.path(root, "data/results/01.1_lterm_eq.RData"))
eqs = eqls[["bh1"]]
srr = srr_om
if (!srr %in% names(openLoop[["SRR"]]) || !srr %in% names(closedLoop[["SRR"]]))
  stop("openLoop/closedLoop SRR has no '", srr, "' — knit 02.1 and 04.2.",
       call. = FALSE)
if (is.null(openLoop[["OM_FMSY"]]) || !srr %in% names(openLoop[["OM_FMSY"]]))
  stop("openLoop$OM_FMSY$", srr, " missing — knit 02.1.", call. = FALSE)
```

```{r digest}
term = function(stk, yr) {
  c(SSB = c(ssb(stk)[, ac(yr)]), F = c(fbar(stk)[, ac(yr)]))
}

# Mean F over the closed-loop backtest window (2015–terminal), matching
# tex/report.tex Table results-current.
win_mean_F = function(stk, y0 = 2015L) {
  yT = dims(stk)$maxyear
  y0 = max(as.integer(dims(stk)$minyear), as.integer(y0))
  mean(c(fbar(window(stk, start = y0, end = yT))), na.rm = TRUE)
}

tab = ldply(as.character(stocks$sid), function(id) {
  if (is.null(openLoop[["SRR"]][[srr]][[id]]))
    stop("Missing openLoop$SRR$", srr, "$", id, " — knit 02.1.", call. = FALSE)
  if (is.null(closedLoop[["SRR"]][[srr]][[id]]))
    stop("Missing closedLoop$SRR$", srr, "$", id, " — knit 02.2.", call. = FALSE)
  yr = dims(oms[[id]])$maxyear
  rp_ices = refpts(eqs[[id]])
  rp_om   = refpts(eqls[[srr]][[id]])
  btrig   = c(rp_ices["MSYBtrigger", "ssb"])
  ftar    = c(rp_ices["FMSY", "harvest"])
  fmsy_om = c(rp_om["msy", "harvest"])
  h  = term(oms[[id]], yr)
  o  = term(openLoop[["SRR"]][[srr]][[id]], yr)
  c_ = term(closedLoop[["SRR"]][[srr]][[id]], yr)
  if (is.null(openLoop[["OM_FMSY"]][[srr]][[id]]))
    stop("Missing openLoop$OM_FMSY$", srr, "$", id, " — knit 02.1.", call. = FALSE)
  om = term(openLoop[["OM_FMSY"]][[srr]][[id]], yr)
  data.frame(
    sid = id,
    name = as.character(stocks$name[match(id, stocks$sid)]),
    year = yr,
    gate = {
      p = lterm_eq$pass
      isTRUE(p$pass[match(id, as.character(p$sid))])
    },
    Ftar       = ftar,
    Fmsy_om    = fmsy_om,
    hist_SSB_Btrig = h["SSB"] / btrig,
    ftar_SSB_Btrig = o["SSB"] / btrig,
    omF_SSB_Btrig  = om["SSB"] / btrig,
    ar_SSB_Btrig   = c_["SSB"] / btrig,
    hist_F         = win_mean_F(oms[[id]]),
    ftar_F         = win_mean_F(openLoop[["SRR"]][[srr]][[id]]),
    omF_F          = win_mean_F(openLoop[["OM_FMSY"]][[srr]][[id]]),
    ar_F           = win_mean_F(closedLoop[["SRR"]][[srr]][[id]]),
    hist_F_Ftar    = win_mean_F(oms[[id]]) / ftar,
    ftar_F_Ftar    = win_mean_F(openLoop[["SRR"]][[srr]][[id]]) / ftar,
    omF_F_Ftar     = win_mean_F(openLoop[["OM_FMSY"]][[srr]][[id]]) / ftar,
    ar_F_Ftar      = win_mean_F(closedLoop[["SRR"]][[srr]][[id]]) / ftar,
    stringsAsFactors = FALSE)
})
save(tab, srr,
     file = file.path(root, "data/results/03.0_digest.RData"))
```

**Table 1**. Terminal $SSB/B_{\mathrm{trigger}}$ and mean $F$ over 2015–terminal
versus ICES $F_{\mathrm{target}}$ and OM $F_{\mathrm{MSY}}$. Reference OM branch
`r srr`. Case studies: Celtic/Irish Sea cod and whiting.

```{r table}
cols = c("name", "sid", "year", "gate",
         "hist_SSB_Btrig", "ftar_SSB_Btrig", "omF_SSB_Btrig", "ar_SSB_Btrig",
         "hist_F", "ftar_F", "omF_F", "ar_F",
         "Ftar", "Fmsy_om")
knitr::kable(
  tab[, cols],
  digits = 2, row.names = FALSE,
  col.names = c("Stock", "SID", "Year", "Gate",
                "Hist SSB/$B_{trig}$",
                "ICES $F_{tar}$ SSB/$B_{trig}$",
                "OM $F_{MSY}$ SSB/$B_{trig}$",
                "AR SSB/$B_{trig}$",
                "Hist $\\bar F$",
                "ICES $F_{tar}$ $\\bar F$",
                "OM $F_{MSY}$ $\\bar F$",
                "AR $\\bar F$",
                "ICES $F_{tar}$", "OM $F_{MSY}$"))
```
