---
title: "Paper figures (funder draft)"
subtitle: "Loads saved results — does not re-run the MSE"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

Contract / funder figures from saved objects (`bh3` OM; ICES HCR inputs on
`bh1`). The published trajectory figure is
[`scripts/main_traj_facet.R`](../scripts/main_traj_facet.R)
(plain ggplot2 facet). This notebook is a short digest plus
`FLCore::plot` of the four case-study `FLStock`s.

`FLCore` exports `plot` for `FLStock` / `FLStocks` (lattice). It does **not**
export `ggplot`; paper facets use ggplot2.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- data_root()
here <- clean_root()
dir.create(file.path(here, "tex/figs"), recursive = TRUE, showWarnings = FALSE)
# knitr is already loaded by rmarkdown::render; do not library() a second copy.
if (!"knitr" %in% loadedNamespaces())
  suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  error     = FALSE,
  cache     = FALSE,
  fig.path  = file.path(here, "tex/figs", "04.0_paper-"),
  fig.width = 9,
  fig.height= 5,
  dev       = "png")
```

```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(ggplot2)
  library(FLBRP)
  library(plyr)
  library(FLBacktest)
})
source(file.path(here, "R/constants.R"))
source(file.path(here, "R/om.R"))
theme_set(theme_paper())
load_om(root)
require_om_gate(root, as.character(stocks$sid))
load(file.path(root, "data/results/02.1_openLoop.RData"))
load(file.path(root, "data/results/02.2_closedLoop.RData"))
load(file.path(root, "data/results/02.5_rebuild.RData"))
eqs <- eqls[["bh1"]]
if (!srr_om %in% names(openLoop[["SRR"]]) ||
    !srr_om %in% names(closedLoop[["SRR"]]) ||
    is.null(openLoop[["OM_FMSY"]][[srr_om]]))
  stop("Need openLoop$SRR / OM_FMSY and closedLoop$SRR for ", srr_om,
       call. = FALSE)
```

# Case-study stocks (FLCore::plot)

```{r flcore-plot, fig.height=8, fig.width=9}
oms4 <- FLStocks(oms[case_sids])
print(FLCore::plot(oms4))
```

Historical assessment series for the four gadoids (lattice `FLCore::plot`).

```{r flcore-arms, fig.height=8, fig.width=9}
id <- case_sids[[1]]
print(FLCore::plot(FLStocks(
  Historical = oms[[id]],
  Ftar       = openLoop[["SRR"]][[srr_om]][[id]],
  OM_FMSY    = openLoop[["OM_FMSY"]][[srr_om]][[id]],
  Advice     = closedLoop[["SRR"]][[srr_om]][[id]]
)))
```

Same four arms for `r case_labs[[1]]`.

# Terminal status (ggplot2)

```{r digest-bars, fig.height=5, fig.width=9}
term_ssb <- function(stk) {
  yr <- dims(stk)$maxyear
  c(ssb(stk)[, ac(yr)])
}
tab <- ldply(case_sids, function(id) {
  btrig <- c(refpts(eqs[[id]])["MSYBtrigger", "ssb"])
  data.frame(
    stock = unname(case_labs[id]),
    arm = c("Historical", "ICES Ftar", "OM FMSY", "Advice rule"),
    SSB_Btrig = c(
      term_ssb(oms[[id]]),
      term_ssb(openLoop[["SRR"]][[srr_om]][[id]]),
      term_ssb(openLoop[["OM_FMSY"]][[srr_om]][[id]]),
      term_ssb(closedLoop[["SRR"]][[srr_om]][[id]])
    ) / btrig,
    stringsAsFactors = FALSE)
})
tab$arm <- factor(tab$arm, levels = c("Historical", "ICES Ftar", "OM FMSY",
                                      "Advice rule"))
tab$stock <- factor(tab$stock, levels = unname(case_labs[case_sids]))
ggplot(tab, aes(arm, SSB_Btrig, fill = arm)) +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey40") +
  geom_col(width = 0.7) +
  facet_wrap(~ stock, nrow = 1) +
  scale_fill_manual(values = paper_cols[c("Historical", "ICES Ftar",
                                          "OM FMSY", "Advice rule")],
                    guide = "none") +
  labs(x = NULL, y = expression(SSB / B[trigger])) +
  theme_paper() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Dashed line is \(MSYB_{\mathrm{trigger}}\). The full Historical / Advice-rule
time series (plus Future) is `tex/figs/traj_facet.pdf` from
`scripts/main_traj_facet.R`.
