---
title: "Multi-SRR sensitivity for peer-review revision"
subtitle: "Four case-study stocks — review notebook only (not part of the funder draft)"
author: "Laurence Kell, Sea++"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: flatly
---

# Purpose

> **Not the funder paper.** The contract / funder draft (`tex/report.tex`,
> abstract) stays on the **bh3** reference Operating Model and single-OM
> tables. This notebook loads saved pipeline objects and puts **all** SRR
> branches side by side for the four Paper-1 case studies, so you can decide
> what (if anything) to promote into a peer-review revision.

Arms compared (Historical window, 2015–terminal for mean $F$; terminal year
for SSB ratios):

| Label | Source |
|-------|--------|
| Historical | `oms` (assessment) |
| ICES $F_{\mathrm{tar}}$ | `openLoop$SRR` — constant published SAG $F_{\mathrm{MSY}}$ |
| OM $F_{\mathrm{MSY}}$ | `openLoop$OM_FMSY` — scenario's own equilibrium $F_{\mathrm{MSY}}$ |
| Advice rule | `closedLoop$SRR` — `hcrICES` from 2015 (`bndTac`) |

ICES HCR inputs ($F_{\mathrm{tar}}$, $B_{\lim}$, $MSYB_{\mathrm{trigger}}$)
remain **bh1** / SAG for every SRR branch. Operating Model $B_{\mathrm{MSY}}$
and $F_{\mathrm{MSY}}$ are performance metrics from `eqls[[srr]]`.

```{r knitr, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE}
source(if (file.exists("../R/paths.R")) "../R/paths.R" else "R/paths.R")
root <- data_root()
here <- clean_root()

dir.create(file.path(here, "tex/figs"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(here, "Rmd/cache", "04.1_multiSRR_review"), recursive = TRUE,
           showWarnings = FALSE)

# knitr is already loaded by rmarkdown::render; do not library() a second copy.
if (!"knitr" %in% loadedNamespaces())
  suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
  echo       = FALSE,
  warning    = FALSE,
  message    = FALSE,
  error     = FALSE,
  cache      = FALSE,
  cache.path = file.path(here, "Rmd/cache", "04.1_multiSRR_review/"),
  fig.path   = file.path(here, "tex/figs", "04.1_multiSRR_review-"),
  fig.width  = 10,
  fig.height = 5,
  dev        = "png")

Fig = 0
Tab = 0
```

```{r libs}
suppressPackageStartupMessages({
  library(FLCore)
  library(ggplot2)
  library(FLBRP)
  library(plyr)
})

source(file.path(here, "R/constants.R"))
theme_set(theme_paper())
update_geom_defaults("line", list(linewidth = 0.85))

cols_pol = c(
  Historical          = "#000000",
  "ICES Ftar"         = "#0072B2",
  "OM FMSY"           = "#009E73",
  "Advice rule"       = "#D55E00")
```

```{r load}
load(file.path(root, "data/om/oms.RData"))
load(file.path(root, "data/om/eqls.RData"))
load(file.path(root, "data/results/02.1_openLoop.RData"))
load(file.path(root, "data/results/02.2_closedLoop.RData"))

rebuild_path = file.path(root, "data/results/02.5_rebuild.RData")
rebuild_ok   = file.exists(rebuild_path)
rebuild_srr  = NA_character_
if (rebuild_ok) {
  # Isolate rebuild objects so they do not overwrite open/closed loop.
  rb_env = new.env(parent = emptyenv())
  load(rebuild_path, envir = rb_env)
  rebuild_srr = if (exists("srr", envir = rb_env, inherits = FALSE))
    as.character(get("srr", envir = rb_env)) else NA_character_
  rebuild_names = ls(envir = rb_env)
}

case_sids = c(
  "cod.27.7e-k",
  "whg.27.7a",
  "cod.27.7a",
  "whg.27.7b-ce-k")
case_labs = c(
  "cod.27.7e-k"    = "Celtic Sea cod",
  "whg.27.7a"      = "Irish Sea whiting",
  "cod.27.7a"      = "Irish Sea cod",
  "whg.27.7b-ce-k" = "Celtic Sea whiting")

# Prefer stocks.csv labels when available; fall back to case_labs.
case_name = setNames(vapply(case_sids, function(id) {
  i = match(id, as.character(stocks$sid))
  if (!is.na(i)) as.character(stocks$name[i]) else unname(case_labs[id])
}, character(1L)), case_sids)

eqs = eqls[["bh1"]]
srr6 = c("bh1", "bh2", "bh3", "rk1", "rk2", "rk3")
srr_all = intersect(srr6, names(eqls))
miss_srr = setdiff(srr6, srr_all)
if (length(miss_srr))
  stop("eqls is missing ", paste(miss_srr, collapse = ", "), call. = FALSE)
# Diagnostics scn0–scn2 stay in eqls but are not open-/closed-loop SRRs.
startYr_ar = 2015L
endYr = setNames(vapply(case_sids, function(id)
  as.integer(dims(oms[[id]])$maxyear), integer(1L)), case_sids)

if (is.null(openLoop[["OM_FMSY"]]))
  stop("openLoop$OM_FMSY missing — knit 02.1.", call. = FALSE)
has_om_fmsy = TRUE
```

# Inventory

```{r inventory}
inv_row = function(nm) {
  ol  = nm %in% names(openLoop[["SRR"]]) &&
    all(case_sids %in% names(openLoop[["SRR"]][[nm]]))
  omf = has_om_fmsy && nm %in% names(openLoop[["OM_FMSY"]]) &&
    all(case_sids %in% names(openLoop[["OM_FMSY"]][[nm]]))
  cl  = nm %in% names(closedLoop[["SRR"]]) &&
    all(case_sids %in% names(closedLoop[["SRR"]][[nm]]))
  miss_cl = if (nm %in% names(closedLoop[["SRR"]]))
    setdiff(case_sids, names(closedLoop[["SRR"]][[nm]])) else case_sids
  data.frame(
    SRR = nm,
    eqls = nm %in% names(eqls),
    openLoop_SRR = ol,
    openLoop_OM_FMSY = omf,
    closedLoop_SRR = cl,
    missing_closed_case = if (length(miss_cl)) paste(miss_cl, collapse = ", ")
      else "—",
    stringsAsFactors = FALSE)
}
inv = ldply(srr_all, inv_row)

n_ol  = sum(inv$openLoop_SRR)
n_omf = sum(inv$openLoop_OM_FMSY)
n_cl  = sum(inv$closedLoop_SRR)
miss_any_cl = inv$SRR[!inv$closedLoop_SRR]
```

**Operating Model branches in `eqls`:** `r paste(srr_all, collapse = ", ")`
(`r length(srr_all)` SRRs).

**Case studies:** `r paste(sprintf("%s (%s)", case_name, case_sids), collapse = "; ")`.

| Object | Case-study coverage |
|--------|---------------------|
| `eqls` | `r length(srr_all)` / `r length(srr_all)` SRRs |
| `openLoop$SRR` (ICES $F_{\mathrm{tar}}$) | `r n_ol` / `r length(srr_all)` |
| `openLoop$OM_FMSY` | `r n_omf` / `r length(srr_all)` |
| `closedLoop$SRR` (Advice rule) | `r n_cl` / `r length(srr_all)` |
| `02.5_rebuild` | `r if (rebuild_ok) paste0("present; SRR = ", ifelse(is.na(rebuild_srr), "unknown", rebuild_srr)) else "file missing"` |

```{r inventory-table}
knitr::kable(
  inv,
  caption = paste0("Table ", Tab <<- Tab + 1,
                   ". SRR inventory for the four case studies."),
  row.names = FALSE)
```

```{r missing-cl-note, results='asis'}
if (length(miss_any_cl)) {
  cat("**Missing closed-loop SRRs for case stocks:** ",
      paste(miss_any_cl, collapse = ", "), ".\n", sep = "")
} else {
  cat("**No missing closed-loop SRRs** among the four case studies;",
      " all branches in `eqls` have Advice-rule trajectories.\n")
}
```

```{r rebuild-note, results='asis'}
rb_lab = if (isTRUE(rebuild_srr == "bh3")) {
  "**bh3-only**"
} else if (rebuild_ok) {
  paste0("SRR = ", rebuild_srr)
} else {
  "unavailable"
}
cat("**Rebuild note.** `02.5_rebuild.RData` is ", rb_lab, ".\n", sep = "")
```

Future / years-to-rebuild comparisons across SRRs are **not** available here.
This notebook therefore focuses on **Historical-window** summaries and
trajectories from open- and closed-loop objects. Open loop always stores all
six SRRs; closed loop defaults to **bh3** (`02.2`). For full closed-loop
coverage re-knit `02.2` with `BM_CLOSED_SRR=all` before this review.
Extending rebuild to every SRR would mean re-knitting
[`02.5_rebuild.Rmd`](02.5_rebuild.Rmd) in a multi-SRR loop; do that only if
peer review needs Future panels by SRR.

# Summary tables

Metrics match `03.0_digest` / `scripts/compute_openloop_metrics.R`:

- **Terminal** $SSB / MSYB_{\mathrm{trigger}}$ (ICES trigger from `eqls$bh1`)
- **Mean** $\bar F$ and $\bar F / F_{\mathrm{tar}}$ over **2015–terminal**
- OM $F_{\mathrm{MSY}}$ vs ICES $F_{\mathrm{tar}}$; $B_{\mathrm{MSY}} / B_{\mathrm{trigger}}$ per SRR
- Optional $\Delta$ (Advice − Historical) in $SSB / B_{\mathrm{trigger}}$

```{r metrics}
term_ssb = function(stk) {
  yr = dims(stk)$maxyear
  x  = ssb(stk)[, ac(yr)]
  if (dims(x)$iter > 1) c(iterMedians(x)) else c(x)
}

win_mean_F = function(stk, y0 = startYr_ar) {
  yT = dims(stk)$maxyear
  y0 = max(as.integer(dims(stk)$minyear), as.integer(y0))
  fb = fbar(window(stk, start = y0, end = yT))
  if (dims(fb)$iter > 1) mean(c(iterMedians(fb)), na.rm = TRUE)
  else mean(c(fb), na.rm = TRUE)
}

arm_metrics = function(stk, btrig, ftar) {
  if (is.null(stk))
    stop("arm_metrics: NULL stock", call. = FALSE)
  Fb = win_mean_F(stk)
  c(SSB_Btrig = term_ssb(stk) / btrig,
    F         = Fb,
    F_Ftar    = Fb / ftar)
}

sum_tab = ldply(srr_all, function(nm) {
  ldply(case_sids, function(id) {
    btrig = c(refpts(eqs[[id]])["MSYBtrigger", "ssb"])
    ftar  = c(refpts(eqs[[id]])["FMSY", "harvest"])
    rp_om = refpts(eqls[[nm]][[id]])
    fmsy  = c(rp_om["msy", "harvest"])
    bmsy  = c(rp_om["msy", "ssb"])

    h  = arm_metrics(oms[[id]], btrig, ftar)
    o  = arm_metrics(openLoop[["SRR"]][[nm]][[id]], btrig, ftar)
    if (!has_om_fmsy || is.null(openLoop[["OM_FMSY"]][[nm]][[id]]))
      stop("Missing openLoop$OM_FMSY$", nm, "$", id, call. = FALSE)
    if (is.null(closedLoop[["SRR"]][[nm]][[id]]))
      stop("Missing closedLoop$SRR$", nm, "$", id, call. = FALSE)
    om = arm_metrics(openLoop[["OM_FMSY"]][[nm]][[id]], btrig, ftar)
    a  = arm_metrics(closedLoop[["SRR"]][[nm]][[id]], btrig, ftar)

    data.frame(
      SRR = nm, sid = id, stock = unname(case_name[id]),
      year = endYr[id],
      Ftar = ftar, Fmsy_om = fmsy,
      Fmsy_over_Ftar = fmsy / ftar,
      Bmsy_over_Btrig = bmsy / btrig,
      hist_SSB_Btrig = h["SSB_Btrig"],
      ftar_SSB_Btrig = o["SSB_Btrig"],
      omF_SSB_Btrig  = om["SSB_Btrig"],
      ar_SSB_Btrig   = a["SSB_Btrig"],
      hist_F = h["F"], ftar_F = o["F"], omF_F = om["F"], ar_F = a["F"],
      hist_F_Ftar = h["F_Ftar"], ftar_F_Ftar = o["F_Ftar"],
      omF_F_Ftar = om["F_Ftar"], ar_F_Ftar = a["F_Ftar"],
      delta_SSB_Btrig = a["SSB_Btrig"] - h["SSB_Btrig"],
      stringsAsFactors = FALSE)
  })
})

sum_tab$stock = factor(sum_tab$stock, levels = unname(case_name[case_sids]))
sum_tab$SRR   = factor(sum_tab$SRR, levels = srr_all)
sum_tab = sum_tab[order(sum_tab$stock, sum_tab$SRR), ]
rownames(sum_tab) = NULL

save(sum_tab, inv, case_sids, srr_all, rebuild_srr,
     file = file.path(root, "data/results/04.1_multiSRR_review.RData"))
```

## Terminal SSB / trigger and mean $F$ (2015–terminal)

Wide table: one row per stock × SRR.

```{r table-main, results='asis'}
fmt2 = function(x) sprintf("%.2f", x)
show = sum_tab[, c(
  "stock", "SRR",
  "hist_SSB_Btrig", "ftar_SSB_Btrig", "omF_SSB_Btrig", "ar_SSB_Btrig",
  "hist_F", "ftar_F", "omF_F", "ar_F",
  "hist_F_Ftar", "ftar_F_Ftar", "omF_F_Ftar", "ar_F_Ftar",
  "delta_SSB_Btrig")]
knitr::kable(
  show, digits = 2, row.names = FALSE,
  col.names = c(
    "Stock", "SRR",
    "Hist SSB/$B_{trig}$", "ICES $F_{tar}$ SSB/$B_{trig}$",
    "OM $F_{MSY}$ SSB/$B_{trig}$", "AR SSB/$B_{trig}$",
    "Hist $\\bar F$", "ICES $F_{tar}$ $\\bar F$",
    "OM $F_{MSY}$ $\\bar F$", "AR $\\bar F$",
    "Hist $F/F_{tar}$", "ICES $F_{tar}$ $F/F_{tar}$",
    "OM $F_{MSY}$ $F/F_{tar}$", "AR $F/F_{tar}$",
    "$\\Delta$ AR−Hist"),
  caption = paste0(
    "Table ", Tab <<- Tab + 1,
    ". Terminal SSB/$B_{trigger}$ and mean $\\bar F$ (2015–terminal) ",
    "across SRRs. Historical SSB/$B_{trigger}$ does not depend on the OM SRR ",
    "(same assessment); $\\bar F$ likewise. Counterfactual columns do."))
```

## OM $F_{\mathrm{MSY}}$ vs ICES $F_{\mathrm{tar}}$; $B_{\mathrm{MSY}}/B_{\mathrm{trigger}}$

```{r table-refpts}
rp_show = sum_tab[, c(
  "stock", "SRR", "Ftar", "Fmsy_om", "Fmsy_over_Ftar", "Bmsy_over_Btrig")]
knitr::kable(
  rp_show, digits = 3, row.names = FALSE,
  col.names = c(
    "Stock", "SRR",
    "ICES $F_{tar}$", "OM $F_{MSY}$",
    "OM $F_{MSY}$ / ICES $F_{tar}$",
    "$B_{MSY}$ / $B_{trigger}$"),
  caption = paste0(
    "Table ", Tab <<- Tab + 1,
    ". Published ICES $F_{tar}$ (fixed) versus Operating Model ",
    "$F_{MSY}$ and $B_{MSY}/B_{trigger}$ by SRR."))
```

```{r heat-delta, fig.height=4.5, fig.width=9}
p_d = ggplot(sum_tab, aes(x = SRR, y = stock, fill = delta_SSB_Btrig)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(aes(label = sprintf("%+.1f", delta_SSB_Btrig)), size = 3) +
  scale_fill_gradient2(
    low = "#D55E00", mid = "#ecf0f1", high = "#0072B2", midpoint = 0,
    name = expression(Delta ~ SSB / B[trig])) +
  labs(x = NULL, y = NULL,
       title = expression(
         Delta ~ (Advice ~ rule - Historical) ~ SSB / B[trigger])) +
  theme_paper() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right")
print(p_d)
```

**Figure `r Fig = Fig + 1; Fig`**. $\Delta$ terminal $SSB/B_{\mathrm{trigger}}$
(Advice rule − Historical). Blue = Advice rule ends higher; orange = lower.

```{r heat-fmsy, fig.height=4.5, fig.width=9}
p_f = ggplot(sum_tab, aes(x = SRR, y = stock, fill = log10(Fmsy_over_Ftar))) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(aes(label = sprintf("%.2f", Fmsy_over_Ftar)), size = 3) +
  scale_fill_gradient2(
    low = "#0072B2", mid = "#ecf0f1", high = "#D55E00", midpoint = 0,
    name = expression(log[10](F[MSY]/F[tar]))) +
  labs(x = NULL, y = NULL,
       title = expression(OM ~ F[MSY] / ICES ~ F[tar] ~ "(labels = ratio)")) +
  theme_paper() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right")
print(p_f)
```

**Figure `r Fig = Fig + 1; Fig`**. OM $F_{\mathrm{MSY}}$ / ICES $F_{\mathrm{tar}}$
by stock × SRR (tile colour on $\log_{10}$ scale; labels are the ratio).

# Trajectories

One stock per figure; columns = SRR; colour = policy. $Y$-scales free within
each stock. Window starts in 2010 for readability (Advice rule still peels off
in 2015). Catch is included as a third panel set below.

```{r traj-data, results='hide'}
as_ts = function(stk, qfun, start = 2010L) {
  yT = dims(stk)$maxyear
  y0 = max(as.integer(dims(stk)$minyear), as.integer(start))
  x  = qfun(window(stk, start = y0, end = yT))
  if (dims(x)$iter > 1) x = iterMedians(x)
  df = as.data.frame(x)
  df[, c("year", "data")]
}

build_traj = function(qfun, qname) {
  ldply(case_sids, function(id) {
    ldply(srr_all, function(nm) {
      if (is.null(openLoop[["SRR"]][[nm]][[id]]))
        stop("Missing openLoop$SRR$", nm, "$", id, call. = FALSE)
      if (!has_om_fmsy || is.null(openLoop[["OM_FMSY"]][[nm]][[id]]))
        stop("Missing openLoop$OM_FMSY$", nm, "$", id, call. = FALSE)
      if (is.null(closedLoop[["SRR"]][[nm]][[id]]))
        stop("Missing closedLoop$SRR$", nm, "$", id, call. = FALSE)
      arms = list(
        Historical = oms[[id]],
        "ICES Ftar" = openLoop[["SRR"]][[nm]][[id]],
        "OM FMSY" = openLoop[["OM_FMSY"]][[nm]][[id]],
        "Advice rule" = closedLoop[["SRR"]][[nm]][[id]])
      ldply(names(arms), function(pol) {
        stk = arms[[pol]]
        if (is.null(stk))
          stop("NULL stock for ", pol, " / ", nm, " / ", id, call. = FALSE)
        df = as_ts(stk, qfun)
        df$sid = id
        df$stock = unname(case_name[id])
        df$SRR = nm
        df$policy = pol
        df$qname = qname
        df
      })
    })
  })
}

traj_ssb = build_traj(ssb, "SSB")
traj_f   = build_traj(fbar, "F")
traj_c   = build_traj(catch, "Catch")
traj = rbind(traj_ssb, traj_f, traj_c)
traj$stock  = factor(traj$stock, levels = unname(case_name[case_sids]))
traj$SRR    = factor(traj$SRR, levels = srr_all)
traj$policy = factor(traj$policy, levels = names(cols_pol))
traj$qname  = factor(traj$qname, levels = c("SSB", "F", "Catch"))
```

```{r traj-ssb, fig.height=8, fig.width=12}
for (id in case_sids) {
  nm = unname(case_name[id])
  d  = subset(traj, sid == id & qname == "SSB")
  btrig = c(refpts(eqs[[id]])["MSYBtrigger", "ssb"])
  p = ggplot(d, aes(year, data, colour = policy)) +
    geom_hline(yintercept = btrig, linetype = 2, colour = bm_col("trigger"),
               linewidth = 0.4) +
    geom_line(na.rm = TRUE) +
    facet_wrap(~ SRR, scales = "free_y", ncol = 3) +
    scale_colour_manual(values = cols_pol) +
    labs(title = paste(nm, "— SSB"), x = "Year", y = "SSB",
         subtitle = "Dashed = ICES MSY Btrigger (fixed)") +
    theme_paper()
  print(p)
}
```

**Figure `r Fig = Fig + 1; Fig`**. SSB trajectories by SRR (facets) for each
case-study stock. Black Historical, blue ICES $F_{\mathrm{tar}}$, green OM
$F_{\mathrm{MSY}}$, orange Advice rule.

```{r traj-f, fig.height=8, fig.width=12}
for (id in case_sids) {
  nm = unname(case_name[id])
  d  = subset(traj, sid == id & qname == "F")
  ftar = c(refpts(eqs[[id]])["FMSY", "harvest"])
  p = ggplot(d, aes(year, data, colour = policy)) +
    geom_hline(yintercept = ftar, linetype = 2, colour = bm_col("f_target"),
               linewidth = 0.4) +
    geom_line(na.rm = TRUE) +
    facet_wrap(~ SRR, scales = "free_y", ncol = 3) +
    scale_colour_manual(values = cols_pol) +
    labs(title = paste(nm, "— Fbar"), x = "Year", y = expression(bar(F)),
         subtitle = "Dashed = ICES Ftar (fixed)") +
    theme_paper()
  print(p)
}
```

**Figure `r Fig = Fig + 1; Fig`**. $\bar F$ trajectories by SRR for each stock.

```{r traj-ssb-grid, fig.height=7, fig.width=11}
d_ssb = subset(traj, qname == "SSB")
p_grid = ggplot(d_ssb, aes(year, data, colour = policy)) +
  geom_line(linewidth = 0.55, na.rm = TRUE) +
  facet_grid(stock ~ SRR, scales = "free_y") +
  scale_colour_manual(values = cols_pol) +
  labs(title = "SSB — stock × SRR overview", x = "Year", y = "SSB") +
  theme_paper() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = rel(0.7)),
        strip.text = element_text(size = rel(0.75)))
print(p_grid)
```

**Figure `r Fig = Fig + 1; Fig`**. SSB overview grid (stocks × SRRs).

# Skim findings

Auto-generated bullets from the summary table (read the heatmaps and tables
above before promoting any into the paper).

```{r findings, results='asis'}
# Sign of Advice − Historical at bh3 (reference) vs other SRRs
bh3 = subset(sum_tab, SRR == "bh3")
flip = ldply(case_sids, function(id) {
  base_sign = sign(bh3$delta_SSB_Btrig[bh3$sid == id])
  sub = subset(sum_tab, sid == id)
  flips = as.character(sub$SRR[sign(sub$delta_SSB_Btrig) != base_sign &
                                 !is.na(sub$delta_SSB_Btrig)])
  data.frame(
    stock = unname(case_name[id]),
    bh3_delta = bh3$delta_SSB_Btrig[bh3$sid == id],
    flip_srrs = if (length(flips)) paste(flips, collapse = ", ") else "none",
    stringsAsFactors = FALSE)
})

hi = subset(sum_tab, Fmsy_over_Ftar >= 2)
lo = subset(sum_tab, Fmsy_over_Ftar <= 0.5)
extreme = subset(sum_tab, Fmsy_over_Ftar >= 5 | Fmsy_over_Ftar <= 0.2)

cat("- **Sign of Δ (Advice − Historical) vs bh3.** ")
cat("Where another SRR flips the sign of the bh3 Δ:\n")
for (i in seq_len(nrow(flip))) {
  cat(sprintf("  - %s (bh3 Δ = %+.2f): %s\n",
              flip$stock[i], flip$bh3_delta[i], flip$flip_srrs[i]))
}

cat("\n- **OM $F_{MSY} \\gg$ ICES $F_{tar}$ (ratio ≥ 2):**\n")
if (!nrow(hi)) cat("  - none\n") else {
  for (i in seq_len(nrow(hi)))
    cat(sprintf("  - %s / %s: ratio = %.2f (OM $F_{MSY}$ = %.3f, $F_{tar}$ = %.3f)\n",
                hi$stock[i], hi$SRR[i], hi$Fmsy_over_Ftar[i],
                hi$Fmsy_om[i], hi$Ftar[i]))
}

cat("\n- **OM $F_{MSY} \\ll$ ICES $F_{tar}$ (ratio ≤ 0.5):**\n")
if (!nrow(lo)) cat("  - none\n") else {
  for (i in seq_len(nrow(lo)))
    cat(sprintf("  - %s / %s: ratio = %.2f\n",
                lo$stock[i], lo$SRR[i], lo$Fmsy_over_Ftar[i]))
}

cat("\n- **Extreme mismatches (ratio ≥ 5 or ≤ 0.2) — treat with caution in revision:**\n")
if (!nrow(extreme)) cat("  - none\n") else {
  for (i in seq_len(nrow(extreme)))
    cat(sprintf("  - %s / %s: OM $F_{MSY}/F_{tar}$ = %.2f; OM $F_{MSY}$ arm terminal SSB/$B_{trig}$ = %.2f\n",
                extreme$stock[i], extreme$SRR[i], extreme$Fmsy_over_Ftar[i],
                extreme$omF_SSB_Btrig[i]))
}

# Irish Sea cod pattern (Historical F << Ftar) across SRRs
isc = subset(sum_tab, sid == "cod.27.7a")
cat("\n- **Irish Sea cod** (Historical $\\bar F \\ll F_{tar}$): Advice-rule ")
cat("terminal SSB/$B_{trig}$ ranges ")
cat(sprintf("%.2f–%.2f across SRRs; ICES $F_{tar}$ arm %.2f–%.2f; ",
            min(isc$ar_SSB_Btrig, na.rm = TRUE),
            max(isc$ar_SSB_Btrig, na.rm = TRUE),
            min(isc$ftar_SSB_Btrig, na.rm = TRUE),
            max(isc$ftar_SSB_Btrig, na.rm = TRUE)))
cat(sprintf("Δ (AR−Hist) is negative for %d / %d SRRs.\n",
            sum(isc$delta_SSB_Btrig < 0, na.rm = TRUE), nrow(isc)))

# Depleted gadoids that gain under Advice
gain = subset(sum_tab, sid %in% c("cod.27.7e-k", "whg.27.7a") &
                delta_SSB_Btrig > 0)
cat("\n- **Celtic Sea cod & Irish Sea whiting:** Advice rule raises ")
cat(sprintf("SSB/$B_{trig}$ relative to Historical for %d / %d stock×SRR cells.\n",
            nrow(gain),
            sum(sum_tab$sid %in% c("cod.27.7e-k", "whg.27.7a"))))
```

# Peer-review checklist (do not edit report.tex from here)

What would need updating **if** multi-SRR material is promoted into a
peer-review revision. Checklist only — **do not** rewrite
`tex/report.tex` / funder Results in this notebook.

1. **Paper tables** — decide whether Results keep a single reference OM
   (bh3 / scn2) with multi-SRR in SI, or promote a compact stock × SRR
   summary (Tables above / `data/results/04.1_multiSRR_review.RData`).
2. **Figures** — pick 1–2 readable trajectory panels (e.g. SSB grid or one
   stock’s SRR facets); avoid dumping all nine SRRs into the main text.
3. **Reference OM statement** — keep the funder draft on bh3; for the
   journal draft, state clearly which OM is narrative and which are
   sensitivity.
4. **`docs/paper_outline.md` Next before submission, item i / related** —
   freeze which OM appears in main figures; note multi-SRR sensitivity
   lives in this notebook (`04.1`) rather than rewriting funder Results.
5. **Rebuild / Future** — only if reviewers ask: extend `02.5` beyond bh3
   and add years-to-trigger / years-to-$B_{\mathrm{MSY}}$ by SRR.
6. **Extreme SRRs** — flag or drop pathological OM $F_{\mathrm{MSY}}$
   branches (see skim findings) before any SI table goes to press.
7. **Do not** change `tex/report.tex` or `abstract.tex` for multi-SRR
   unless / until the funder paper is intentionally revised.

Saved metrics: `r file.path("data/results", "04.1_multiSRR_review.RData")`.
Source notebooks: [`02.1_openLoop.Rmd`](02.1_openLoop.Rmd),
[`02.2_closedLoop.Rmd`](02.2_closedLoop.Rmd),
[`03.0_digest.Rmd`](03.0_digest.Rmd),
[`04.0_paper_figures.Rmd`](04.0_paper_figures.Rmd).
